### 2014-03-31 main.tex
temp commit

\documentclass[a4paper]{article}

\usepackage{authblk}

% TODO this suggests only one result: the improved lower bound
%      reference to other results?
\title{A Kochen-Specker Vector System has at least 22 vertices}

\author{Bas Westerbaan}
\author{Sander Uijlen}
\affil{Institute for Computing and Information Sciences\\
       Radboud Universiteit Nijmegen\\
   \{\texttt{bwesterb},\texttt{suijlen}\}\texttt{@cs.ru.nl}}

\begin{document}

\maketitle

\begin{abstract}
    TODO
\end{abstract}

\section{Introduction}
TODO

\section{The 33 vector system}
TODO

\section{An improved lower bound}
TODO

\section{Embeddability}
TODO

\end{document}

% vim: ft=tex.latex

### 2014-03-31 code/helper.py

@@ -1,0 +1,26 @@
+def load_graph6(s):
+    """ Loads a graph from graph6 format """
+    c = s[0]
+    if not 63 <= ord(c) and ord(c) <= 63 + 62:
+        raise NotImplementedError
+    n = ord(c) - 63
+    bitmap = []
+    for i in xrange(1, len(s)):
+        for j in xrange(6):
+            bitmap.append(((ord(s[i]) - 63) >> (5 - j)) & 1 == 1)
+    x = 0
+    y = 1
+    g = {}
+    for i in xrange(n):
+        g[i] = set()
+    for value in bitmap:
+        if value:
+            g[x].add(y)
+            g[y].add(x)
+        x += 1
+        if x == y:
+            x = 0
+            y += 1
+            if y == n:
+                break
+    return g

### 2014-03-31 code/plot.py

@@ -1,0 +1,8 @@
+from helper import load_graph6
+
+def plot(g):
+    print g
+
+if __name__ == '__main__':
+    with open('../graphs/ks33.g6') as f:
+        plot(load_graph6(f.readline()[:-1]))

### 2014-04-02 code/helper.py
helper: add write_graph6

@@ -1,3 +1,30 @@
+def write_graph6(g):
+    """ Writes a graph to graph6 format """
+    ret = ''
+    if len(g) > 62:
+        raise NotImplementedError
+    keys = list(g)
+    ret += chr(len(g) + 63)
+    bitmap = []
+    x = 0
+    y = 1
+    while True:
+        bitmap.append(int(keys[y] in g[keys[x]]))
+        x += 1
+        if x == y:
+            x = 0
+            y += 1
+            if y == len(g):
+                break
+    for i in xrange(0, len(bitmap), 6):
+        v = 0
+        for m in xrange(6):
+            v *= 2
+            if i + m < len(bitmap):
+                v += bitmap[i + m]
+        ret += chr(v + 63)
+    return ret
+
 def load_graph6(s):
     """ Loads a graph from graph6 format """
     c = s[0]

### 2014-04-08 main.tex
Add bibliography
abstract and correct order of authors

@@
 %      reference to other results?
-\title{A Kochen-Specker Vector System has at least 22 vertices}
+\title{A Kochen-Specker system has at least 22 vertices}
 
+\author{Sander Uijlen}
 \author{Bas Westerbaan}
-\author{Sander Uijlen}
+
+% TODO will we keep these @cs.ru.nl addresses?
 \affil{Institute for Computing and Information Sciences\\
@@
 \begin{abstract}
-    TODO
+    A Kochen-Specker system is a set of points on the sphere,
+    such that there is no~$\{0,1\}$-coloring such that
+    at most one of two orthogonal points is assigned~$1$
+    and of three pairwise orthogonal points exactly one
+    is assigned~$1$.
+    The existence of such a system is the basis of the Kochen-Specker
+    and Free Will theorems.  At the time of writing, the smallest known
+    KS system has 31 vectors.
+
+    Arends, Ouaknine and Wampler have shown that a KS system has at least
+    18 vectors, by reducing the problem to the existence of graphs
+    with a certain topological embeddability and non-colorability property.
+    Deciding embeddability and the sheer number of graphs on more than~$17$
+    vertices, proved the bottleneck in their search.
+
+    Continuing their effort, we restrict our enumeration to smaller class of
+    graphs and develop a practical decision procedure for embeddability, to
+    establish an improved lower bound of 22.
 \end{abstract}
@@
 
+\bibliography{main}{}
+\bibliographystyle{plain}
+
 \end{document}

### 2014-04-08 main.tex
Also for email adresses
Add attributions and amsthm
Improve abstract

@@
 
+\usepackage{paralist}
 \usepackage{authblk}
+\usepackage{amsthm}
+
+\newcounter{main}
+\newtheorem{prop}[main]{Proposition}
+\newtheorem{thm}[main]{Theorem}
+\newtheorem{lem}[main]{Lemma}
+\theoremstyle{definition}
+\newtheorem{dfn}[main]{Definition}
+\theoremstyle{remark}
+\newtheorem{rem}[main]{Remark}
 
@@
        Radboud Universiteit Nijmegen\\
-   \{\texttt{bwesterb},\texttt{suijlen}\}\texttt{@cs.ru.nl}}
+   \{\texttt{suijlen},\texttt{bwesterb}\}\texttt{@cs.ru.nl}}
 
@@
 \begin{abstract}
-    A Kochen-Specker system is a set of points on the sphere,
-    such that there is no~$\{0,1\}$-coloring such that
-    at most one of two orthogonal points is assigned~$1$
+    At the heart of the Conway's Free Will theorems and Kochen-Specker's
+        argument
+    is the existence of a Kochen-Specker (KS) system:
+    a set of points on the sphere,
+    that has no~$\{0,1\}$-coloring such that
+    at most one of two orthogonal points are colored~$1$
     and of three pairwise orthogonal points exactly one
-    is assigned~$1$.
-    The existence of such a system is the basis of the Kochen-Specker
-    and Free Will theorems.  At the time of writing, the smallest known
+    is colored~$1$.
+    In public lectures, Conway encouraged the search for small
+    KS systems.  
+    At the time of writing, the smallest known
     KS system has 31 vectors.
@@
     18 vectors, by reducing the problem to the existence of graphs
-    with a certain topological embeddability and non-colorability property.
+    with a topological embeddability and non-colorability property.
     Deciding embeddability and the sheer number of graphs on more than~$17$
@@
 \section{Introduction}
+
+% TODO whose idea is this orthogonality graph
 TODO
@@
 
+\section{Acknowledgments}
+We wish to thank the following for their generous contribution to the
+distributed computation:
+    the Digital Security group, Intelligent Systems group
+    and the C\&CZ service of the Radboud University;
+    Wouter Geraedts and
+    Jille Timmermans.
+
+% attribute mckay for answering questions?
 \bibliography{main}{}

### 2014-04-08 main.tex
Improve abstract

@@
     graphs and develop a practical decision procedure for embeddability, to
-    establish an improved lower bound of 22.
+    improve the lower bound to 22.
 \end{abstract}

### 2014-04-08 main.tex
start of introduction§

@@
 % TODO whose idea is this orthogonality graph
-TODO
+\subsection{The experiment}
+
+Let's perform the following experiment.  Shoot a deuterium atom,
+or any other spin-1 particle,
+along, say: the x-axis, through a uniform magnetic field.
+
+Depending on the direction of the magnetic field,
+the particle will move undisturbed
+or deviate.
+
+Quantum Mechanics only predicts the chance, given the initial configuration
+of the field, whether the particle will deviate.
+Its probabilistic prediction has been thouroughly tested.
+One could wonder: is there a deterministic theory.
 

### 2014-04-09 main.tex
Some more updates

@@
 
+\usepackage{draftwatermark}
 \usepackage{paralist}
@@
 \usepackage{amsthm}
+
+% TODO remove this in the final version
+\newcommand{\keyword}[1]{\textbf{#1}}
 
@@
 \newtheorem{dfn}[main]{Definition}
+\newtheorem*{spin}{SPIN Axiom \cite{ck09}}
 \theoremstyle{remark}
@@
 \begin{abstract}
-    At the heart of the Conway's Free Will theorems and Kochen-Specker's
-        argument
+    At the heart of the Conway's Free Will theorems and Kochen and Specker's
+        argument against noncontextual hidden variable theories
     is the existence of a Kochen-Specker (KS) system:
@@
 
+% TODO correct description of experiment
 Let's perform the following experiment.  Shoot a deuterium atom,
@@
 along, say: the x-axis, through a uniform magnetic field.
-
 Depending on the direction of the magnetic field,
@@
 Its probabilistic prediction has been thouroughly tested.
-One could wonder: is there a deterministic theory.
+One wonders: is there a deterministic theory predicting the
+outcome of this experiment?
 
-\section{The 33 vector system}
-TODO
+Kochen and Specker proved that such a theory cannot satisfy:
+% TODO note contextuality
+\begin{spin}
+    Given three pairwise orthogonal directions.
+    In exactly one of the directions, the particle will not deviate.
+\end{spin}
+Their argument is based on the existence of a Kochen-Specker system.
+\begin{dfn}
+    A \keyword{Kochen-Specker (KS) system} is
+    a finite set of points on the sphere
+    for which each pair is not antipodal and
+    there is no~\keyword{010-coloring}.
+    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
+    \begin{enumerate}
+        \item
+            no pair of orthogonal points are both colored~$1$ and
+        \item
+            of three pairwise orthogonal points exactly one is colored~$1$;
+            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
+    \end{enumerate}
+\end{dfn}
+% TODO shorten this proof.
+For suppose there is a KS system and  a deterministic theory satisfying
+the SPIN Axiom.
+Then we color a point of the system~$0$,
+whenever the theory predicts that the particle will deviate
+if the magnetic field is directed along that point and~$1$ otherwise.
+Given two orthogonal points of the system,
+we can find a third point orthogonal to both of them.
+The SPIN axiom implies exactly one of them is colored~$1$.
+Thus they cannot both be colored~$1$.
+And similarly, given three pairwise orthogonal vectors in the system,
+the SPIN axiom implies exactly one of them is colored~$1$.
+Hence the KS system is 010-colorable, quod non.  Thus the deterministic theory
+cannot satisfy the SPIN Axiom.
+
+The KS system proposed by Kochen and Specker contained 117 points\cite{ks}.
+% TODO reference (x2)
+Penrose and Peres independently found a smaller system of 33 points.
+The current record is the 31 point system of Conway.
+As pointed out by \cite{c00,aow11}, finding small KS systems
+is of both theoretical and practical interest.
+In public lectures, Conway himself, stressed the search for small KS systems.
+
+% TODO refer to non-3d systems
+In \cite{aow11} Arends, Ouaknine and Wampler show that a KS system must have
+at least 18 vectors.  We have improved upon their result, abd
+
+\subsection{The 33 vector KS system of Penrose and Peres}
 

### 2014-04-09 main.tex
Written down AOW's search

@@
 \usepackage{authblk}
+\usepackage{amsmath}
 \usepackage{amsthm}
@@
 \newtheorem{prop}[main]{Proposition}
+\newtheorem{cathm}[main]{Computer-aided Theorem}
 \newtheorem{thm}[main]{Theorem}
 \newtheorem{lem}[main]{Lemma}
+\newtheorem{fact}[main]{Fact}
 \theoremstyle{definition}
@@
 % TODO refer to non-3d systems
-In \cite{aow11} Arends, Ouaknine and Wampler show that a KS system must have
-at least 18 vectors.  We have improved upon their result, abd
+In \cite{aow11} Arends, Ouaknine and Wampler give a computer aided proof
+that a KS system must have at least 18 vectors.  We improve their lower bound
+and show that a KS must have at least 22 vectors.
 
 \subsection{The 33 vector KS system of Penrose and Peres}
+TODO
+
+\subsection{Kochen-Specker graphs}
+We follow \cite{aow11} and reduce the search for Kochen-Specker systems
+to the search of a certain class of graphs.
+First note that in a Kochen-Specker system we may replace a point with its
+antipodal point.  They are both orthogonal to the same points and hence
+the non-010-colorability is preserved.
+Thus, we may assume antipodal points are identified on the sphere.
+That is: a Kochen-Specker system is a finite subset of the projective plane
+that is not 010-colorable.
+\begin{dfn}
+Given a finite subset~$S$ of the projective plane
+(or equivalently, of the northern hemisphere without equator).
+Define its \keyword{orthogonality graph}~$G(S)$ as follows.
+The vertices are the points of~$S$.
+Two vertices are joined by an edge, if their corresponding points
+are orthogonal.
+\end{dfn}
+\begin{dfn}
+A graph~$G$ is called~\keyword{embeddable},
+if it occurs as a orthogonality graph;
+that is: if there is a finite subset~$S$ of the projective plane,
+such that~$G(S) = G$.
+\end{dfn}
+\begin{dfn}
+A graph is~\keyword{010-colorable}
+if there is a~$\{0,1\}$-coloring, such that
+\begin{enumerate}
+\item
+for each triangle there is exactly one vertex that is colored~$1$ and
+\item
+adjacent vertices are not both colored~$1$.
+\end{enumerate}
+\end{dfn}
+\begin{dfn}
+A \keyword{Kochen-Specker graph}
+is a embeddable graph that is not 010-colorable.
+\end{dfn}
+It is an easy, but important, consequence of the definitions that:
+\begin{fact}
+    A finite subset~$S$ of the projective plane
+    is a Kochen-Specker system,
+    if and only if its orthogonality graph~$G(S)$
+    is Kochen-Specker.
+\end{fact}
+To prove there is no Kochen-Specker system on~$17$ points,
+it is sufficient to enumerate all graphs on~$17$ vertices
+and check they are not 010-colorable or not embeddable.
+This is infeasable, there are~${\sim}10^{26}$ non-isomorphic
+graphs on~$17$ points.
+% TODO OEIS A000088
+However, we can restrict ourselves to certain classes of graphs.
+\begin{prop}[\cite{aow11}]
+    An embeddable graph does not contain a square.
+\end{prop}
+\begin{proof}
+    Given points~$a\neq b$.
+    Consider the points orthogonal to~$a$.
+    This is a great circle.
+    The points orthognal to~$b$ is a different great circle.
+    They intersect in precisely two antipodal points.
+    Hence, if~$c$ and~$d$ are both orthogonal to~$a$ and~$b$,
+    then~$c$ and~$d$ are equivalent.
+    Thus in an embeddable graph, there cannot be a square~$acbd$.
+\end{proof}
+The squarefreeness is a considerable restriction.  There are
+only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$ vertices.
+We can restrict ourselves to connected graphs.
+% TODO OEIS A006786
+\begin{prop}[\cite{aow11}]
+    A minimal Kochen-Specker graph is connected.
+\end{prop}
+\begin{proof}
+    Suppose~$G$ is a non-connected Kochen-Specker graph.
+    Then one of its components is not 010-colorable.
+    As a subgraph of an embeddable graph, is embeddable,
+    this component is embeddable as well.
+    Hence it is a smaller connected Kochen-Specker graph.
+\end{proof}
+The gain, however, is small.
+There are only~${\sim}10^9$ non-isomorphic squarefree graphs on~$17$
+vertices that are not connected.
+We have verified the main result of \cite{aow11}:
+\begin{cathm}
+The unique non-010-colorable squarefree connected graph on~$17$
+vertices is. TODO
+It is not embeddable, see TODO, and hence a Kochen-Specker
+system has at least 18 points.
+\end{cathm}
 
 \section{An improved lower bound}
+Continuing their effort, we consider another restriction.
+\begin{prop}
+    A minimal Kochen-Specker graph has minimal vertex-order three.
+\end{prop}
+\begin{proof}
+    Given a Kochen-Specker graph~$G$.
+    Suppose~$v$ is a vertex with order less than~$2$.
+    Let~$G'$ be~$G$ with~$v$ removed.
+    Clearly~$G'$ is embeddable.
+    Suppose~$G'$ is 010-colorable.
+    Then we can extend the coloring to a coloring of~$G$ as follows.
+    If~$v$ is adjacent to only one or no vertex,
+    then we can color~$v$ with~$0$.
+    Suppose~$v$ is adjacent to two vertices, say~$w$ and~$w'$.
+    If one of~$w$ or~$w'$ is colored~$1$, we can color~$v$ with~$0$.
+    If both~$w$ and~$w'$ are colored~$0$, we can color~$v$ with~$1$.
+    Thus~$G$ is 010-colorable, quod non.  Thus~$G'$ is a smaller
+    Kochen-Specker graph.
+\end{proof}
+There are only~${\sim}10^7$
+squarefree non-isomorphic graphs on 17 vertices with minimal vertex order 3.
+Even though Arends, Ouaknine and Wampler
+note this restriction once,
+surprisingly, they did not restrict their graph enumeration
+to graphs with minimal vertex order 3.
+
+\begin{prop}
+
+\end{prop}
+
+\section{Embeddability}
 TODO
 
-\section{Embeddability}
+\section{Future research}
 TODO

### 2014-04-09 main.tex
More text on connectedness

@@
 \newtheorem{prop}[main]{Proposition}
-\newtheorem{cathm}[main]{Computer-aided Theorem}
+\newtheorem{comp}[main]{Computation}
+\newtheorem{cor}[main]{Corollary}
 \newtheorem{thm}[main]{Theorem}
@@
 % TODO OEIS A006786
-\begin{prop}[\cite{aow11}]
+\begin{prop}[\cite{aow11}]\label{prop:ks-conn}
     A minimal Kochen-Specker graph is connected.
@@
 We have verified the main result of \cite{aow11}:
-\begin{cathm}
+\begin{comp}
 The unique non-010-colorable squarefree connected graph on~$17$
@@
 system has at least 18 points.
-\end{cathm}
+\end{comp}
 
@@
 
+We continue with a strengthening of Proposition~\ref{prop:ks-conn}.
 \begin{prop}
-
+A minimal Kochen-Specker graph is biconnected,
+that is: removing any single edge leaves the graph connected.
 \end{prop}
+\begin{proof}
+    TODO (Sander)
+\end{proof}
+We believe a minimal KS graph is also triconnected.
+Although these arguments are pleasing,
+they are of little use:
+% TODO Add argument or computation for {2,3}-connected
+\begin{comp}
+    There are five non-isomorphic minimal
+    squarefree connected graphs
+    with minimal vertex order 3 and they have 10 vertices.
+\end{comp}
+\begin{cor}
+    Any unconnected
+    squarefree graph with minimal vertex order 3
+    has at least 20 vertices, for it has two connected components,
+    each with at least 10 vertices.
+\end{cor}
 

### 2014-04-10 main.tex
test
Add modified proof of Sander

@@
 A graph~$G$ is called~\keyword{embeddable},
-if it occurs as a orthogonality graph;
+if it occurs as a subgraph of an orthogonality graph;
 that is: if there is a finite subset~$S$ of the projective plane,
-such that~$G(S) = G$.
+such that~$G \leq G(S)$.
 \end{dfn}
+% TODO add remark on subgraph definition?
 \begin{dfn}
@@
 \begin{proof}
-    TODO (Sander)
+Suppose we have a connected but not biconnected graph~$G$.
+Then we can decompose
+this graph into its biconnected components.  A biconnected components
+is a maximal subgraph that is biconnected.
+We can consider the graph of biconnected components~$B(G)$:
+two biconnected components are adjacent if there is an edge
+between a vertex in one component and a vertex in the other.
+Note that there can at most be one edge between 
+the vertices of biconnected components.
+$B(G)$~does not contain loops: if it would
+then the union of the biconnected components in the graph,
+would be itself biconnected.
+
+Now consider a leaf~$A$ in the biconnected component graph~$B(G)$.
+Let~$B$ be the union of the remaining biconnected components.
+There is exactly one pair of vertices~$(a,b)$
+such that~$a \in A$, $b \in B$ and $a$ is adjacent to~$b$.
+
+Take any~$b_2 \in B$ with~$b_2 \neq b$.
+By assumption~$b_2$ is not adjacent to~$a$.
+Define~$G'$ to be the graph~$G$ extended with the edge~$(a,b_2)$.
+Clearly, $G'$ is not~010-colorable.  We will show that~$G'$ is embeddable.
+
+Consider an embedding of~$G$.
+Observe that a rotation of all (points corresponding to) elements of~$A$
+along the axis spanned
+by (the point corresponding to)~$b$, preserves all relevant
+orthogonality relations.
+
+Rotate the vectors of the embedding
+along~$b$ such that~$b_2$ becomes orthogonal to~$a$.
+It could that there are elements~$a'$ in~$A$
+and~$b'$ in~$B$ that are rotated onto the same point.
+Then consider rotations of all elements of~$B$ along~$a$.
+There are finitely many angles for which two elements will overlap.
+Hence, there is a rotation of all elements of~$B$ along~$a$
+such that~$A$ and~$B$ do not overlap.
+Note that all existing orthogonality relations are preserved.
+
+These two rotations yield a new set of points on the sphere,
+that demonstrate~$G'$ is embeddable. 
 \end{proof}

### 2014-04-13 main.tex
Updates

@@
 
+% TODO replace 21 with 22
 % TODO remove this in the final version
@@
 %      reference to other results?
-\title{A Kochen-Specker system has at least 22 vertices}
+\title{A Kochen-Specker system has at least 21 vertices}
 
@@
     graphs and develop a practical decision procedure for embeddability, to
-    improve the lower bound to 22.
+    improve the lower bound to 21.
 \end{abstract}
@@
 that a KS system must have at least 18 vectors.  We improve their lower bound
-and show that a KS must have at least 22 vectors.
+and show that a KS must have at least 21 vectors.
 
@@
 % TODO Add argument or computation for {2,3}-connected
+% TODO note that restricting to (bi/tri)-connected is a waste of time
+% TODO add graph 
 \begin{comp}
@@
 \end{cor}
+% TODO is core x years a proper expression?
+Now, we can state our main computation, which took roughly a week
+    on a 64-core Opteron 6276.
+\begin{comp}
+    Let~$C_n$ denote the number of non-010 colorable squarefree
+    graphs with minimal vertex order 3 on~$n$ nodes.  Then:
+
+    \begin{tabular}{ll}
+        $n$ & $C_n$ \\
+        \hline
+        $\leq 16$ & $0$ \\
+        $17$ & $1$ \\
+        $18$ & $2$ \\
+        $19$ & $19$ \\
+        $20$ & $441$
+    \end{tabular}
+
+    All these 463 graphs are not embeddable.
+    % TODO add reference
+\end{comp}
 
 \section{Embeddability}
-TODO
 

### 2014-04-13 main.tex
updates
Add overview

@@
 % TODO correct description of experiment
-Let's perform the following experiment.  Shoot a deuterium atom,
+Consider the following experiment.  Shoot a deuterium atom,
 or any other spin-1 particle,
-along, say: the x-axis, through a uniform magnetic field.
+along, say: the x-axis, through a inhomogenous magnetic field.
 Depending on the direction of the magnetic field,
@@
 % TODO refer to non-3d systems
-In \cite{aow11} Arends, Ouaknine and Wampler give a computer aided proof
+\subsection{Overview}
+In \cite{aow11} Arends, Ouaknine and Wampler (AOW) give a computer aided proof
 that a KS system must have at least 18 vectors.  We improve their lower bound
-and show that a KS must have at least 21 vectors.
+and show that a KS system must have at least 21 vectors.
+
+First, in Subsection~\ref{sec:ks33} we consider the 33 vector KS system
+of Penrose and Peres.  Then, in Subsection~\ref{sec:ksgraphs},
+we repeat a part of AOW's work, in particuar the reduction of
+KS systems to graphs.
+The bottleneck of their search was the sheer number of graphs
+and the deciding embeddability.
+
+In Section~\ref{sec:ilb},
+we improve upon their reduction,
+to cut down the number of graphs drastically,
+and state the results of our main computation.
+Finally, in Section~\ref{sec:emb},
+we describe our pratical embeddability test.
 
 \subsection{The 33 vector KS system of Penrose and Peres}
+\label{sec:ks33}
 TODO
@@
 \subsection{Kochen-Specker graphs}
+\label{sec:ksgraphs}
 We follow \cite{aow11} and reduce the search for Kochen-Specker systems
@@
 \section{An improved lower bound}
+\label{sec:ilb}
 Continuing their effort, we consider another restriction.
@@
 
-    \begin{tabular}{ll}
-        $n$ & $C_n$ \\
+    \begin{center}
+    \begin{tabular}{l|lllll}
+        $n$ & $\leq 16$
+            & $17$
+            & $18$
+            & $19$
+            & $20$ \\
         \hline
-        $\leq 16$ & $0$ \\
-        $17$ & $1$ \\
-        $18$ & $2$ \\
-        $19$ & $19$ \\
-        $20$ & $441$
+        $C_n$ & $0$
+            & $1$
+            & $2$
+            & $19$
+            & $441$
     \end{tabular}
+    \end{center}
 
@@
 \section{Embeddability}
+\label{sec:emb}
 

### 2014-04-13 main.tex
unemb proof of 10.2

@@
 \usepackage{paralist}
+\usepackage{graphicx}
 \usepackage{authblk}
 \usepackage{amsmath}
+\usepackage{wrapfig}
 \usepackage{amsthm}
@@
 \label{sec:emb}
+\begin{wrapfigure}{r}{0.40\textwidth}
+\begin{center}
+\includegraphics[width=50mm]{graphs/unemb-10-2.jpg}
+\end{center}
+\caption{One of the two minimal
+        non-embeddable graphs
+\label{fig:unemb-10-2}}
+\end{wrapfigure}
+
+Our computation has yielded a few hundred non-010-colorable graphs.
+If we show one of them is embeddable, we have found a new KS system.
+If we demonstrate all of them are not embeddable, we have
+proven a lower bound on the size of a minimal KS system.
+
+In~\cite{aow11}, Arends, Wampler and Ouaknine discuss several methods
+to test embeddability of a graph.  None of the methods could decide
+of all graphs considered, whether they were embeddable or not.
+% TODO give a small description of their deficiencies
+% TODO give intermediate computation.
+
+We propose a new method,
+which of all graphs we considered,
+could decide
+whether they were embeddable or not.
+It is based on the following derivation.
+\begin{prop}
+The graph in Figure~\ref{fig:unemb-10-2}
+is not embeddable.
+\end{prop}
+\begin{proof}
+Suppose it is embeddable.
+Consider~$p_1$.
+It is orthogonal to both~$a$ and~$v$.
+$a$ and~$v$ are not collinear,
+hence~$p_1$ must be collinear to~$v \times a$,
+the cross-product of~$v$ and~$a$.
+Similarly, $p_2$ is collinear to~$v \times p_1 = v \times (v \times a)$.
+Continuing in this fashion,
+we see that
+\begin{equation}\label{eq:ue1}
+    a \text{ is collinear to }
+    x \times (x \times( w \times (w\times (v \times (v \times a))))).
+\end{equation}
+Now, we may assume that~$z=(0,0,1)$ and $x=(1,0,0)$.
+Thus: $v=(v_1,v_2,0)$;
+$w = (w_1,w_2,0)$
+and $a = (0, a_2,a_3)$ for some~$-1 \leq v_1,v_2,w_1,w_2,a_2,a_3 \leq 1$
+with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
+Now, \eqref{eq:ue1} becomes:
+\begin{equation*}
+\begin{pmatrix}
+0\\
+a_2\\
+a_3\\
+\end{pmatrix}
+\text{ is collinear to }
+\begin{pmatrix}
+0\\
+-a_2 v_1w_2 (v_1w_1 + v_2w_2) \\
+-a_3 (v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2)\\
+\end{pmatrix}.
+\end{equation*}
+And thus
+\begin{align*}
+v_1w_2 \left<v,w\right> & =
+v_1w_2 (v_1w_1 + v_2w_2) \\
+& = v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2 \\
+& = (v_1^2 + v_2^2)w_1^2 + (v_1^2 + v_2^2)w_2^2 \\
+& = w_1^2 + w_2^2 = 1.
+\end{align*}
+Since~$v$ and~$w$ are not collinear,
+we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
+Recall~$|v_1|, |w_2| \leq 1$.
+Thus: $|v_1w_2\left<v,w\right>| < 1$.
+Contradiction.
+\end{proof}
 

### 2014-04-13 main.tex
Redlog part

@@
 \usepackage{draftwatermark}
+\usepackage{algorithm2e}
 \usepackage{paralist}
@@
 
-In~\cite{aow11}, Arends, Wampler and Ouaknine discuss several methods
+In~\cite{aow11}, Arends, Wampler and Ouaknine discuss several
+computer-aided methods
 to test embeddability of a graph.  None of the methods could decide
@@
 whether they were embeddable or not.
-It is based on the following derivation.
-\begin{prop}
+First we give a pen-and-paper example.
+\begin{prop}\label{prop:unemb-10-2}
 The graph in Figure~\ref{fig:unemb-10-2}
@@
 
+% TODO Can we prove: if G is embeddable, then we can add any new vertex of
+%      order 2 and G is still embeddable?
+
+In the previous proof, we fixed, without loss of generality, the position
+of a few vertices.  Then we derived cross-product expressions for the
+remaining vertices.  Finally, we find an equation relating some of
+the cross-product expressions and show it is unsatisfiable.
+We can automate this reasoning as follows.
+
+\begin{algorithm}
+    \While{there are unassigned vertices}{
+        pick an unassigned vertex~$v$\;
+        set~$V(v)=v$\;
+        mark~$v$ as free\;
+        \While{there are unassigned vertices adjacent to two
+                different assigned vertices}{
+            pick such a vertex~$w$ adjacent to the assigned ~$w_1$ and~$w_2$\;
+            set~$V(w)=V(w_1) \times V(w_2)$\;
+            mark edges~$(v,w_1)$ and~$(v,w_2)$ as accounted for\;
+        }   
+    }
+    \For{each pair of vertices~$(v_1, v_2)$}{
+        \If{$(v_1,v_2)$ is not an edge} {
+            record requirement:~``$V(v_1)$ is not collinear to $V(v_2)$''\;
+        }
+    }
+    \For{each edge~$(v_1,v_2)$ not accounted for}{
+        record requirement:~``$V(v_1)$ is orthogonal to~$V(v_2)$''\;
+    }
+\end{algorithm}
+At two points in the algorithm, there is a choice which vertex to pick.
+Depending on the vertices chosen, the number of recorded requirements
+and free points may significantly.  By considering all possible choices,
+one can find the one with least free points.
+
+The requirements can be mechanically converted
+to a formal sentence
+in the language of the real numbers.
+This sentence is true if and only if the graph is embeddable.
+Using variations on Tarski's quantifier elimination procedure,
+we can decide the truth of these sentences.
+We have used the redlog\cite{redlog} package of the reduce algebra
+system.  In Appendix~\ref{adx:unemb-10-2} the Reader can find
+the reduce script generated mechanically for the graph
+in Figure~\ref{fig:unemb-10-2}.
+% TODO references to quantifier elimination
+%
+
 \section{Future research}
@@
 
+\appendix
+\section{Example reduce script}\label{adx:unemb-10-2}
+The following is (a part of) the reduce script mechanically generated
+to prove the non-embeddability of the graph of Figure~\ref{fig:unemb-10-2}.
+The algorithm choose a different assignment, than we did in the
+proof of Proposition~\ref{prop:unemb-10-2}.
+As free points it picked, in order,~$z$, $x$, $v$ and~$p_3$.
+The point~$w$ is assigned~$p_3 \times z$.
+\begin{verbatim}
+load_package redlog;
+rlset R;
+
+procedure d(x,y);
+    (first x) * (first y) +
+    (second x) * (second y) +
+    (third x) * (third y);
+
+procedure k(x,y);
+    {(second x)*(third y) - (third x)*(second y),
+     (third x)*(first y) - (first x)*(third y),
+     (first x)*(second y) - (second x)*(first y)};
+
+v0c1 := 1; v0c2 := 0; v0c3 := 0;
+v1c1 := 0; v1c2 := 1; v1c3 := 0;
+
+v0 := {v0c1, v0c2, v0c3}; 
+v1 := {v1c1, v1c2, v1c3}; 
+v2 := {v2c1, v2c2, v2c3}; 
+v3 := {v3c1, v3c2, v3c3}; 
+v2c1 := 0;
+neq0 := k(v0,k(v3,v1)); 
+\end{verbatim}
+\begin{center} \emph{(snip)} \end{center}
+\begin{verbatim}
+neq29 := k(k(k(k(v3,v1),v1),v2),k(k(v3,v0),v3)); 
+phi := 
+       (first neq0 neq 0 or
+        second neq0 neq 0 or
+        third neq0 neq 0) and 
+\end{verbatim}
+\begin{center} \emph{(snip)} \end{center}
+\begin{verbatim}
+       (first neq29 neq 0 or
+        second neq29 neq 0 or
+        third neq29 neq 0) and 
+       d(v2,v0) = 0 and 
+       d(k(k(v3,v0),v3),k(k(k(k(v3,v1),v1),v2),v2)) = 0 and 
+        true;
+rlqe ex(v3c3,
+     ex(v3c2,
+     ex(v3c1,
+     ex(v2c3,
+     ex(v2c2,phi)))));
+\end{verbatim}
+
 % attribute mckay for answering questions?

### 2014-04-13 main.tex
Sander's work

@@
 
-    Continuing their effort, we restrict our enumeration to smaller class of
-    graphs and develop a practical decision procedure for embeddability, to
-    improve the lower bound to 21.
+    Continuing their effort, we prove a restriction on the class of graphs
+    we need to consider and develop a more practical decision procedure for
+    embeddability to improve the lower bound to 21.
 \end{abstract}
@@
 or any other spin-1 particle,
-along, say: the x-axis, through a inhomogenous magnetic field.
+along, say: the x-axis, through a inhomogeneous magnetic field.
 Depending on the direction of the magnetic field,
@@
 of the field, whether the particle will deviate.
-Its probabilistic prediction has been thouroughly tested.
+Its probabilistic prediction has been thoroughly tested.
 One wonders: is there a deterministic theory predicting the
@@
 % TODO shorten this proof.
-For suppose there is a KS system and  a deterministic theory satisfying
+A point on the sphere obviously corresponds to a direction in space.
+Because of this, the term point, vector and direction
+can be used interchangeably. Antipodal points correspond to oppositve
+vectors and these span the same direction in space.
+
+Suppose there is a KS system and  a deterministic theory satisfying
 the SPIN Axiom.
@@
 of Penrose and Peres.  Then, in Subsection~\ref{sec:ksgraphs},
-we repeat a part of AOW's work, in particuar the reduction of
+we repeat a part of AOW's work, in particular the reduction of
 KS systems to graphs.
@@
 Finally, in Section~\ref{sec:emb},
-we describe our pratical embeddability test.
+we describe our practical embeddability test.
 
@@
 the non-010-colorability is preserved.
-Thus, we may assume antipodal points are identified on the sphere.
+Therefore, we may assume antipodal points are identified on the sphere.
 That is: a Kochen-Specker system is a finite subset of the projective plane
@@
 To prove there is no Kochen-Specker system on~$17$ points,
-it is sufficient to enumerate all graphs on~$17$ vertices
-and check they are not 010-colorable or not embeddable.
-This is infeasable, there are~${\sim}10^{26}$ non-isomorphic
+it would be sufficient to enumerate all graphs on~$17$ vertices
+and check these are not 010-colorable or not embeddable.
+However, this is infeasable as there are
+already~${\sim}10^{26}$ non-isomorphic
 graphs on~$17$ points.
 % TODO OEIS A000088
-However, we can restrict ourselves to certain classes of graphs.
+Luckily, we can restrict ourselves to certain classes of graphs.
 \begin{prop}[\cite{aow11}]
@@
 \begin{proof}
-    Given points~$a\neq b$.
+    Given two non antipodal points~$a\neq b$.
     Consider the points orthogonal to~$a$.
@@
     then~$c$ and~$d$ are equivalent.
-    Thus in an embeddable graph, there cannot be a square~$acbd$.
+    Therefore, an embeddable graph cannot contain a square.
 \end{proof}
@@
 \label{sec:ilb}
-Continuing their effort, we consider another restriction.
+Continuing the effort of Arends, Ouaknine and Wambler,
+we consider another restriction.
 \begin{prop}
@@
     Given a Kochen-Specker graph~$G$.
-    Suppose~$v$ is a vertex with order less than~$2$.
+    Suppose~$v$ is a vertex with order less than or equal~$2$.
     Let~$G'$ be~$G$ with~$v$ removed.
@@
     If both~$w$ and~$w'$ are colored~$0$, we can color~$v$ with~$1$.
-    Thus~$G$ is 010-colorable, quod non.  Thus~$G'$ is a smaller
-    Kochen-Specker graph.
+    This would imply~$G$ is 010-colorable, quod non.
+    Therefore~$G'$ is a smaller
+    Kochen-Specker graph, which contradicts minimality.
 \end{proof}

### 2014-04-14 main.tex
Some more work.

@@
 
-% TODO this suggests only one result: the improved lower bound
-%      reference to other results?
 \title{A Kochen-Specker system has at least 21 vertices}
@@
     A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
+        \footnote{
+                In other papers, like \cite{aow11},
+                the~$0$ and~$1$ are swapped:
+                101-colorings.
+                Then the colors correspond directly to the squared measurement
+                outcome of the spin.}
     \begin{enumerate}
@@
 The KS system proposed by Kochen and Specker contained 117 points\cite{ks}.
-% TODO reference (x2)
-Penrose and Peres independently found a smaller system of 33 points.
-The current record is the 31 point system of Conway.
+Penrose and Peres\cite{peres} independently found a smaller system of 33 points.
+The current record is the 31 point system of Conway\cite[p.~197]{qtcm}.
 As pointed out by \cite{c00,aow11}, finding small KS systems
@@
 
-First, in Subsection~\ref{sec:ks33} we consider the 33 vector KS system
-of Penrose and Peres.  Then, in Subsection~\ref{sec:ksgraphs},
+First, in Subsection~\ref{sec:ksgraphs},
 we repeat a part of AOW's work, in particular the reduction of
@@
 we improve upon their reduction,
-to cut down the number of graphs drastically,
+to cut down the number of graphs to consider drastically,
 and state the results of our main computation.
@@
 we describe our practical embeddability test.
-
-\subsection{The 33 vector KS system of Penrose and Peres}
-\label{sec:ks33}
-TODO
 
@@
 This sentence is true if and only if the graph is embeddable.
-Using variations on Tarski's quantifier elimination procedure,
-we can decide the truth of these sentences.
+Famously, Tarski proved\cite{tarski}
+that such sentences are decidable.
+His decision procedure has an impractical complexity.
+However, its practical value has been improved
+by, for instance, the method of cylindrical algebraic decomposition\cite{qecad}.
 We have used the redlog\cite{redlog} package of the reduce algebra
-system.  In Appendix~\ref{adx:unemb-10-2} the Reader can find
+system, which implements a variant of Tarski's quantifier elimination.
+In Appendix~\ref{adx:unemb-10-2} the Reader can find
 the reduce script generated mechanically for the graph
 in Figure~\ref{fig:unemb-10-2}.
-% TODO references to quantifier elimination
-%
 
-\section{Future research}
+\section{Conclusion and future research}
 TODO
@@
 % attribute mckay for answering questions?
+% TODO attribute geng
 \bibliography{main}{}

### 2014-04-14 main.tex
changes

@@
 
-\title{A Kochen-Specker system has at least 21 vertices}
+\title{A Kochen-Specker system has at least 21 vertices\\
+        {\small (extended abstract)}}
 
@@
 % TODO will we keep these @cs.ru.nl addresses?
-\affil{Institute for Computing and Information Sciences\\
-       Radboud Universiteit Nijmegen\\
+\affil{{\small Institute for Computing and Information Sciences\\
+        Radboud Universiteit Nijmegen}\\
    \{\texttt{suijlen},\texttt{bwesterb}\}\texttt{@cs.ru.nl}}
@@
 
+We are grateful to prof.~McKay for discussing
+the feasability of certain graph restrictions.
+
 \appendix

### 2014-04-14 main.tex
work

@@
 
+Different assignments give different sentences.  In our tests,
+some assignments would yield sentences that were decided within milliseconds,
+whereas another assignment with less free vertices would
+yield a sentence that could not be decided (directly).
+Therefore, when determining embeddability of a graph,
+we try several assignments in parallel.
+
+In this way, there were still a few graphs that we could not decide.
+By hand, we determined these graphs to be embeddable.
+We addapted the algorithm, to in parallel for some assignments
+to guess the position of one of the vectors.  If the corresponding
+sentence turns out false, we know nothing.  However,
+if the sentence is true, we know the graph is embeddable.
+
+With this method, we have decided the embeddability of every squarefree
+graph with minimal vertex order three of 13 vertices of less.
+In particular:
+\begin{comp}
+    Every squarefree graph of minimal vertex order three
+    that is not 010-colorable
+    of order less than or equal 20
+    contains as a subgraph one of
+    \begin{center}
+        \includegraphics[width=120mm]{graphs/unemb-base-20.jpg}
+    \end{center}
+    which are unembeddable.  The left and middle graph
+    are the only minimal unembeddable squarefree graph.
+\end{comp}
+For the first graph, we have proven directly that it is unembeddable.
+See Proposition~\ref{prop:unemb-10-2}.
+For the second graph, we also have a direct proof.
+
 \section{Conclusion and future research}

### 2014-04-14 main.tex
bla

--- 2014-04-14 main.tex
@@
 
-Quantum Mechanics only predicts the chance, given the initial configuration
+Quantum Mechanics only predicts the probability, given the configuration
 of the field, whether the particle will deviate.
 Its probabilistic prediction has been thoroughly tested.
-One wonders: is there a deterministic theory predicting the
+One wonders: is there a deterministic noncontextual theory predicting the
 outcome of this experiment?
@@
                 In other papers, like \cite{aow11},
-                the~$0$ and~$1$ are swapped:
-                101-colorings.
-                Then the colors correspond directly to the squared measurement
-                outcome of the spin.}
+                the~$0$ and~$1$ are swapped, they consider 101-colorings.
+                These colorings are of course equivalent and the difference
+                arises from considering either squared spin measurements $S^2$, or $1-S^2$.
+                %Then the colors correspond directly to the squared measurement
+               	%outcome of the spin.
+               	}
     \begin{enumerate}
@@
 Because of this, the term point, vector and direction
-can be used interchangeably. Antipodal points correspond to oppositve
+can be used interchangeably. Antipodal points correspond to opposite
 vectors and these span the same direction in space.
 
-Suppose there is a KS system and  a deterministic theory satisfying
+Suppose there is a KS system and a noncontextual deterministic theory satisfying
 the SPIN Axiom.
-Then we color a point of the system~$0$,
-whenever the theory predicts that the particle will deviate
-if the magnetic field is directed along that point and~$1$ otherwise.
+Then we color a point of this system~$0$,
+whenever this theory predicts that the particle will deviate
+if the spin is measured in the direction corresponding to that point, and~$1$ otherwise.
 Given two orthogonal points of the system,
 we can find a third point orthogonal to both of them.
-The SPIN axiom implies exactly one of them is colored~$1$.
-Thus they cannot both be colored~$1$.
-And similarly, given three pairwise orthogonal vectors in the system,
+The SPIN axiom implies exactly one of them is colored~$1$, so they cannot both be colored~$1$.
+Similarly, given three pairwise orthogonal vectors in the system,
 the SPIN axiom implies exactly one of them is colored~$1$.
-Hence the KS system is 010-colorable, quod non.  Thus the deterministic theory
-cannot satisfy the SPIN Axiom.
+Hence there would be a 010-coloring of the KS system, quod non. Therefore a deterministic noncontextual theory cannot satisfy the SPIN Axiom.\\
 
@@
 The bottleneck of their search was the sheer number of graphs
-and the deciding embeddability.
+and the deciding whether such graphs are embeddable.
 
@@
 and check these are not 010-colorable or not embeddable.
-However, this is infeasable as there are
+However, this is infeasible as there are
 already~${\sim}10^{26}$ non-isomorphic
@@
 \begin{comp}
-The unique non-010-colorable squarefree connected graph on~$17$
-vertices is. TODO
+There is a unique non-010-colorable squarefree connected graph on~$17$ or less
+vertices. It looks as follows TODO
 It is not embeddable, see TODO, and hence a Kochen-Specker
@@
 
-Now consider a leaf~$A$ in the biconnected component graph~$B(G)$.
+Now consider a leaf~$A$ in the biconnected component graph~$B(G)$. That is, $A$ is a component which is only connected to one other component.
 Let~$B$ be the union of the remaining biconnected components.
@@
 that demonstrate~$G'$ is embeddable.
+Therefore, in generating all relevant graphs, we find both $G$ and $G'$.
+Since they have the same order, we might as well ignore the graph $G$.
+This shows it is sufficient to consider only biconnected graphs.
 \end{proof}
@@
 \end{cor}
+
 % TODO is core x years a proper expression?
@@
 computer-aided methods
-to test embeddability of a graph.  None of the methods could decide
-of all graphs considered, whether they were embeddable or not.
+to test embeddability of a graph.  None of these methods could decide
+for all graphs considered, whether they were embeddable or not.
 % TODO give a small description of their deficiencies
@@
 We propose a new method,
-which of all graphs we considered,
+which for all graphs we considered,
 could decide
@@
 $w = (w_1,w_2,0)$
-and $a = (0, a_2,a_3)$ for some~$-1 \leq v_1,v_2,w_1,w_2,a_2,a_3 \leq 1$
+and $a = (0, a_2,a_3)$ for some~$-1 < v_1,v_2,w_1,w_2,a_2,a_3 < 1$,
 with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
@@
 \end{equation*}
-And thus
+And therefore
 \begin{align*}
@@
 Depending on the vertices chosen, the number of recorded requirements
-and free points may significantly.  By considering all possible choices,
+and free points may significantly vary. By considering all possible choices,
 one can find the one with least free points.

### 2014-04-14 main.tex
bla

@@
                 In other papers, like \cite{aow11},
-                the~$0$ and~$1$ are swapped, they consider 101-colorings.
-                These colorings are of course equivalent and the difference
-                arises from considering either squared spin measurements $S^2$, or $1-S^2$.
+                the~$0$ and~$1$ are swapped; they consider 101-colorings.
+                These colorings are of course equivalent and the
+                difference arises from considering either squared
+                spin measurements $S^2$, or $1-S^2$.
                 %Then the colors correspond directly to the squared measurement
@@
 whenever this theory predicts that the particle will deviate
-if the spin is measured in the direction corresponding to that point, and~$1$ otherwise.
+if the spin is measured in the direction corresponding to that
+point, and~$1$ otherwise.
 Given two orthogonal points of the system,
 we can find a third point orthogonal to both of them.
-The SPIN axiom implies exactly one of them is colored~$1$, so they cannot both be colored~$1$.
+The SPIN axiom implies exactly one of them is colored~$1$, so they
+cannot both be colored~$1$.
 Similarly, given three pairwise orthogonal vectors in the system,
 the SPIN axiom implies exactly one of them is colored~$1$.
-Hence there would be a 010-coloring of the KS system, quod non. Therefore a deterministic noncontextual theory cannot satisfy the SPIN Axiom.\\
+Hence there would be a 010-coloring of the KS system, quod non.
+Therefore a deterministic noncontextual theory cannot satisfy the
+SPIN Axiom.
 
@@
 
-Now consider a leaf~$A$ in the biconnected component graph~$B(G)$. That is, $A$ is a component which is only connected to one other component.
+Now consider a leaf~$A$ in the biconnected component graph~$B(G)$.
+That is, $A$ is a biconnected component which is
+only connected to one other biconnected component.
 Let~$B$ be the union of the remaining biconnected components.
@@
 that demonstrate~$G'$ is embeddable.
-Therefore, in generating all relevant graphs, we find both $G$ and $G'$.
-Since they have the same order, we might as well ignore the graph $G$.
-This shows it is sufficient to consider only biconnected graphs.
+Therefore, in generating all relevant graphs, we find both $G$ and
+$G'$. Since they have the same order, we might as well ignore the
+graph $G$. This shows it is sufficient to consider only biconnected
+graphs.
 \end{proof}
@@
 
-In this way, there were still a few graphs that we could not decide.
+In this way, there were still a few (010-colorable) graphs of which
+we could not decide embeddability.
 By hand, we determined these graphs to be embeddable.

### 2014-04-14 main.tex
future research

@@
     that is not 010-colorable
-    of order less than or equal 20
-    contains as a subgraph one of
+    of order less than or equal to 20
+    contains, as a subgraph, one of the following three graphs:
     \begin{center}
@@
     \end{center}
-    which are unembeddable.  The left and middle graph
+    These three graphs are unembeddable.  The left and middle graph
     are the only minimal unembeddable squarefree graph.
@@
 See Proposition~\ref{prop:unemb-10-2}.
-For the second graph, we also have a direct proof.
+For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.
 
 \section{Conclusion and future research}
-TODO
+At the time of writing, we are running a computation to find all possible KS graphs of order 22.
+We can then check if these graphs contain any of our found non embeddable subgraphs.
+We can use our method to check the embeddability of the graphs that have not been excluded.
+In theory, we can do this for all graphs up to order 31, or until we find an embeddable smaller graph.
+However, it might be that the amount of these graphs becomes too big.
+Even to an extend where they exceed even the number of particles in the universe.
 

### 2014-04-14 main.tex
conclusie

--- 2014-04-14 main.tex
@@
 
-With this method, we have decided the embeddability of every squarefree
-graph with minimal vertex order three of 13 vertices of less.
+With this method, we have decided in an hour the embeddability
+of every squarefree graph with minimal vertex order three of 13
+vertices of less.
 In particular:
@@
 \section{Conclusion and future research}
-TODO
+Arends, Ouaknine and Wampler struggled with two problems:
+enumerating candidate graphs of less than 31 vertices
+and testing their embeddability.
+We have verified most of their computation claims.
+Then, we enumerated all candidate graphs
+up to 21 vertices.
+Furthermore, we have proposed a new decision procedure
+which was able to decide embeddability,
+for all candidate graphs we found.
+Therefore, we demonstrate: a Kochen-Specker system must have at least
+21 points.
+
+Enumerating all candidate graphs of less than 31 vertices
+is computationally infeasable.
+To bridge the enormous the gap between 21 and 31,
+requires a new insight.
+For instance: another restriction on which graphs to consider.
+
+The Reader, interested in pursuing this line of research,
+is encouraged to read the master thesis\cite{a09} of Arends,
+in which he discusses in detail several other
+properties that a minimal KS system must enjoy, as well as
+some failed attempts.
 

### 2014-04-14 main.tex

@@
 See Proposition~\ref{prop:unemb-10-2}.
-For the second graph, we also have a direct proof.
+For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.
 

### 2014-04-14 main.tex
bla

@@
 Kochen and Specker proved that such a theory cannot satisfy:
-% TODO note contextuality
 \begin{spin}
@@
     All these 463 graphs are not embeddable.
-    % TODO add reference
+    See Computation~\ref{comp:unemb20}.
 \end{comp}
@@
 In particular:
-\begin{comp}
+\begin{comp}\label{comp:unemb20}
     Every squarefree graph of minimal vertex order three
@@
 and testing their embeddability.
-We have verified most of their computation claims.
-Then, we enumerated all candidate graphs
-up to 21 vertices.
-Furthermore, we have proposed a new decision procedure
-which was able to decide embeddability,
+We have verified most of their computations.
+Then we enumerated all candidate graphs
+up to and including 20 vertices.
+Furthermore, we have proposed a new decision procedure,
+which was able to decide embeddability
 for all candidate graphs we found.
@@
 21 points.
+At the time of writing,
+we are computing
+whether there is a KS system of 21 vertices.\footnote{
+The authors have a wager whether there is a minimal KS system of less
+than 25 vertices.}
 

### 2014-04-14 main.tex
updates

@@
-\documentclass[a4paper]{article}
+\documentclass[a4paper,titlepage]{article}
 
@@
 \end{proof}
+\begin{wrapfigure}{r}{0.40\textwidth}
+\begin{center}
+\includegraphics[width=50mm]{graphs/c17.jpg}
+\end{center}
+\caption{Unique non 010-colorable squarefree connected
+            graph on 17 vertices.}
+\label{fig:c17}
+\end{wrapfigure}
 The squarefreeness is a considerable restriction.  There are
@@
 There is a unique non-010-colorable squarefree connected graph on~$17$ or less
-vertices. It looks as follows TODO
-It is not embeddable, see TODO, and hence a Kochen-Specker
+vertices. Its depicted in Figure~\ref{fig:c17}
+It is not embeddable, as the graph in Figure~\ref{fig:unemb-10-2}
+is an unembeddable subgraph.  See Proposition~\ref{prop:unemb-10-2}.
+Hence a Kochen-Specker
 system has at least 18 points.

### 2014-04-14 main.tex
changes

--- 2014-04-14 main.tex
@@
 
-\usepackage{draftwatermark}
 \usepackage{algorithm2e}
+\usepackage{hyperref}
 \usepackage{paralist}
@@
     is the existence of a Kochen-Specker (KS) system:
-    a set of points on the sphere,
+    a set of points on the sphere
     that has no~$\{0,1\}$-coloring such that
@@
     with a topological embeddability and non-colorability property.
-    Deciding embeddability and the sheer number of graphs on more than~$17$
-    vertices, proved the bottleneck in their search.
+    The bottleneck in their search
+    proved to be the sheer number of graphs on more than~$17$
+    vertices and deciding embeddability.
 
@@
 \end{abstract}
+
+\clearpage
 
@@
     A \keyword{Kochen-Specker (KS) system} is
-    a finite set of points on the sphere
+    a finite set of points on the sphere\footnote{
+            We define KS systems to be three dimensional,
+            as in the original proof of Kochen and Specker.
+            Later, higher dimensional systems have been studied.
+            See, for instance~\cite[p.~201]{qtcm}.
+        }
     for which each pair is not antipodal and
@@
 vertices that are not connected.
+In our computations, checking for connectedness
+required more time than would be gained by reducing the number of graphs.
+
 We have verified the main result of \cite{aow11}:
@@
 There is a unique non-010-colorable squarefree connected graph on~$17$ or less
-vertices. It looks as follows TODO
-It is not embeddable, see TODO, and hence a Kochen-Specker
+vertices:
+\begin{center}
+\includegraphics[width=50mm]{graphs/c17.jpg}
+\end{center}
+It is not embeddable, as the graph in Figure~\ref{fig:unemb-10-2}
+is an unembeddable subgraph.  For our proof,
+see Proposition~\ref{prop:unemb-10-2}.
+Hence a Kochen-Specker
 system has at least 18 points.
@@
 We believe a minimal KS graph is also triconnected.
-Although these arguments are pleasing,
-they are of little use:
+However, we did not find a proof.
+
+Although these restrictions are theoretically pleasing,
+they seem to be of little use as a practical restriction.
+Concerning excluding unconnected graphs:
 % TODO Add argument or computation for {2,3}-connected
@@
     each with at least 10 vertices.
+    With 20 vertices, there are exactly 25 of these.
 \end{cor}
+This justifies, at this stage, not checking for connectedness.
+Similarly, we believe there are very few connected but not biconnected graphs.
 
 % TODO is core x years a proper expression?
-Now, we can state our main computation, which took roughly a week
-    on a 64-core Opteron 6276.
+Now we can state our main computation.
 \begin{comp}
@@
 \end{comp}
+The computation took roughly a week on a 64-core Opteron 6276.
+It was executed as follows.
+We enumerated all squarefree graphs with minimal vertex
+order 3 on less than or equal~$20$ vertices,
+using the~\texttt{geng} util of the nauty software package,
+which uses the isomorphism-free exhaustive generation
+method of McKay\cite{geng}.
+The output of~\texttt{geng}, we passed through
+a custom heuristic backtracker written in~C++
+to decide 010-colorability of these graphs.
 
@@
 $w = (w_1,w_2,0)$
-and $a = (0, a_2,a_3)$ for some~$-1 < v_1,v_2,w_1,w_2,a_2,a_3 < 1$,
+and $a = (0, a_2,a_3)$ for some~$-1 \leq v_1,v_2,w_1,w_2,a_2,a_3 \leq 1$,
 with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
@@
         pick an unassigned vertex~$v$\;
-        set~$V(v)=v$\;
+        assign~$V(v)=v$\;
         mark~$v$ as free\;
@@
             pick such a vertex~$w$ adjacent to the assigned ~$w_1$ and~$w_2$\;
-            set~$V(w)=V(w_1) \times V(w_2)$\;
+            assign~$V(w)=V(w_1) \times V(w_2)$\;
             mark edges~$(v,w_1)$ and~$(v,w_2)$ as accounted for\;
@@
 By hand, we determined these graphs to be embeddable.
-We addapted the algorithm, to in parallel for some assignments
-to guess the position of one of the vectors.  If the corresponding
+We addapted the algorithm, as to guess
+for some assignments the position of one of the vectors.
+If the corresponding
 sentence turns out false, we know nothing.  However,
@@
 
+\clearpage
 \appendix
@@
 % TODO attribute geng
+\clearpage
 \bibliography{main}{}

### 2014-05-06 main.tex
minor comments of anonymous reviewer #1

@@
 \begin{prop}[\cite{aow11}]
-    An embeddable graph does not contain a square.
+    An embeddable graph is squarefree.
+    That is: it does not contain the square as a subgraph.
+    \footnote{Some authors call a graph squarefree, if it does not
+        contain the square as induced subgraph.
+        For them the full graph on four nodes is squarefree.
+        We follow~Weisstein\cite{sf-weisstein} and~Sloane\cite{sf-sloane} and
+        call a graph squarefree if it does not
+        contain the square as subgraph.
+        For us the full graph on four nodes is not squarefree.}
 \end{prop}
@@
     This is a great circle.
-    The points orthognal to~$b$ is a different great circle.
+    The points orthogonal to~$b$ is a different great circle.
     They intersect in precisely two antipodal points.
@@
 then the union of the biconnected components in the graph,
-would be itself biconnected.
+would be itself biconnected.  Thus~$B(G)$ is a tree.
 
@@
 along~$b$ such that~$b_2$ becomes orthogonal to~$a$.
-It could that there are elements~$a'$ in~$A$
+It could be that there are elements~$a'$ in~$A$
 and~$b'$ in~$B$ that are rotated onto the same point.
@@
 By hand, we determined these graphs to be embeddable.
-We addapted the algorithm, as to guess
+We adapted the algorithm, as to guess
 for some assignments the position of one of the vectors.

### 2014-05-06 main.tex
minor update

@@
     That is: it does not contain the square as a subgraph.
-    \footnote{Some authors call a graph squarefree, if it does not
+    \footnote{Some authors call a graph squarefree if it does not
         contain the square as induced subgraph.
-        For them the full graph on four nodes is squarefree.
+        For them the complete graph on four vertices is squarefree.
         We follow~Weisstein\cite{sf-weisstein} and~Sloane\cite{sf-sloane} and
@@
         contain the square as subgraph.
-        For us the full graph on four nodes is not squarefree.}
+        For us the complete graph on four vertices is not squarefree.}
 \end{prop}

### 2014-05-09 code/plot.py
code: remove plot

@@ -1,8 +1,0 @@
-from helper import load_graph6
-
-def plot(g):
-    print g
-
-if __name__ == '__main__':
-    with open('../graphs/ks33.g6') as f:
-        plot(load_graph6(f.readline()[:-1]))

### 2014-05-11 main.tex
Reference to A000088

@@
 already~${\sim}10^{26}$ non-isomorphic
-graphs on~$17$ points.
-% TODO OEIS A000088
+graphs on~$17$ points.\cite{oeisA000088}
 Luckily, we can restrict ourselves to certain classes of graphs.

### 2014-05-11 main.tex
Fix biconnected and add a reference for the public lectures

@@
 is of both theoretical and practical interest.
-In public lectures, Conway himself, stressed the search for small KS systems.
+In public lectures, Conway himself, stressed the search for small KS
+systems.\cite{OC}
 
@@
 We continue with a strengthening of Proposition~\ref{prop:ks-conn}.
-\begin{prop}
+\begin{prop}\label{prop:ks-biconn}
 A minimal Kochen-Specker graph is biconnected,
@@
 \end{prop}
+We need some preparation, before we can prove this Proposition.
+\begin{dfn}
+Given a graph~$G$ and a vertex~$v$ of~$G$.
+We say, \keyword{$v$ has fixed color~$c$ (in~$G$)},
+if~$G$ is~$010$-colorable
+and for every~$010$-coloring of~$G$,
+the vertex~$v$ is assigned color~$c$.
+\end{dfn}
+We are interested in these graphs because of the following observation.
+\begin{lem}
+If there is an embeddable graph~$G$ on~$n$ vertices with a vertex
+with fixed color~$1$,
+then there is a Kochen-Specker graph on~$2n$ vertices.
+\end{lem}
 \begin{proof}
-Suppose we have a connected but not biconnected graph~$G$.
+Let~$G$ be a graph and~$v$ a vertex of~$G$ with fixed color~$1$.
+Consider two copies of the graph~$G$.
+Connect the two instances of~$v$ with an edge.
+Call this graph~$G'$.
+Clearly, $G'$ is not~$010$-colorable.
+
+We need to show~$G'$ is embeddable.
+Given an embedding~$S$ of~$G$.
+We may assume that the point in~$S$ corresponding to~$v$
+is the north pole.
+Furthermore, we may assume that there is no point on the $x$-axis,
+by rotating points along the north pole.
+Let~$S'$ be~$S$ rotated~$90$ degrees along the~$y$-axis.
+Some points of~$S$ and~$S'$ might overlap.
+That is: there might be a point~$s$ in~$S$ and~$s'$ in~$S'$
+that are equal or antipodal.
+Observe that if no points of~$S'$ and~$S$ overlap,
+then~$S \cup S'$ is an embedding of~$G'$.
+
+Suppose there are points in~$S'$ and~$S$
+that overlap.
+Note that the north pole (and south pole) is not in~$S'$. 
+Let~$S''$ be~$S'$ rotated along the north pole at some angle~$\alpha$.
+There are finitely many angles such that there are overlaping points.
+Thus there is an angle such that~$S \cup S''$ is an embedding of~$G'$.
+\end{proof}
+Unfortunately, these graphs are not small.
+\begin{comp}\label{comp:bic1}
+Let~$F^{3,1}_n$ denote the number of connected graphs
+with minimal vertex-order~$3$
+on~$n$ vertices
+with a vertex with fixed color~$1$.
+Then:
+\begin{center}
+\begin{tabular}{l|lll}
+    $n$ & $\leq 13$
+        & $14$
+        & $15$ \\
+    \hline
+    $F^{3,1}_n$ & $0$
+        & $4$
+        & $59$
+\end{tabular}
+\end{center}
+Using the methods of Section~\ref{sec:emb},
+we have determined these graphs unembeddable.
+\end{comp}
+We need one final computational result.
+\begin{comp}\label{comp:bic2}
+There are no connected graphs
+on less than~$15$ vertices
+with minimal vertex-order~$2$
+with a vertex with fixed color~$0$. 
+\end{comp}
+We are ready to prove that a minimal Kochen-Specker graph
+is biconnected.
+\begin{proof}[Proof of Proposition~\ref{prop:ks-biconn}]
+Suppose we are given a connected but not biconnected graph~$G$.
 Then we can decompose
@@
 
+Furthermore suppose~$G$ is a minimal Kochen-Specker graph.
 Now consider a leaf~$A$ in the biconnected component graph~$B(G)$.
-That is, $A$ is a biconnected component which is
-only connected to one other biconnected component.
-Let~$B$ be the union of the remaining biconnected components.
-There is exactly one pair of vertices~$(a,b)$
-such that~$a \in A$, $b \in B$ and $a$ is adjacent to~$b$.
+That is, $A$ is a biconnected component that is connected to only
+one other biconnected component.
+Let~$B$ be the remainder of the graph.
+There is exactly one pair~$(a,b)$ with~$a$ in~$A$ and~$b$ in~$B$
+and~$a$ adjacent to~$b$.
 
-Take any~$b_2 \in B$ with~$b_2 \neq b$.
-By assumption~$b_2$ is not adjacent to~$a$.
-Define~$G'$ to be the graph~$G$ extended with the edge~$(a,b_2)$.
-Clearly, $G'$ is not~010-colorable.  We will show that~$G'$ is embeddable.
+Note that~$A$ and~$B$ are~$010$-colorable.
+Suppose that~$a$ does not have fixed color~$1$ in~$A$.
+Then there is a~$010$-coloring of~$A$
+which assigns~$0$ to~$a$.
+But then~$G$ is~$010$-colorable, quod non.
+Thus~$A$ is an embeddable graph with a vertex with fixed color~$1$.
+Similarly~$B$ is an embeddable graph with a vertex with fixed color~$1$.
+We may assume~$G$ is two copies of~$A$ connected at the
+copies of the vertex with fixed color~$1$.
 
-Consider an embedding of~$G$.
-Observe that a rotation of all (points corresponding to) elements of~$A$
-along the axis spanned
-by (the point corresponding to)~$b$, preserves all relevant
-orthogonality relations.
+The graph~$G$ has minimal vertex-order~$3$.
+Thus every vertex of~$A$ has minimal order~$3$ except for~$a$,
+which has minimal order~$2$.
+Suppose~$a$ has minimal order~$3$ in~$A$.
+Then by Computation~\ref{comp:bic1},
+the graph~$A$ has at least~$16$ vertices.
+Thus~$G$ has at least~$32$ vertices, quod non.
 
-Rotate the vectors of the embedding
-along~$b$ such that~$b_2$ becomes orthogonal to~$a$.
-It could be that there are elements~$a'$ in~$A$
-and~$b'$ in~$B$ that are rotated onto the same point.
-Then consider rotations of all elements of~$B$ along~$a$.
-There are finitely many angles for which two elements will overlap.
-Hence, there is a rotation of all elements of~$B$ along~$a$
-such that~$A$ and~$B$ do not overlap.
-Note that all existing orthogonality relations are preserved.
-
-These two rotations yield a new set of points on the sphere,
-that demonstrate~$G'$ is embeddable.
-Therefore, in generating all relevant graphs, we find both $G$ and
-$G'$. Since they have the same order, we might as well ignore the
-graph $G$. This shows it is sufficient to consider only biconnected
-graphs.
+Thus~$a$ must have order~$2$ in~$A$.
+Let~$c,d \in A$ be the vertices connected to~$a$ in~$A$.
+For~$a$ to have fixed color~$1$ in~$A$,
+the vertices~$c$ and~$d$ must have fixed color~$0$ in~$A$.
+By Computation~\ref{comp:bic2},
+the graph~$A$ has at least~$15+1$ vertices.
+Thus~$G$ has at least~$32$ vertices. Contradiction.
 \end{proof}

### 2014-05-12 main.tex
Another citation, no pagenumber of the first page
Remove stray line
spellchecker

@@
 
+\usepackage{draftwatermark}
+
 % TODO replace 21 with 22
-% TODO remove this in the final version
 \newcommand{\keyword}[1]{\textbf{#1}}
@@
 \maketitle
+\thispagestyle{empty}
 
@@
     At the heart of the Conway's Free Will theorems and Kochen and Specker's
-        argument against noncontextual hidden variable theories
+        argument against non-contextual hidden variable theories
     is the existence of a Kochen-Specker (KS) system:
@@
 Its probabilistic prediction has been thoroughly tested.
-One wonders: is there a deterministic noncontextual theory predicting the
+One wonders: is there a deterministic non-contextual theory predicting the
 outcome of this experiment?
@@
 
-Suppose there is a KS system and a noncontextual deterministic theory satisfying
+Suppose there is a KS system and a non-contextual deterministic theory satisfying
 the SPIN Axiom.
@@
 Hence there would be a 010-coloring of the KS system, quod non.
-Therefore a deterministic noncontextual theory cannot satisfy the
+Therefore a deterministic non-contextual theory cannot satisfy the
 SPIN Axiom.
@@
 The squarefreeness is a considerable restriction.  There are
-only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$ vertices.
+only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$
+vertices.\cite{sf-sloane}
 We can restrict ourselves to connected graphs.
-% TODO OEIS A006786
 \begin{prop}[\cite{aow11}]\label{prop:ks-conn}
@@
 Let~$S''$ be~$S'$ rotated along the north pole at some angle~$\alpha$.
-There are finitely many angles such that there are overlaping points.
+There are finitely many angles such that there are overlapping points.
 Thus there is an angle such that~$S \cup S''$ is an embedding of~$G'$.
@@
 We are grateful to prof.~McKay for discussing
-the feasability of certain graph restrictions.
+the feasibility of certain graph restrictions.
 

### 2014-05-12 main.tex
Fix an overflow.
update biconnectedness proof

@@
 Given a finite subset~$S$ of the projective plane
-(or equivalently, of the northern hemisphere without equator).
+(or equivalently, a finite subset of the northern
+hemisphere without equator).
 Define its \keyword{orthogonality graph}~$G(S)$ as follows.
@@
 We are interested in these graphs because of the following observation.
-\begin{lem}
+\begin{lem}\label{lem:biconn}
 If there is an embeddable graph~$G$ on~$n$ vertices with a vertex
@@
 
-Furthermore suppose~$G$ is a minimal Kochen-Specker graph.
+Furthermore suppose, reasoning towards contradication,
+that~$G$ is a minimal Kochen-Specker graph.
 Now consider a leaf~$A$ in the biconnected component graph~$B(G)$.
@@
 Similarly~$B$ is an embeddable graph with a vertex with fixed color~$1$.
-We may assume~$G$ is two copies of~$A$ connected at the
-copies of the vertex with fixed color~$1$.
 
-The graph~$G$ has minimal vertex-order~$3$.
+Let~$A'$ denote the graph containing two copies of~$A$
+connected at the copies of the vertex with fixed color~$1$.
+Using the reasoning in the proof of Lemma~\ref{lem:biconn},
+we see~$A'$ is a Kochen-Specker graph that is not biconnected.
+Define~$B'$ similarly. Observe that~$A'$ and~$B'$ must have the same
+number of vertices as~$G$ by the minimality of~$G$.
+
+The graph~$A'$ has minimal vertex-order~$3$.
 Thus every vertex of~$A$ has minimal order~$3$ except for~$a$,
@@
 the graph~$A$ has at least~$16$ vertices.
-Thus~$G$ has at least~$32$ vertices, quod non.
+Thus~$A'$ has at least~$32$ vertices, quod non.
 
@@
 the graph~$A$ has at least~$15+1$ vertices.
-Thus~$G$ has at least~$32$ vertices. Contradiction.
+Thus~$A'$ has at least~$32$ vertices. Contradiction.
 \end{proof}

### 2014-05-20 code/subgraph.py
code: subgraph: use igraph

@@ -1,40 +1,19 @@
-from nauty import orbits_of
+import igraph
 
-def is_subgraph_of(smallg, largeg):
-    """ Checks whether smallg is a subgraph of largeg """
-    todo = list(smallg)
-    avail = frozenset(largeg)
-    todo.sort(key=lambda node: -len(smallg[node]))
-    mapping = {}
-    for node in smallg:
-        mapping[node] = None
-    stack = [(mapping, tuple(todo), avail)]
-    N = 0
-    while stack:
-        N += 1
-        mapping, todo, avail = stack.pop()
-        if not todo:
-            return mapping
-        node = todo[0]
-        todo = todo[1:]
-        orbits = orbits_of(largeg, set(largeg)-avail)
-        for orbit in orbits:
-            target = tuple(orbit)[0]
-            if target not in avail:
-                continue
-            if len(largeg[target]) < len(smallg[node]):
-                continue
-            ok = True
-            for neighbour in smallg[node]:
-                if mapping[neighbour] is None:
-                    continue
-                if mapping[neighbour] not in largeg[target]:
-                    ok = False
-                    break
-            if not ok:
-                continue
-            new_mapping = dict(mapping)
-            new_avail = avail - frozenset([target])
-            new_mapping[node] = target
-            stack.append((new_mapping, todo, new_avail))
-    return False
+def to_igraph(g):
+    ret = igraph.Graph()
+    ret.add_vertices(len(g))
+    for v in g:
+        for w in g[v]:
+            if v >= w: continue
+            ret.add_edges([(v, w)])
+    return ret
+
+def find_mono(smallg, largeg):
+    """ Tries to find an embedding of smallg into largeg.
+        Returns None if there is none. """
+    found_one, mapping = to_igraph(largeg).subisomorphic_lad(to_igraph(smallg),
+                            return_mapping=True)
+    if not found_one:
+        return None
+    return mapping

### 2014-05-20 code/sphereEmbeddable.py
code: add sphere embeddability checker

@@ -1,0 +1,291 @@
+""" Tries to determine, using quantifier elimination, whether a graph is
+    embeddable on the sphere. """
+
+import StringIO
+import random
+
+def find_assignments(g):
+    """ Find all possible crossproduct assignments for the graph g. """
+    # TODO compute automorphism group first?
+    # Create list of edges
+    edges_without_duplicates = set()
+    edges = set()
+    for v in g:
+        for w in g[v]:
+            if (w, v) not in edges:
+                edges_without_duplicates.add((v, w))
+            edges.add((v,w))
+    # Fill stack with choices of initial edge
+    frame = collections.namedtuple('frame',
+                                ('first_edge',
+                                 'to_visit',
+                                 'nvar',
+                                 'ortho',
+                                 'eqs',
+                                 'var',
+                                 'assign',
+                                 'easily_embeddable',
+                                 'iterative_assignment',
+                                 'base',
+                                 'touched_by',
+                                 'edges_used'))
+    batch = list()
+    for first_edge in edges_without_duplicates:
+        f = frame(
+                first_edge=first_edge,
+                to_visit=set(first_edge),
+                nvar=0,
+                ortho=[],
+                eqs=[],
+                var=[first_edge[0], first_edge[1]],
+                assign={first_edge[0]: 0,
+                        first_edge[1]: 1},
+                easily_embeddable=True,
+                iterative_assignment=True,
+                base=set(first_edge),
+                touched_by={first_edge[0]: first_edge[1],
+                            first_edge[1]: first_edge[0]},
+                edges_used=set([first_edge, (first_edge[1], first_edge[0])]))
+        for z in (frozenset(g[first_edge[0]]) &
+                  frozenset(g[first_edge[1]])):
+            f.base.add(z)
+            f.assign[z] = (0, 1)
+            f.edges_used.add((first_edge[0], z))
+            f.edges_used.add((first_edge[1], z))
+            f.edges_used.add((z, first_edge[0]))
+            f.edges_used.add((z, first_edge[1]))
+            f.to_visit.add(z)
+            f.touched_by[z] = first_edge[0]
+        batch.append(f)
+    completed = []
+    while batch:
+        new_batch = []
+        for f in batch:
+            # First, deduce all assignments
+            while f.to_visit:
+                v = f.to_visit.pop()
[…]

### 2014-05-20 code/nauty.py
code: nauty: add canonical_labeling option

@@ -6,12 +6,14 @@ import subprocess
 from helper import load_graph6
 
 def geng(until=None, minimal_vertex_degree=3, squarefree=True,
-                            connected=False):
+                            connected=False, canonical_labeling=False):
     """ Generates graphs using the geng util """
     i = 0
     while True:
         i += 1
         args = '-d%s' % minimal_vertex_degree
+        if canonical_labeling:
+            args += 'l'
         if squarefree:
             args += 'f'
         if connected:

### 2014-05-20 code/vectorsystem.py
code: add vectorsystem

@@ -1,0 +1,27 @@
+
+def orthogonality_graph(vs):
+    """ Returns the orthogonality graph of a set of vectors. """
+    g = dict()
+    for i in xrange(len(vs)):
+        g[i] = []
+    for i, v in enumerate(vs):
+        for j, w in enumerate(vs):
+            if j >= i:
+                continue
+            if v[0] * w[0] + v[1]*w[1] + v[2]*w[2] == 0:
+                g[i].append(j)
+                g[j].append(i)
+    return g
+
+def full_grid(n):
+    """ Returns all vectors on the n x n x n grid. """
+    ret = list()
+    for rv1 in xrange(n):
+        for rv2 in xrange(n):
+            v1 = 2*rv1 - n + 1
+            v2 = 2*rv2 - n + 1
+            ret.append((v1,v2,n-1))
+            ret.append((v1,n-1,v2))
+            ret.append((n-1,v1,v2))
+    ret = list(set([x if x[0] >= 0 else tuple([-y for y in x]) for x in ret]))
+    return ret

### 2014-05-20 code/sphereEmbeddable.py
code: fix sphereEmbeddable

@@ -1,7 +1,11 @@
 """ Tries to determine, using quantifier elimination, whether a graph is
     embeddable on the sphere. """
 
+import collections
+import subprocess
+import threading
 import StringIO
+import textwrap
 import random
 
 def find_assignments(g):
@@ -105,8 +109,8 @@ def find_assignments(g):
                 for v1, v2 in edges_without_duplicates - f.edges_used:
                     f.ortho.append((f.assign[v1], f.assign[v2]))
                     f = f._replace(easily_embeddable=False)
-                if f.easily_embeddable:
-                    return (True, 'easily')
+                #if f.easily_embeddable:
+                #    return (True, 'easily')
                 completed.append(f)
                 continue
             # If not: consider every possible node for the new variable
@@ -159,6 +163,11 @@ def check_sphere_embeddability(g, assignment, guess=False):
         find out, and if it does, it will be quite a bit faster than
         without guessing. """
     # Generate the reduce script.
+    edges = set()
+    for v in g:
+        for w in g[v]:
+            edges.add((v,w))
+    f = assignment
     def assign_to_reduce(x):
         if isinstance(x, tuple):
             return 'k({},{})'.format(*map(assign_to_reduce, x))
@@ -271,9 +280,9 @@ def check_sphere_embeddability(g, assignment, guess=False):
             embed = (bit == 'true')
             break
     assert embed is not None
-    if not embed and guessing:
+    if not embed and guess:
         return None
-    return embed
+    return (embed, io.getvalue())
 
 def popen_communicate_timeout(pipe, s, timeout=2.5):
     """ Communicates with a Popen object with a timeout """

### 2014-05-22 code/comp1.py
code: add comp1

@@ -1,0 +1,130 @@
+""" Computation 1: decides for square-free graphs of minimal vertex order 3,
+    whether they are embeddable. """
+
+import os
+import json
+import binascii
+
+import nauty
+from helper import write_graph6, load_graph6
+from vectorsystem import full_grid, orthogonality_graph
+from subgraph import find_mono
+from sphereEmbeddable import find_assignments, check_sphere_embeddability
+
+class LenientJsonEncoder(json.JSONEncoder):
+    def default(self, obj):
+        if isinstance(obj, set):
+            return list(obj)
+        return json.JSONEncoder.default(self, obj)
+
+def check_for_result(graph):
+    if os.path.exists("../graphs/class/%s.json" % binascii.hexlify(
+                                                    write_graph6(graph))):
+        return True
+    return False
+
+def write_result(graph, result):
+    with open('../graphs/class/%s.json' % binascii.hexlify(write_graph6(graph)),
+            'w') as f:
+        json.dump(result, f, cls=LenientJsonEncoder)
+
+def main():
+    minimal_unembeddable = []
+    print 'Loading previously computed minimally unembeddable fd3 graphs'
+    with open('../graphs/min-unemb-fd3.g6', 'r') as f:
+        for l in f:
+            minimal_unembeddable.append(load_graph6(l[:-1]))
+    print 'Computing grids'
+    grids = []
+    for i in xrange(10):
+        vectors = full_grid(2*i+3)
+        grids.append((orthogonality_graph(vectors),
+                        vectors))
+    print 'Considering graphs'
+    for graph in nauty.geng(15, canonical_labeling=True):
+        if check_for_result(graph):
+            continue
+        result = {}
+        print write_graph6(graph)
+        found_unemb_subgraph = False
+        print ' checking for unembeddable subgraph'
+        for unemb in minimal_unembeddable:
+            mono = find_mono(unemb, graph)
+            if mono is not None:
+                result['embeddable'] = False
+                result['reason'] = 'unembeddable-subgraph'
+                result['subgraph'] = write_graph6(unemb)
+                result['mono'] = mono
+                print '  yes'
+                found_unemb_subgraph = True
+                break
+        if found_unemb_subgraph:
+            write_result(graph, result)
+            continue
+        print ' computing cross-product assignments'
+        assignments = find_assignments(graph)
+        assignment_index = 0
+        guessing = False
+        grid_index = 0
+        embeddable = None
+        decided_embeddability = False
+        decided_grid_embeddability = False
[…]

### 2014-05-23 code/comp1.py
code: update comp1 and sphereEmbeddable

@@ -30,9 +30,11 @@ def write_result(graph, result):
 
 def main():
     minimal_unembeddable = []
+    minimal_unembeddable_names = set()
     print 'Loading previously computed minimally unembeddable fd3 graphs'
     with open('../graphs/min-unemb-fd3.g6', 'r') as f:
         for l in f:
+            minimal_unembeddable_names.add(l[:-1])
             minimal_unembeddable.append(load_graph6(l[:-1]))
     print 'Computing grids'
     grids = []
@@ -49,6 +51,7 @@ def main():
         found_unemb_subgraph = False
         print ' checking for unembeddable subgraph'
         for unemb in minimal_unembeddable:
+            assert write_graph6(graph) not in minimal_unembeddable_names
             mono = find_mono(unemb, graph)
             if mono is not None:
                 result['embeddable'] = False

### 2014-05-23 code/sphereEmbeddable.py

@@ -279,7 +279,8 @@ def check_sphere_embeddability(g, assignment, guess=False):
         if bit in ('false', 'true'):
             embed = (bit == 'true')
             break
-    assert embed is not None
+    if embed is None:
+        return None
     if not embed and guess:
         return None
     return (embed, io.getvalue())

### 2014-05-30 code/comp1.py
comp1: check up to 14

@@ -43,7 +43,7 @@ def main():
         grids.append((orthogonality_graph(vectors),
                         vectors))
     print 'Considering graphs'
-    for graph in nauty.geng(15, canonical_labeling=True):
+    for graph in nauty.geng(14, canonical_labeling=True):
         if check_for_result(graph):
             continue
         result = {}

### 2014-05-30 code/comp2.py
code: add comp2

@@ -1,0 +1,59 @@
+""" Computation 2: for those graphs that were found embeddable, but not
+        <10-grid embeddable in computation 1, check <20-grid embeddable """
+
+
+import os
+import json
+import msgpack
+import binascii
+
+from helper import write_graph6, load_graph6
+from vectorsystem import full_grid, orthogonality_graph
+from subgraph import find_mono
+
+def load_grids():
+    path = '../graphs/grids.msgpack'
+    if os.path.exists(path):
+        print 'Loading grids'
+        with open(path) as f:
+            return msgpack.load(f)
+    print 'Computing grids'
+    grids = []
+    for i in xrange(24):
+        vectors = full_grid(2*i+3)
+        grids.append((orthogonality_graph(vectors),
+                        vectors))
+        print ' %s: %s vectors' % (i, len(vectors))
+    with open(path, 'w') as f:
+        msgpack.dump(grids, f)
+    return grids
+
+def main():
+    grids = load_grids()
+    dirpath = '../graphs/class'
+    for fn in os.listdir(dirpath):
+        path = os.path.join(dirpath, fn)
+        with open(path) as f:
+            result = json.load(f)
+            if (result['reason'] != 'quantifier-elimination' or
+                    not result['embeddable']):
+                continue
+            graph = load_graph6(binascii.unhexlify(fn.split('.')[0]))
+            print fn
+            for i, grid in tuple(enumerate(grids))[9:]:
+                print i
+                grid_graph, grid_vectors = grid
+                mono = find_mono(graph, grid_graph)
+                if mono is None:
+                    continue
+                print ' yes!'
+                result['reason'] = 'grid-embeddable'
+                result['gridEmbeddable'] = True
+                result['grid'] = i
+                result['embedding'] = [grid_vectors[x] for x in mono]
+                with open(path, 'w') as f:
+                    json.dump(result, f)
+                break
+
+if __name__ == '__main__':
+    main()

### 2014-05-30 code/generateSite.py
code: add generateSite

@@ -1,0 +1,64 @@
+""" Generates the graph pages on the site. """
+
+import os
+import json
+import binascii
+
+from helper import load_graph6
+
+import yaml
+import pydot
+
+class literal_str(unicode): pass
+def yaml_literal_str_representer(dumper, data):
+    return dumper.represent_scalar(u'tag:yaml.org,2002:str', data, style='|')
+yaml.add_representer(literal_str, yaml_literal_str_representer)
+
+def main():
+    dirname = '../graphs/class'
+    for fn in os.listdir(dirname):
+        path = os.path.join(dirname, fn)
+        with open(path) as f:
+            result = json.load(f)
+        graph_safe_name = fn.split('.')[0]
+        graph_name = binascii.unhexlify(graph_safe_name)
+        print graph_name
+        graph = load_graph6(graph_name)
+        if len(graph) >= 14:
+            continue
+        dot = pydot.Dot('G', graph_type='graph', splines='true')
+        for v in graph:
+            dot.add_node(pydot.Node(str(v)))
+        for v in graph:
+            for w in graph[v]:
+                if v >= w:
+                    continue
+                dot.add_edge(pydot.Edge(str(v), str(w)))
+        image = graph_safe_name + '.svg'
+        dot.write_svg(os.path.join('../site/smallGraphs/', image), prog='fdp')
+        front_matter = {
+                'layout': 'smallGraph',
+                'unembeddableSubgraph': (result['subgraph'],
+                                         binascii.hexlify(result['subgraph']),
+                                         result['mono'])
+                                            if 'subgraph' in result else None,
+                'title': graph_name,
+                'image': image,
+                'reduce_script': literal_str(result['reduce-script'])
+                                if 'reduce-script' in result else None,
+                'safe_name': graph_safe_name,
+                'n_vertices': len(graph),
+                'embeddable': result.get('embeddable'),
+                'minimally_unembeddable':
+                        result.get('reason') == 'quantifier-elimination',
+                'gridEmbeddable': result['grid']*2+3
+                        if 'grid' in result else False,
+                'embedding': result.get('embedding')}
+        with open('../site/smallGraphs/%s.markdown' % graph_safe_name, 'w') as f:
+            f.write('---\n')
+            yaml.dump(front_matter, f, default_flow_style=False)
+            f.write('---\n')
+
+
+if __name__ == '__main__':
+    main()

### 2014-05-30 code/generateSite.py
code/generateSite: show progress

@@ -16,13 +16,15 @@ yaml.add_representer(literal_str, yaml_literal_str_representer)
 
 def main():
     dirname = '../graphs/class'
-    for fn in os.listdir(dirname):
+    fns = os.listdir(dirname)
+    for i, fn in enumerate(fns):
+        if i % 10 == 0:
+            print '%s/%s' % (i, len(fns))
         path = os.path.join(dirname, fn)
         with open(path) as f:
             result = json.load(f)
         graph_safe_name = fn.split('.')[0]
         graph_name = binascii.unhexlify(graph_safe_name)
-        print graph_name
         graph = load_graph6(graph_name)
         if len(graph) >= 14:
             continue

### 2014-06-01 code/comp3.py
code: add comp3

@@ -1,0 +1,38 @@
+""" Computation 3: check whether the candidate graphs contain unembeddable
+        subgraphs. """
+
+from helper import write_graph6, load_graph6
+from subgraph import find_mono
+
+import binascii
+import json
+
+def main():
+    candidates = []
+    unemb = []
+    print 'Reading candidates ...'
+    with open('../graphs/candidates.g6') as f:
+        for l in f:
+            candidates.append((l[:-1], load_graph6(l[:-1])))
+    print 'Reading minimal unembeddable graphs ...'
+    with open('../graphs/min-unemb-fd3.g6') as f:
+        for l in f:
+            unemb.append((l[:-1], load_graph6(l[:-1])))
+    print 'Checking subgraphs ...'
+    for i, g_pair in enumerate(candidates):
+        gname, g = g_pair
+        if i % 100 == 0:
+            print i, len(candidates)
+        for j, ug_pair in enumerate(unemb):
+            ugname, ug = ug_pair
+            mono = find_mono(ug, g)
+            if mono is None:
+                continue
+            break
+        gsname = binascii.hexlify(gname)
+        with open('../graphs/candidates/%s.json' % gsname, 'w') as f:
+            json.dump({'subgraph': ugname,
+                       'mono': mono}, f)
+
+if __name__ == '__main__':
+    main()

### 2014-06-01 code/generateSite.py
site: updates.

@@ -14,7 +14,46 @@ def yaml_literal_str_representer(dumper, data):
     return dumper.represent_scalar(u'tag:yaml.org,2002:str', data, style='|')
 yaml.add_representer(literal_str, yaml_literal_str_representer)
 
-def main():
+def generate_candidates():
+    dirname = '../graphs/candidates'
+    fns = os.listdir(dirname)
+    for i, fn in enumerate(fns):
+        if i % 10 == 0:
+            print '%s/%s' % (i, len(fns))
+        path = os.path.join(dirname, fn)
+        with open(path) as f:
+            result = json.load(f)
+        graph_safe_name = fn.split('.')[0]
+        graph_name = binascii.unhexlify(graph_safe_name)
+        graph = load_graph6(graph_name)
+        if len(graph) >= 21:
+            break
+        dot = pydot.Dot('G', graph_type='graph', splines='true')
+        for v in graph:
+            dot.add_node(pydot.Node(str(v)))
+        for v in graph:
+            for w in graph[v]:
+                if v >= w:
+                    continue
+                dot.add_edge(pydot.Edge(str(v), str(w)))
+        image = graph_safe_name + '.svg'
+        dot.write_svg(os.path.join('../site/candidates/', image), prog='fdp')
+        front_matter = {
+                'layout': 'candidate',
+                'unembeddableSubgraph': (result['subgraph'],
+                                         binascii.hexlify(result['subgraph']),
+                                         result['mono'])
+                                            if 'subgraph' in result else None,
+                'title': graph_name,
+                'image': image,
+                'safe_name': graph_safe_name,
+                'n_vertices': len(graph)}
+        with open('../site/candidates/%s.markdown' % graph_safe_name, 'w') as f:
+            f.write('---\n')
+            yaml.dump(front_matter, f, default_flow_style=False)
+            f.write('---\n')
+
+def generate_smallGraphs():
     dirname = '../graphs/class'
     fns = os.listdir(dirname)
     for i, fn in enumerate(fns):
@@ -27,7 +66,7 @@ def main():
         graph_name = binascii.unhexlify(graph_safe_name)
         graph = load_graph6(graph_name)
         if len(graph) >= 14:
-            continue
+            break
         dot = pydot.Dot('G', graph_type='graph', splines='true')
         for v in graph:
             dot.add_node(pydot.Node(str(v)))
@@ -61,6 +100,11 @@ def main():
[…]

### 2014-09-08 paper/main.tex
Move paper into subdirectory
paper: convert to eptcs

--- 2014-05-12 main.tex
@@
-\documentclass[a4paper]{article}
+\documentclass[adraft,copyright,creativecommons]{eptcs}
 
 \usepackage{algorithm2e}
+\usepackage{breakurl}
 \usepackage{hyperref}
@@
 \usepackage{graphicx}
-\usepackage{authblk}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
-
-\usepackage{draftwatermark}
 
@@
 
-\title{A Kochen-Specker system has at least 21 vertices\\
+\title{A Kochen-Specker system has at least 21 vectors\\
         {\small (extended abstract)}}
+\def\titlerunning{A Kochen-Specker system has at least 21 vectors
+            (extended abstract)}
+\def\authorrunning{Uijlen \& Westerbaan}
 
-\author{Sander Uijlen}
-\author{Bas Westerbaan}
-
-% TODO will we keep these @cs.ru.nl addresses?
-\affil{{\small Institute for Computing and Information Sciences\\
-        Radboud Universiteit Nijmegen}\\
-   \{\texttt{suijlen},\texttt{bwesterb}\}\texttt{@cs.ru.nl}}
+\author{Sander Uijlen
+    \institute{Radboud Universiteit}
+    \email{suijlen@cs.ru.nl}
+\and
+    Bas Westerbaan
+    \institute{Radboud Universiteit}
+    \email{bwesterb@cs.ru.nl}}
 
@@
 \begin{center}
-\includegraphics[width=50mm]{graphs/c17.jpg}
+\includegraphics[width=50mm]{../graphs/c17.jpg}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=50mm]{graphs/unemb-10-2.jpg}
+\includegraphics[width=50mm]{../graphs/unemb-10-2.jpg}
 \end{center}
@@
     \begin{center}
-        \includegraphics[width=120mm]{graphs/unemb-base-20.jpg}
+        \includegraphics[width=120mm]{../graphs/unemb-base-20.jpg}
     \end{center}
@@
 \bibliography{main}{}
-\bibliographystyle{plain}
+\bibliographystyle{eptcs}
 

### 2014-09-08 paper/main.tex
paper: two fixes
paper: improve introduction

@@
 \maketitle
-\thispagestyle{empty}
 
@@
 
-% TODO correct description of experiment
-Consider the following experiment.  Shoot a deuterium atom,
-or any other spin-1 particle,
-along, say: the x-axis, through a inhomogeneous magnetic field.
-Depending on the direction of the magnetic field,
-the particle will move undisturbed
-or deviate.
+Consider the following experiment. Measure the SPIN of a spin 1
+particle along a direction. Or more concretely: shoot a deuterium
+atom through a certain fixed inhomogeneous magnetic field. Depending
+on the direction it was shot through the field, the particle will
+move undisturbed or deviate.
 
-Quantum Mechanics only predicts the probability, given the configuration
-of the field, whether the particle will deviate.
+Quantum Mechanics only predicts the probability, given the direction,
+whether the particle will deviate.
 Its probabilistic prediction has been thoroughly tested.
-One wonders: is there a deterministic non-contextual theory predicting the
+One wonders: is there a deterministic theory predicting the
 outcome of this experiment?
 
-Kochen and Specker proved that such a theory cannot satisfy:
+Kochen and Specker have shown that such a deterministic theory must
+be odd: it cannot satisfy the plausible SPIN axiom, that is:
 \begin{spin}

### 2014-09-08 paper/main.tex
paper: update figures.  add attribution to Judith

@@
 
+\begin{wrapfigure}{r}{0.30\textwidth}
+\begin{center}
+\includegraphics[width=40mm]{../graphs/ks31.pdf}
+\end{center}
+\caption{Conway's 31 vector Kochen-Specker system
+\label{fig:conway31}}
+\end{wrapfigure}
 Suppose there is a KS system and a non-contextual deterministic theory satisfying
@@
 \section{Embeddability}
-\label{sec:emb}
-\begin{wrapfigure}{r}{0.40\textwidth}
-\begin{center}
-\includegraphics[width=50mm]{../graphs/unemb-10-2.jpg}
-\end{center}
-\caption{One of the two minimal
-        non-embeddable graphs
-\label{fig:unemb-10-2}}
-\end{wrapfigure}
-
 Our computation has yielded a few hundred non-010-colorable graphs.
@@
 % TODO give intermediate computation.
+\label{sec:emb}
+\begin{wrapfigure}{r}{0.30\textwidth}
+\begin{center}
+\includegraphics[width=50mm]{../graphs/unemb10-1-a.pdf}
+\end{center}
+\caption{One of the two minimal
+        non-embeddable graphs
+\label{fig:unemb-10-2}}
+\end{wrapfigure}
 
@@
     \begin{center}
-        \includegraphics[width=120mm]{../graphs/unemb-base-20.jpg}
+        \includegraphics[width=30mm]{../graphs/unemb10-1.pdf} \qquad
+        \includegraphics[width=30mm]{../graphs/unemb10-2.pdf} \qquad
+        \includegraphics[width=30mm]{../graphs/unemb10-3.pdf}
     \end{center}
@@
 We are grateful to prof.~McKay for discussing
-the feasibility of certain graph restrictions.
+the feasibility of certain graph restrictions and
+to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
+for drawing the figures in this paper.
 

### 2014-09-08 paper/main.tex
paper: add event; add forward reference

@@
 \documentclass[adraft,copyright,creativecommons]{eptcs}
+
+\providecommand{\event}{QPL2014}
 
@@
 \title{A Kochen-Specker system has at least 21 vectors\\
-        {\small (extended abstract)}}
+        {\small (extended abstract)
+            \footnote{This is a condensed version of the article
+                        to be published in Ohmsha-Springer's
+                        New Generation Computing.
+                    One can find a preprint of the full version at
+                    \url{http://westerbaan.name/\~bas/math/ks.pdf}}}}
 \def\titlerunning{A Kochen-Specker system has at least 21 vectors

### 2014-09-08 paper/main.tex
paper: add crossproduct illustration

@@
 that is not 010-colorable.
+\begin{wrapfigure}{r}{0.25\textwidth}
+\begin{center}
+\includegraphics[width=40mm]{../graphs/kruisproduct.pdf}
+\end{center}
+\caption{squarefreeness of an embeddable graph
+\label{fig:crossproduct}}
+\end{wrapfigure}
 \begin{dfn}
@@
 \begin{proof}
-    Given two non antipodal points~$a\neq b$.
-    Consider the points orthogonal to~$a$.
+    Given two non antipodal points~$v\neq w$.
+    See Figure~\ref{fig:crossproduct}.
+    Consider the points orthogonal to~$v$.
     This is a great circle.
-    The points orthogonal to~$b$ is a different great circle.
+    The points orthogonal to~$w$ is a different great circle.
     They intersect in precisely two antipodal points.
-    Hence, if~$c$ and~$d$ are both orthogonal to~$a$ and~$b$,
+    Hence, if~$c$ and~$d$ are both orthogonal to~$v$ and~$w$,
     then~$c$ and~$d$ are equivalent.

### 2014-09-08 paper/main.tex
paper: 21 -> 22

--- 2014-09-08 paper/main.tex
@@
 
-% TODO replace 21 with 22
 \newcommand{\keyword}[1]{\textbf{#1}}
@@
 
-\title{A Kochen-Specker system has at least 21 vectors\\
-        {\small (extended abstract)
+\title{A Kochen-Specker system has at least 22 vectors\\
+        {\small (extended abstract)%
             \footnote{This is a condensed version of the article
@@
                     \url{http://westerbaan.name/\~bas/math/ks.pdf}}}}
-\def\titlerunning{A Kochen-Specker system has at least 21 vectors
+\def\titlerunning{A Kochen-Specker system has at least 22 vectors
             (extended abstract)}
@@
     we need to consider and develop a more practical decision procedure for
-    embeddability to improve the lower bound to 21.
+    embeddability to improve the lower bound to 22.
 \end{abstract}
@@
     there is no~\keyword{010-coloring}.
-    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
+    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that%
         \footnote{
@@
 \end{wrapfigure}
+
 Suppose there is a KS system and a non-contextual deterministic theory satisfying
@@
 that a KS system must have at least 18 vectors.  We improve their lower bound
-and show that a KS system must have at least 21 vectors.
+and show that a KS system must have at least 22 vectors.
 
@@
 
+\begin{wrapfigure}{r}{0.25\textwidth}
+\begin{center}
+\includegraphics[width=40mm]{../graphs/kruisproduct.pdf}
+\end{center}
+\caption{squarefreeness of an embeddable graph
+\label{fig:crossproduct}}
+\end{wrapfigure}
+
 \subsection{Kochen-Specker graphs}
@@
 that is not 010-colorable.
+
 \begin{dfn}
@@
     An embeddable graph is squarefree.
-    That is: it does not contain the square as a subgraph.
+    That is: it does not contain the square as a subgraph.%
     \footnote{Some authors call a graph squarefree if it does not
@@
     \begin{center}
-    \begin{tabular}{l|lllll}
+    \begin{tabular}{l|llllll}
         $n$ & $\leq 16$
@@
             & $19$
-            & $20$ \\
+            & $20$
+            & $21$ \\
         \hline
@@
             & $441$
+            & $11876$
     \end{tabular}
@@
 
-    All these 463 graphs are not embeddable.
+    All these 12339 graphs are not embeddable.
     See Computation~\ref{comp:unemb20}.
 \end{comp}
-The computation took roughly a week on a 64-core Opteron 6276.
+The computation was distributed on approximately 300 CPU cores
+and took roughly three months.
 It was executed as follows.
 We enumerated all squarefree graphs with minimal vertex
-order 3 on less than or equal~$20$ vertices,
+order 3 on less than or equal~$21$ vertices,
 using the~\texttt{geng} util of the nauty software package,
@@
 % TODO give intermediate computation.
+
 \label{sec:emb}
@@
 
-With this method, we have decided in an hour the embeddability
-of every squarefree graph with minimal vertex order three of 13
-vertices of less.
+With this method, we have decided in an a day the embeddability
+of every squarefree graph with minimal vertex order three of 14
+vertices of less except for one.\footnote{
+A list of all squarefree graphs with minimal vertex order three of
+14 vertices of less together with their embeddability can
+be found here:
+\url{http://kochen-specker.info/smallGraphs/}
+}
 In particular:
@@
     are the only minimal unembeddable squarefree graph.
-\end{comp}
 For the first graph, we have proven directly that it is unembeddable.
@@
 For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.
+
+Every squarefree graph of minimal vertex order three that is
+not~$010$-colorable of order 21 contains an unembeddable subgraph.%
+\footnote{A list of these graphs together with their unembeddable
+subgraphs, can be found here:
+\url{http://kochen-specker.info/candidates/}}
+\end{comp}
 
@@
 Then we enumerated all candidate graphs
-up to and including 20 vertices.
+up to and including 21 vertices.
 Furthermore, we have proposed a new decision procedure,
@@
 Therefore, we demonstrate: a Kochen-Specker system must have at least
-21 points.
+22 points.
 At the time of writing,
 we are computing
-whether there is a KS system of 21 vertices.\footnote{
+whether there is a KS system of 22 vertices.\footnote{
 The authors have a wager whether there is a minimal KS system of less
@@
 is computationally infeasable.
-To bridge the enormous the gap between 21 and 31,
+To bridge the enormous the gap between 22 and 31,
 requires a new insight.

### 2014-09-09 paper/main.tex
paper: TeXify the 17 vertex unembeddable graph
paper: convert minimally unembeddable graphs
paper: replace unemb10-3.pdf
paper: replace unemb10-1-a

@@
 \usepackage{amsthm}
+\usepackage{tikz}
 
@@
 \newtheorem{rem}[main]{Remark}
+
+\tikzstyle{every node}=[circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt]
 
@@
 \begin{center}
-\includegraphics[width=50mm]{../graphs/c17.jpg}
+    \begin{tikzpicture}[thick,scale=1.4]
+        \draw
+            \foreach \x in {0,90,180,270} {
+                (\x+45:1) node{} -- (\x+65:1.3) node{}
+                (\x+45:1)  -- (\x+25:1.3) node{}
+                (\x+65:1.3)  -- (\x+25:1.3)
+            }
+            \foreach \x in {0,180} {
+                (\x:0.8) node{} -- (\x+25:1.3)
+                (\x:0.8) -- (\x-25:1.3)
+                (\x+25:1.3) -- (\x-25:1.3)
+                (\x+65:1.3) -- (\x+115:1.3)
+            }
+            (0:0.8) -- (180:0.8)
+            (90:0.2) node{} -- (0:0.8)
+            (90:0.2) -- (180:0.8)
+
+            (90:0.2) -- (90:0.6) node{}
+            (90:0.6) -- (270+45:1)
+            (90:0.6) -- (270-45:1)
+            (90:0.6) -- (90+45:1)
+            (90:0.6) -- (90-45:1)
+
+            (90:0.2) -- (270:0.4) node{}
+            (270:0.4) -- (270+25:1.3)
+            (270:0.4) -- (270-25:1.3)
+            (270:0.4) -- (90+25:1.3)
+            (270:0.4) -- (90-25:1.3)
+        ;
+    \end{tikzpicture}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=50mm]{../graphs/unemb10-1-a.pdf}
+        \begin{tikzpicture}[thick,scale=1.4]
+            \draw
+                \foreach \x in {90,210,330} {
+                    (\x:0.7) -- (0:0)
+                    (\x-30:1.3) -- (\x:0.7)
+                    (\x+30:1.3) -- (\x:0.7)
+                    (\x-30:1.3) -- (\x+30:1.3) 
+                    (\x+30:1.3) -- (\x-30+120:1.3) 
+                }
+                (90:0.7) node{$w$}
+                (210:0.7) node{$v$}
+                (330:0.7) node{$x$}
+                (0:0) node{$z$}
+                (90+30:1.3) node{$p_3$}
+                (90-30:1.3) node{$p_4$}
+                (210+30:1.3) node{$p_1$}
+                (210-30:1.3) node{$p_2$}
+                (330+30:1.3) node{$p_5$}
+                (330-30:1.3) node{$a$}
+                ;
+        \end{tikzpicture}
 \end{center}
@@
     \begin{center}
-        \includegraphics[width=30mm]{../graphs/unemb10-1.pdf} \qquad
-        \includegraphics[width=30mm]{../graphs/unemb10-2.pdf} \qquad
-        \includegraphics[width=30mm]{../graphs/unemb10-3.pdf}
+        \begin{tikzpicture}[thick,scale=1.4]
+            \draw
+                \foreach \x in {90,210,330} {
+                    (\x:0.7) node{} -- (0:0) node{}
+                    (\x-30:1.3) node{} -- (\x:0.7)
+                    (\x+30:1.3) node{} -- (\x:0.7)
+                    (\x-30:1.3) -- (\x+30:1.3) 
+                    (\x+30:1.3) -- (\x-30+120:1.3) 
+                }
+            ;
+        \end{tikzpicture}
+        \qquad
+        \begin{tikzpicture}[thick,scale=1.4]
+            \draw
+                \foreach \x in {0,180} {
+                    (\x+15:1.3) node{} -- (\x+90:0.9) node{}
+                    (\x+15:1.3) -- (\x+45:0.5) node{}
+                    (\x+45:0.5) -- (\x+90:0.9)
+
+                    (\x-15:1.3) node{} -- (\x-90:0.9)
+                    (\x-15:1.3) -- (\x-45:0.5) node{}
+                    (\x-45:0.5) -- (\x-90:0.9)
+
+                    (\x-15:1.3) -- (\x+15:1.3)
+                }
+                (45:0.5) -- (45+180:0.5)
+                (45+90:0.5) -- (45+90+180:0.5)
+
+            ;
+        \end{tikzpicture}
+        \qquad
+        \begin{tikzpicture}[thick,scale=1.4]
+            \draw
+                (45:0.5) node{} -- (135:0.5) node{}
+                                -- (225:0.5) node{}
+                                -- (315:0.5) node{}
+                (15:1.3) node{}
+                (345:1.3) node{}
+                (195:1.3) node{}
+                (165:1.3) node{}
+
+                (75:1) node{} -- (15:1.3) -- (45:0.5) -- (75:1)
+                (285:1) node{} -- (345:1.3) -- (315:0.5) -- (285:1)
+
+                (0:1.3) node{} -- (180:0.8) node{}
+                (0:1.3) -- (15:1.3)
+                (0:1.3) -- (345:1.3)
+
+                (180:0.8) -- (195:1.3)
+                (180:0.8) -- (165:1.3)
+                (195:1.3) -- (165:1.3)
+                (180:0.8) -- (225:0.5)
+                (180:0.8) -- (135:0.5)
+                (195:1.3) -- (285:1)
+                (165:1.3) -- (75:1)
+
+            ;
+        \end{tikzpicture}
     \end{center}

### 2014-09-09 paper/main.tex
paper: improve labels of figures

--- 2014-09-08 paper/main.tex
@@
 \begin{center}
-\includegraphics[width=50mm]{../graphs/c17.jpg}
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+        \draw
+            \foreach \x in {0,90,180,270} {
+                (\x+45:1) node[p]{} -- (\x+65:1.3) node[p]{}
+                (\x+45:1)  -- (\x+25:1.3) node[p]{}
+                (\x+65:1.3)  -- (\x+25:1.3)
+            }
+            \foreach \x in {0,180} {
+                (\x:0.8) node[p]{} -- (\x+25:1.3)
+                (\x:0.8) -- (\x-25:1.3)
+                (\x+25:1.3) -- (\x-25:1.3)
+                (\x+65:1.3) -- (\x+115:1.3)
+            }
+            (0:0.8) -- (180:0.8)
+            (90:0.2) node[p]{} -- (0:0.8)
+            (90:0.2) -- (180:0.8)
+
+            (90:0.2) -- (90:0.6) node[p]{}
+            (90:0.6) -- (270+45:1)
+            (90:0.6) -- (270-45:1)
+            (90:0.6) -- (90+45:1)
+            (90:0.6) -- (90-45:1)
+
+            (90:0.2) -- (270:0.4) node[p]{}
+            (270:0.4) -- (270+25:1.3)
+            (270:0.4) -- (270-25:1.3)
+            (270:0.4) -- (90+25:1.3)
+            (270:0.4) -- (90-25:1.3)
+        ;
+    \end{tikzpicture}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=50mm]{../graphs/unemb10-1-a.pdf}
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+            \draw
+                \foreach \x in {90,210,330} {
+                    (\x:0.7) -- (0:0)
+                    (\x-30:1.3) -- (\x:0.7)
+                    (\x+30:1.3) -- (\x:0.7)
+                    (\x-30:1.3) -- (\x+30:1.3) 
+                    (\x+30:1.3) -- (\x-30+120:1.3) 
+                }
+                (90:0.7) node[p]{}
+                (210:0.7) node[p]{}
+                (330:0.7) node[p]{}
+                (0:0) node[p]{}
+                (90+30:1.3) node[p]{}
+                (90-30:1.3) node[p]{}
+                (210+30:1.3) node[p]{}
+                (210-30:1.3) node[p]{}
+                (330+30:1.3) node[p]{}
+                (330-30:1.3) node[p]{}
+
+                (90:0.7) node[above]{$w$}
+                (210:0.7) node[above]{$v$}
+                (330:0.7) node[above]{$x$}
+                (0:0) node[below]{$z$}
+                (90+30:1.3) node[above]{$p_3$}
+                (90-30:1.3) node[above]{$p_4$}
+                (210+30:1.3) node[below]{$p_1$}
+                (210-30:1.3) node[left]{$p_2$}
+                (330+30:1.3) node[right]{$p_5$}
+                (330-30:1.3) node[below]{$a$}
+                ;
+        \end{tikzpicture}
 \end{center}
@@
     \begin{center}
-        \includegraphics[width=30mm]{../graphs/unemb10-1.pdf} \qquad
-        \includegraphics[width=30mm]{../graphs/unemb10-2.pdf} \qquad
-        \includegraphics[width=30mm]{../graphs/unemb10-3.pdf}
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+            \draw
+                \foreach \x in {90,210,330} {
+                    (\x:0.7) node[p]{} -- (0:0) node[p]{}
+                    (\x-30:1.3) node[p]{} -- (\x:0.7)
+                    (\x+30:1.3) node[p]{} -- (\x:0.7)
+                    (\x-30:1.3) -- (\x+30:1.3) 
+                    (\x+30:1.3) -- (\x-30+120:1.3) 
+                }
+            ;
+        \end{tikzpicture}
+        \qquad
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+            \draw
+                \foreach \x in {0,180} {
+                    (\x+15:1.3) node[p]{} -- (\x+90:0.9) node[p]{}
+                    (\x+15:1.3) -- (\x+45:0.5) node[p]{}
+                    (\x+45:0.5) -- (\x+90:0.9)
+
+                    (\x-15:1.3) node[p]{} -- (\x-90:0.9)
+                    (\x-15:1.3) -- (\x-45:0.5) node[p]{}
+                    (\x-45:0.5) -- (\x-90:0.9)
+
+                    (\x-15:1.3) -- (\x+15:1.3)
+                }
+                (45:0.5) -- (45+180:0.5)
+                (45+90:0.5) -- (45+90+180:0.5)
+
+            ;
+        \end{tikzpicture}
+        \qquad
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+            \draw
+                (45:0.5) node[p]{} -- (135:0.5) node{}
+                                -- (225:0.5) node{}
+                                -- (315:0.5) node{}
+                (15:1.3) node[p]{}
+                (345:1.3) node[p]{}
+                (195:1.3) node[p]{}
+                (165:1.3) node[p]{}
+
+                (75:1) node[p]{} -- (15:1.3) -- (45:0.5) -- (75:1)
+                (285:1) node[p]{} -- (345:1.3) -- (315:0.5) -- (285:1)
+
+                (0:1.3) node[p]{} -- (180:0.8) node[p]{}
+                (0:1.3) -- (15:1.3)
+                (0:1.3) -- (345:1.3)
+
+                (180:0.8) -- (195:1.3)
+                (180:0.8) -- (165:1.3)
+                (195:1.3) -- (165:1.3)
+                (180:0.8) -- (225:0.5)
+                (180:0.8) -- (135:0.5)
+                (195:1.3) -- (285:1)
+                (165:1.3) -- (75:1)
+
+            ;
+        \end{tikzpicture}
     \end{center}

### 2014-09-09 paper/main.tex
paper: fix Judith's attribution

@@
 to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
-for drawing the figures in this paper.
+for drawing the first two figures in this paper.
 

### 2014-09-09 paper/main.tex
paper: replace kruisproduct

@@
 \begin{center}
-\includegraphics[width=40mm]{../graphs/kruisproduct.pdf}
+\begin{tikzpicture}[scale=1.9,thick,
+    partial ellipse/.style args={#1:#2:#3}{
+            insert path={+ (#1:#3) arc (#1:#2:#3)}
+        }]
+    \draw (0,0) [partial ellipse=360:180:1 and 0.3] ;
+    \draw[gray,dashed] (0,0) [partial ellipse=0:180:1 and 0.3] ;
+    \draw (0,0) [rotate=315,partial ellipse=360:180:1 and 0.3] ;
+    \draw[gray,dashed] (0,0) [rotate=315,partial ellipse=0:180:1 and 0.3] ;
+    \draw (0,0) circle (1) ;
+    \node[circle,inner sep=1pt,fill] at (90:1) {};
+    \node[above] at (90:1) {$v$};
+    \node[circle,inner sep=1pt,fill] at (45:1) {};
+    \node[right] at (45:1) {$w$};
+\end{tikzpicture}
 \end{center}
@@
             \draw
-                (45:0.5) node[p]{} -- (135:0.5) node{}
-                                -- (225:0.5) node{}
-                                -- (315:0.5) node{}
+                (45:0.5) node[p]{} -- (135:0.5) node[p]{}
+                                -- (225:0.5) node[p]{}
+                                -- (315:0.5) node[p]{}
                 (15:1.3) node[p]{}

### 2014-09-09 paper/main.tex
paper: add triconnectedness proof

@@
 \end{proof}
-We believe a minimal KS graph is also triconnected.
-However, we did not find a proof.
+
+We can go one step further.
+\begin{prop}
+    A minimal Kochen-Specker graph is edge-triconnected.
+\end{prop}
+\begin{proof}
+    Given a minimal Kochen-Specker graph~$G$.
+    Suppose it is not edge-triconnected.
+    Then it splits into two graphs~$A$ and~$B$
+    together with verteces~$a_1, a_2 \in A$
+    and~$b_1,b_2 \in B$
+    such that~$(a_1, b_1)$ and~$(a_2,b_2)$
+    are the only edges between~$A$ and~$B$.
+    Note that~$A$ and~$B$ must be $010$-colorable,
+    for otherwise~$G$ would not be a minimal Kochen-Specker graph.
+    \begin{enumerate}
+    \item Suppose~$a_1 = a_2$ and~$b_1 = b_2$.
+                Then~$G$ is biconnected.  Contradiction with
+                    Proposition~\ref{prop:ks-biconn}.
+    \item Suppose~$a_1 \neq a_2$ and~$b_1 = b_2$.
+        \label{tricase1}
+        Suppose~$b_1=b_2$ does not have a fixed color in~$B$.
+        Then any coloring of~$A$ can be extended with some coloring
+        in~$B$ to a coloring of~$G$.  Contradiction.
+        Apparently~$b_1=b_2$ has a fixed color in~$B$.
+        \begin{enumerate}
+            \item
+            Suppose~$b_1=b_2$ has fixed color~$1$ in~$B$.
+            Note~$\#B \geq 16$.
+
+            Suppose there is a coloring of~$A$
+            in which both~$a_1$ and~$a_2$ have color~$0$.
+            Then, regardless whether~$a_1$ and~$a_2$ are adjacent or not,
+            this coloring can be extended with a coloring of~$B$
+            (in which~$b_1=b_2$ must be colored~$1$)
+            to a coloring~$G$.  Contradiction.
+
+            Thus the type of~$(a_1,a_2)$ in~$A$ cannot contain~$(0,0)$.
+            Thus~$\#A \geq 16$.
+            Consequently~$\#G \geq 32$. Contradiction with minimality.
+
+            \item
+            Apparently~$b_1=b_2$ has fixed color~$0$ in~$B$.
+            Note~$\#B \geq 15$.
+
+            Suppose~$a_1$ is not adjacent to~$a_2$.
+            Then any coloring of~$A$ can be extended with a coloring
+            of~$B$ to a coloring of~$G$. Contradiction.
+
+            Apparently~$a_1$ is adjacent to~$a_2$.
+
+            The type of~$(a_1,a_2)$ in~$A$ cannot contain~$(1,0)$
+            or~$(0,1)$ for otherwise~$G$ can be colored.
+            It also cannot contain~$(1,1)$ as~$a_1$ and~$a_2$
+            are adjacent.
+            Thus both~$a_1$ and~$a_2$ have fixed color~$0$ in~$A$.
+            Hence~$\#A\geq 17$. (?) % FIXME check and ref
+            Consequently~$\#G \geq 32$. Contradiction
+            with minimality.
+        \end{enumerate}
+    \item Suppose~$a_1 = a_2$ and~$b_1 \neq b_2$.
+            This leads to a contradication in the same way
+            as in case~\ref{tricase1}.
+    \item Apparently~$a_1 \neq a_2$ and~$b_1 \neq b_2$.
+        The type of~$(a_1,a_2)$ in~$A$ cannot contain~$(0,0)$,
+        for otherwise~$G$ is colorable.
+        Similarly, the type of~$(b_1,b_2)$ in~$B$ cannot contain~$(0,0)$.
+        Thus both~$\#A \geq 16$ and~$\#B \geq 16$.
+        Hence~$\#G \geq 32$. Contradiction with minimality. \qedhere
+    \end{enumerate}
+
+\end{proof}
 

### 2014-09-10 paper/main.tex
paper: *not* biconnected
paper: edge-{tri,bi}connected

@@
 \begin{prop}\label{prop:ks-biconn}
-A minimal Kochen-Specker graph is biconnected,
+A minimal Kochen-Specker graph is edge-biconnected,
 that is: removing any single edge leaves the graph connected.
@@
 We are ready to prove that a minimal Kochen-Specker graph
-is biconnected.
+is edge-biconnected.
 \begin{proof}[Proof of Proposition~\ref{prop:ks-biconn}]
-Suppose we are given a connected but not biconnected graph~$G$.
+Suppose we are given a connected but not edge-biconnected graph~$G$.
 Then we can decompose
-this graph into its biconnected components.  A biconnected components
-is a maximal subgraph that is biconnected.
-We can consider the graph of biconnected components~$B(G)$:
-two biconnected components are adjacent if there is an edge
+this graph into its edge-biconnected components.  A edge-biconnected components
+is a maximal subgraph that is edge-biconnected.
+We can consider the graph of edge-biconnected components~$B(G)$:
+two edge-biconnected components are adjacent if there is an edge
 between a vertex in one component and a vertex in the other.
 Note that there can at most be one edge between 
-the vertices of biconnected components.
+the vertices of edge-biconnected components.
 $B(G)$~does not contain loops: if it would
-then the union of the biconnected components in the graph,
-would be itself biconnected.  Thus~$B(G)$ is a tree.
+then the union of the edge-biconnected components in the graph,
+would be itself edge-biconnected.  Thus~$B(G)$ is a tree.
 
@@
 that~$G$ is a minimal Kochen-Specker graph.
-Now consider a leaf~$A$ in the biconnected component graph~$B(G)$.
-That is, $A$ is a biconnected component that is connected to only
-one other biconnected component.
+Now consider a leaf~$A$ in the edge-biconnected component graph~$B(G)$.
+That is, $A$ is a edge-biconnected component that is connected to only
+one other edge-biconnected component.
 Let~$B$ be the remainder of the graph.
@@
 Using the reasoning in the proof of Lemma~\ref{lem:biconn},
-we see~$A'$ is a Kochen-Specker graph that is not biconnected.
+we see~$A'$ is a Kochen-Specker graph that is not edge-biconnected.
 Define~$B'$ similarly. Observe that~$A'$ and~$B'$ must have the same
@@
     A minimal Kochen-Specker graph is edge-triconnected.
+    That is: removing any two edges keeps the graph connected.
 \end{prop}
@@
     \item Suppose~$a_1 = a_2$ and~$b_1 = b_2$.
-                Then~$G$ is biconnected.  Contradiction with
+                Then~$G$ is not edge-biconnected.  Contradiction with
                     Proposition~\ref{prop:ks-biconn}.
@@
 This justifies, at this stage, not checking for connectedness.
-Similarly, we believe there are very few connected but not biconnected graphs.
+Similarly, we believe there are very few connected but not
+edge-biconnected graphs.
 

### 2014-09-11 paper/main.tex
paper: add non-edge-triconnected graph figure
Improved connectivity proofs

@@
 \usepackage{graphicx}
+\usepackage{mdwlist}
 \usepackage{amsmath}
 \usepackage{wrapfig}
+\usepackage{picins}
 \usepackage{amsthm}
@@
 \newtheorem{dfn}[main]{Definition}
+\newtheorem{exa}[main]{Example}
 \newtheorem*{spin}{SPIN Axiom \cite{ck09}}
@@
 Unfortunately, these graphs are not small.
+
 \begin{comp}\label{comp:bic1}
-Let~$F^{3,1}_n$ denote the number of connected graphs
-with minimal vertex-order~$3$
-on~$n$ vertices
-with a vertex with fixed color~$1$.
-Then:
-\begin{center}
-\begin{tabular}{l|lll}
-    $n$ & $\leq 13$
-        & $14$
-        & $15$ \\
-    \hline
-    $F^{3,1}_n$ & $0$
-        & $4$
-        & $59$
-\end{tabular}
-\end{center}
-Using the methods of Section~\ref{sec:emb},
-we have determined these graphs unembeddable.
+    There are no embeddable graphs with fixed color~$1$
+    on less than 17 vertices.
 \end{comp}
-We need one final computational result.
-\begin{comp}\label{comp:bic2}
-There are no connected graphs
-on less than~$15$ vertices
-with minimal vertex-order~$2$
-with a vertex with fixed color~$0$. 
-\end{comp}
+
 We are ready to prove that a minimal Kochen-Specker graph
@@
 \begin{proof}[Proof of Proposition~\ref{prop:ks-biconn}]
-Suppose we are given a connected but not edge-biconnected graph~$G$.
-Then we can decompose
-this graph into its edge-biconnected components.  A edge-biconnected components
-is a maximal subgraph that is edge-biconnected.
-We can consider the graph of edge-biconnected components~$B(G)$:
-two edge-biconnected components are adjacent if there is an edge
-between a vertex in one component and a vertex in the other.
-Note that there can at most be one edge between 
-the vertices of edge-biconnected components.
-$B(G)$~does not contain loops: if it would
-then the union of the edge-biconnected components in the graph,
-would be itself edge-biconnected.  Thus~$B(G)$ is a tree.
-
-Furthermore suppose, reasoning towards contradication,
-that~$G$ is a minimal Kochen-Specker graph.
-Now consider a leaf~$A$ in the edge-biconnected component graph~$B(G)$.
-That is, $A$ is a edge-biconnected component that is connected to only
-one other edge-biconnected component.
-Let~$B$ be the remainder of the graph.
-There is exactly one pair~$(a,b)$ with~$a$ in~$A$ and~$b$ in~$B$
-and~$a$ adjacent to~$b$.
-
-Note that~$A$ and~$B$ are~$010$-colorable.
-Suppose that~$a$ does not have fixed color~$1$ in~$A$.
-Then there is a~$010$-coloring of~$A$
-which assigns~$0$ to~$a$.
-But then~$G$ is~$010$-colorable, quod non.
-Thus~$A$ is an embeddable graph with a vertex with fixed color~$1$.
-Similarly~$B$ is an embeddable graph with a vertex with fixed color~$1$.
-
-Let~$A'$ denote the graph containing two copies of~$A$
-connected at the copies of the vertex with fixed color~$1$.
-Using the reasoning in the proof of Lemma~\ref{lem:biconn},
-we see~$A'$ is a Kochen-Specker graph that is not edge-biconnected.
-Define~$B'$ similarly. Observe that~$A'$ and~$B'$ must have the same
-number of vertices as~$G$ by the minimality of~$G$.
-
-The graph~$A'$ has minimal vertex-order~$3$.
-Thus every vertex of~$A$ has minimal order~$3$ except for~$a$,
-which has minimal order~$2$.
-Suppose~$a$ has minimal order~$3$ in~$A$.
-Then by Computation~\ref{comp:bic1},
-the graph~$A$ has at least~$16$ vertices.
-Thus~$A'$ has at least~$32$ vertices, quod non.
-
-Thus~$a$ must have order~$2$ in~$A$.
-Let~$c,d \in A$ be the vertices connected to~$a$ in~$A$.
-For~$a$ to have fixed color~$1$ in~$A$,
-the vertices~$c$ and~$d$ must have fixed color~$0$ in~$A$.
-By Computation~\ref{comp:bic2},
-the graph~$A$ has at least~$15+1$ vertices.
-Thus~$A'$ has at least~$32$ vertices. Contradiction.
+Given a minimal Kochen-Specker graph~$G$.
+\parpic(0cm,1cm)[r]{%
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+            \draw [rotate=90] (0,-0.8) ellipse (0.6 and 0.5) ;
+            \draw [rotate=90] (0,0.8) circle (0.6 and 0.5) ;
+            \draw 
+                  (-0.7, 0.1) node[left] {$a$}
+                  (0.7, 0.1) node[right] {$b$}
+                  (-0.7, 0.1) node[p] {} -- (0.7, 0.1) node[p] {}
+                  (-0.8, -0.6) node[below] {$A$}
+                  (0.8, -0.6) node[below] {$B$}
+                ;
+        \end{tikzpicture}}
+Recall it must be connected.  Suppose it is not edge-biconnected.
+Then there must be an edge~$(a,b)$ in~$G$,
+which removal disconnects~$G$.
+Thus~$G$ decomposes into two connected graphs~$A$ and~$B$
+such that~$a\in A$, $b \in B$ and~$(a,b)$ is the only edge
+between~$A$ and~$B$.
+Clearly~$A$ and~$B$ are embeddable.
+
+Note that~$A$ must be~$010$-colorable,
+for if it were not~$010$-colorable,
+then~$A$ is a Kochen-Specker graph,
+in contradiction with~$G$'s minimality.
+Similarly~$B$ is~$010$-colorable.
+Suppose there is a~$010$-coloring of~$A$ in which~$a$ is colored~$0$.
+Then we can extend this coloring with any $010$-coloring of~$B$
+to a~$010$-coloring of~$G$, which is absurd.
+Thus~$a$ must have fixed color~$1$ in~$A$.
+Similarly~$b$ must have fixed color~$1$ in~$B$.
+Thus by Computation~\ref{comp:bic1},
+we have~$\#A \geq 17$ and~$\#B \geq 17$.
+Consequently~$\#G \geq 34$.
+Contradiction with~$G$'s minimality.
 \end{proof}
@@
 We can go one step further.
-\begin{prop}
+
+\begin{prop}\label{prop:ks-triconn}
     A minimal Kochen-Specker graph is edge-triconnected.
@@
 \end{prop}
-\begin{proof}
+
+Again, we need some preparation for the proof.  First, we generalize
+the notion of fixed color.
+
+\begin{dfn}
+   Given a graph~$G$ together with selected vertices~$v_1, \ldots, v_n \in G$.
+   Let~$C(G)$ denote the set of $010$-colorings of~$G$.
+   The \keyword{type of~$(v_1,\ldots,v_n)$ (in~$G$)}
+   is the set of all possible ways it can be colored;
+   that is: $t_G(v_1, \ldots, v_n) := \{ (c(v_1), \ldots, c(v_n));
+                \ c \in C(G) \}.$ 
+    A type of~$n$ vertices is
+    called an~\keyword{$n$-type}.
+\end{dfn}
+
+Note that~$v$ has the~$1$-type~$(1)$ in~$G$ if and only if
+it has fixed color~$1$ in~$G$.
+Just like points with fixed color,
+we are interested in types,
+because they do not occur in small graphs.
+
+\begin{comp}\label{comp:types}
+    We have enumerated all embeddable graphs of less than~$17$ vertices
+    and determined a lower bound at which a particular $1$- or $2$-type occurs.
+    \begin{center}
+    \begin{tabular}{ll}
+        $1/2$-type & $\#G$ \\ \hline
+        $\{(0,0), (1,0), (0,1)\}$ &
+            non-trivially $\geq 10$\\ 
+        $\{(0,0), (1,0), (1,1)\}$ & $\geq 10$ \\
+        $\{(0,0), (0,1), (1,1)\}$ & $\geq 10$ \\
+        $\{(0,0), (0,1))\}$ & $\geq 15$ \\
+        $\{(0,0), (1,0))\}$ & $\geq 15$ \\
+        $\{(0)\}$ & $\geq 15$ \\
+        $\{(0,1),(1,0)\}$ & $\geq 16$ \\
+        other & $\geq 17$
+    \end{tabular}
+    \end{center}
+    The type~$\{(0,0),(1,0),(0,1)\}$ occurs in the embeddable two-vertex
+    graph.  However, since
+    \begin{tikzpicture}[thick,scale=1.4,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=5pt}]
+            \draw
+            (0,0) node[p]{} -- (0.3,0) node[p]{} ;
+    \end{tikzpicture}
+    This occurance is called trivial, for the two vertices are adjacent.
+\end{comp}
+
+\begin{proof}[Proof of Proposition~\ref{prop:ks-triconn}]
+
     Given a minimal Kochen-Specker graph~$G$.
+
+\parpic(0cm,1cm)[r]{%
+\begin{tikzpicture}[thick,scale=1.4,
+p/.style={circle, draw, fill=white,
+                inner sep=0pt, minimum width=5pt}]
+    \draw [rotate=90] (0,-0.8) ellipse (0.8 and 0.5) ;
+    \draw [rotate=90] (0,0.8) circle (0.8 and 0.5) ;
+    \draw 
+          (-0.7, 0.3) node[left] {$a_1$}
+          (-0.7, -0.3) node[left] {$a_2$}
+          (0.7, 0.3) node[right] {$b_1$}
+          (0.7, -0.3) node[right] {$b_2$}
+          (-0.7, -0.3) node[p] {} -- (0.7, -0.3) node[p] {}
+          (-0.7, 0.3) node[p] {} -- (0.7, 0.3) node[p] {}
+          (-0.8, -0.8) node[below] {$A$}
+          (0.8, -0.8) node[below] {$B$}
+        ;
+\end{tikzpicture}}
+
     Suppose it is not edge-triconnected.
@@
     for otherwise~$G$ would not be a minimal Kochen-Specker graph.
+
     \begin{enumerate}
@@
             Suppose~$b_1=b_2$ has fixed color~$1$ in~$B$.
-            Note~$\#B \geq 16$.
+            Note~$\#B \geq 17$ by Computation~\ref{comp:types}.
 
@@
             Thus the type of~$(a_1,a_2)$ in~$A$ cannot contain~$(0,0)$.
-            Thus~$\#A \geq 16$.
-            Consequently~$\#G \geq 32$. Contradiction with minimality.
+            Thus, by Computation~\ref{comp:types}, $\#A \geq 17$.
+            Consequently~$\#G \geq 34$. Contradiction with minimality.
 
@@
             Apparently~$b_1=b_2$ has fixed color~$0$ in~$B$.
-            Note~$\#B \geq 15$.
+            Hence, by Computation~\ref{comp:types}, $\#B \geq 15$.
 
@@
             Thus both~$a_1$ and~$a_2$ have fixed color~$0$ in~$A$.
-            Hence~$\#A\geq 17$. (?) % FIXME check and ref
+            Hence~$\#A\geq 17$ by Computation~\ref{comp:types}.
             Consequently~$\#G \geq 32$. Contradiction
@@
         Similarly, the type of~$(b_1,b_2)$ in~$B$ cannot contain~$(0,0)$.
-        Thus both~$\#A \geq 16$ and~$\#B \geq 16$.
-        Hence~$\#G \geq 32$. Contradiction with minimality. \qedhere
+        Thus both~$\#A \geq 17$ and~$\#B \geq 17$.
+        Hence~$\#G \geq 34$. Contradiction with minimality. \qedhere
     \end{enumerate}
-
 \end{proof}

### 2014-09-11 paper/main.tex
paper: ammend a bit

@@
     The type~$\{(0,0),(1,0),(0,1)\}$ occurs in the embeddable two-vertex
-    graph.  However, since
+    graph
     \begin{tikzpicture}[thick,scale=1.4,
@@
             (0,0) node[p]{} -- (0.3,0) node[p]{} ;
-    \end{tikzpicture}
-    This occurance is called trivial, for the two vertices are adjacent.
+    \end{tikzpicture}.
+    Because the two vertices are adjacent, this occurance of
+    the type is called trivial.
 \end{comp}

### 2014-09-12 paper/main.tex
paper: suggesties van Sander

@@
 (or equivalently, a finite subset of the northern
-hemisphere without equator).
+hemisphere without equator\footnote{%
+    A subset of the projective plane can be identified with
+    a subset of the closed northern hemisphere.
+    For a finite subset we can always rotate in such a way
+    that no points lie on the equator.}).
 Define its \keyword{orthogonality graph}~$G(S)$ as follows.
@@
 \begin{dfn}
-A graph is~\keyword{010-colorable}
-if there is a~$\{0,1\}$-coloring, such that
+A graph is called~\keyword{010-colorable}
+if there is a~$\{0,1\}$-coloring of the vertices,
+such that
 \begin{enumerate}
@@
 vertices.\cite{sf-sloane}
-We can restrict ourselves to connected graphs.
+Next, we show we can restrict ourselves to connected graphs.
 \begin{prop}[\cite{aow11}]\label{prop:ks-conn}
@@
     A minimal Kochen-Specker graph has minimal vertex-order three.
+    That is: every vertex is adjacent to at least three other vertices.
 \end{prop}
 \begin{proof}
-    Given a Kochen-Specker graph~$G$.
+    Given a minimal Kochen-Specker graph~$G$.
     Suppose~$v$ is a vertex with order less than or equal~$2$.
@@
     Therefore~$G'$ is a smaller
-    Kochen-Specker graph, which contradicts minimality.
+    Kochen-Specker graph, which contradicts the minimality of~$G$.
 \end{proof}
@@
    Let~$C(G)$ denote the set of $010$-colorings of~$G$.
-   The \keyword{type of~$(v_1,\ldots,v_n)$ (in~$G$)}
-   is the set of all possible ways it can be colored;
-   that is: $t_G(v_1, \ldots, v_n) := \{ (c(v_1), \ldots, c(v_n));
+   The \keyword{type~$t$ of~$(v_1,\ldots,v_n)$ (in~$G$)}
+   is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be colored;
+   that is: $t = \{ (c(v_1), \ldots, c(v_n));
                 \ c \in C(G) \}.$ 
@@
 
-Note that~$v$ has the~$1$-type~$(1)$ in~$G$ if and only if
+\begin{exa}
+\begin{itemize}
+\item The triangle has~$3$-type~$\{(1,0,0), (0,1,0), (0,0,1)\}$.
+\item Every vertex in a Kochen-Specker graph has type~$\emptyset$.
+\item
+    A vertex~$v$ has the~$1$-type~$\{(1)\}$ in~$G$ if and only if
         it has fixed color~$1$ in~$G$.
-Just like points with fixed color,
+\end{itemize}
+\end{exa}
+
+Just as verteces with fixed color are rare,
 we are interested in types,
-because they do not occur in small graphs.
+because most types do not occur in small graphs.
 
@@
     We have enumerated all embeddable graphs of less than~$17$ vertices
-    and determined a lower bound at which a particular $1$- or $2$-type occurs.
+    and determined a lower bound at which a particular $1$- or $2$-type occurs,
+    omitting the trivial types~$\{(0),(1)\}$
+    and~$\{(0,0),(0,1),(1,0),(1,1)\}$.
     \begin{center}
@@
 \section{Embeddability}
-Our computation has yielded a few hundred non-010-colorable graphs.
+Our computation has yielded over nine-thousand non-010-colorable graphs.
 If we show one of them is embeddable, we have found a new KS system.
@@
 It is orthogonal to both~$a$ and~$v$.
-$a$ and~$v$ are not collinear,
-hence~$p_1$ must be collinear to~$v \times a$,
+Since $a$ and~$v$ are not collinear,
+$p_1$ must be collinear to~$v \times a$,
 the cross-product of~$v$ and~$a$.
@@
 \end{equation*}
-And therefore
+Consequently
 \begin{align*}
-v_1w_2 \left<v,w\right> & =
-v_1w_2 (v_1w_1 + v_2w_2) \\
+    v_1w_2 (v_1w_1 + v_2w_2)
 & = v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2 \\
 & = (v_1^2 + v_2^2)w_1^2 + (v_1^2 + v_2^2)w_2^2 \\
-& = w_1^2 + w_2^2 = 1.
+& = w_1^2 + w_2^2 \\
+&= 1.
 \end{align*}
@@
 we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
-Recall~$|v_1|, |w_2| \leq 1$.
-Thus: $|v_1w_2\left<v,w\right>| < 1$.
-Contradiction.
+Now we find the contradiction:
+\begin{equation*}
+1 > |v_1w_2\left<v,w\right>| = |v_1w_2(v_1w_1+v_2w_2)| = 1. \qedhere
+\end{equation*}
 \end{proof}
@@
 We have used the redlog\cite{redlog} package of the reduce algebra
-system, which implements a variant of Tarski's quantifier elimination.
-In Appendix~\ref{adx:unemb-10-2} the Reader can find
-the reduce script generated mechanically for the graph
-in Figure~\ref{fig:unemb-10-2}.
+system, which implements a variant of Tarski's quantifier
+elimination.\footnote{%
+    The reader can find the reduce script generated mechanically
+    for the graph in Figure~\ref{fig:unemb-10-2} here:
+    \url{http://kochen-specker.info/smallGraphs/49743f49514769444f.html}.  }
 
@@
 we could not decide embeddability.
-By hand, we determined these graphs to be embeddable.
-We adapted the algorithm, as to guess
+With some guessing,
+we determined embeddings for these graphs by hand.
+Once we knew the troublesome graphs were embeddable,
+we adapted the algorithm, as to guess
 for some assignments the position of one of the vectors.
@@
 
-With this method, we have decided in an a day the embeddability
-of every squarefree graph with minimal vertex order three of 14
-vertices of less except for one.\footnote{
+With this method, we have decided in a day the embeddability
+of every squarefree graph with minimal vertex order three of
+less than 15, except for one.\footnote{
 A list of all squarefree graphs with minimal vertex order three of
-14 vertices of less together with their embeddability can
+less than 15 vertices together with their embeddability can
 be found here:
-\url{http://kochen-specker.info/smallGraphs/}
+\url{http://kochen-specker.info/smallGraphs/}.
+The graph for which we could not determine
+embeddability can be found
+here: \url{http://kochen-specker.info/smallGraphs/4d4b3f4b3f603f47414641654953625f3f.html}.
 }
@@
 to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
-for drawing the first two figures in this paper.
-
-\clearpage
-\appendix
-\section{Example reduce script}\label{adx:unemb-10-2}
-The following is (a part of) the reduce script mechanically generated
-to prove the non-embeddability of the graph of Figure~\ref{fig:unemb-10-2}.
-The algorithm choose a different assignment, than we did in the
-proof of Proposition~\ref{prop:unemb-10-2}.
-As free points it picked, in order,~$z$, $x$, $v$ and~$p_3$.
-The point~$w$ is assigned~$p_3 \times z$.
-\begin{verbatim}
-load_package redlog;
-rlset R;
-
-procedure d(x,y);
-    (first x) * (first y) +
-    (second x) * (second y) +
-    (third x) * (third y);
-
-procedure k(x,y);
-    {(second x)*(third y) - (third x)*(second y),
-     (third x)*(first y) - (first x)*(third y),
-     (first x)*(second y) - (second x)*(first y)};
-
-v0c1 := 1; v0c2 := 0; v0c3 := 0;
-v1c1 := 0; v1c2 := 1; v1c3 := 0;
-
-v0 := {v0c1, v0c2, v0c3}; 
-v1 := {v1c1, v1c2, v1c3}; 
-v2 := {v2c1, v2c2, v2c3}; 
-v3 := {v3c1, v3c2, v3c3}; 
-v2c1 := 0;
-neq0 := k(v0,k(v3,v1)); 
-\end{verbatim}
-\begin{center} \emph{(snip)} \end{center}
-\begin{verbatim}
-neq29 := k(k(k(k(v3,v1),v1),v2),k(k(v3,v0),v3)); 
-phi := 
-       (first neq0 neq 0 or
-        second neq0 neq 0 or
-        third neq0 neq 0) and 
-\end{verbatim}
-\begin{center} \emph{(snip)} \end{center}
-\begin{verbatim}
-       (first neq29 neq 0 or
-        second neq29 neq 0 or
-        third neq29 neq 0) and 
-       d(v2,v0) = 0 and 
-       d(k(k(v3,v0),v3),k(k(k(k(v3,v1),v1),v2),v2)) = 0 and 
-        true;
-rlqe ex(v3c3,
-     ex(v3c2,
-     ex(v3c1,
-     ex(v2c3,
-     ex(v2c2,phi)))));
-\end{verbatim}
+for drawing figures for the draft version of this paper.
 

### 2014-09-12 paper/main.tex
paper: more suggestions of Sander

@@
 
-Consider the following experiment. Measure the SPIN of a spin 1
-particle along a direction. Or more concretely: shoot a deuterium
-atom through a certain fixed inhomogeneous magnetic field. Depending
+Consider the following experiment. Measure the spin of a spin 1
+particle along a direction. For example, shoot a deuterium
+atom through a certain fixed inhomogeneous magnetic field,
+such as that in the Stern-Gerlach experiment. Depending
 on the direction it was shot through the field, the particle will
@@
 Its probabilistic prediction has been thoroughly tested.
-One wonders: is there a deterministic theory predicting the
+One wonders: is there a
+\emph{deterministic} theory predicting the
 outcome of this experiment?
 
-Kochen and Specker have shown that such a deterministic theory must
-be odd: it cannot satisfy the plausible SPIN axiom, that is:
+Kochen and Specker have shown that such a non-contextual deterministic
+theory must be odd: it cannot satisfy the plausible SPIN
+axiom, that is:
 \begin{spin}

### 2014-09-13 paper/main.tex
paper: replace ks31 by Sander's tikz
paper: improve other pictures

@@
 \usepackage{paralist}
+\usepackage{etoolbox}
 \usepackage{graphicx}
@@
 
-\begin{wrapfigure}{r}{0.30\textwidth}
-\begin{center}
-\includegraphics[width=40mm]{../graphs/ks31.pdf}
-\end{center}
-\caption{Conway's 31 vector Kochen-Specker system
-\label{fig:conway31}}
-\end{wrapfigure}
+\piccaption{Conway's 31 vector Kochen-Specker system\label{fig:conway31}}
+\parpic[r]{%
+%  Based on  <http://www.latex-community.org/viewtopic.php?f=4&t=2111>
+\begin{tikzpicture}[scale=0.8,
+        style={inner sep=0pt,%
+              outer sep=2pt,%
+              mark coordinate/.style={%
+                inner sep=0pt,outer sep=0pt,minimum size=4pt,
+                fill=black,circle}%
+            }]
+\def\R{2.5} % sphere radius
+\def\St{1.41421356237}
+\def\ST{1.73205080757}
+\filldraw[ball color=white] (0,0) circle (\R);
+\coordinate[mark coordinate] (N) at (0,0);
+\coordinate[mark coordinate] (N) at (0,-\R/\St);
+\coordinate[mark coordinate] (N) at (0,-\R*\St);
+\coordinate[mark coordinate] (N) at (0,\R*\St/4);
+\coordinate[mark coordinate] (N) at (0,\R*\St/2);
+\coordinate[mark coordinate] (N) at (0,3*\R*\St/4);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/8,3*\R*\St/8);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/8,5*\R*\St/8);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/4,0);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/4,-\R*\St/4);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/4,-\R*\St/2);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/4,-3*\R*\St/4);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/4,\R*\St/4);
+\coordinate[mark coordinate] (N) at (-\R*\St*\ST/4,\R*\St/2);
+\coordinate[mark coordinate] (N) at (-3*\R*\St*\ST/8,\R*\St/8);
+\coordinate[mark coordinate] (N) at (-3*\R*\St*\ST/8,-\R*\St/8);
+\coordinate[mark coordinate] (N) at (-3*\R*\St*\ST/8,-3*\R*\St/8);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/8,3*\R*\St/8);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/8,5*\R*\St/8);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/8,-\R*\St/8);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/8,-3*\R*\St/8);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/8,-5*\R*\St/8);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/4,0);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/4,\R*\St/4);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/4,\R*\St/2);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/4,-\R*\St/4);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/4,-\R*\St/2);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/4,-3*\R*\St/4);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/2,-\R/\St);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/2,0);
+\coordinate[mark coordinate] (N) at (\R*\St*\ST/2,\R/\St);
+\draw (0,0) -- (0,-\R*\St) -- (-\R*\St*\ST/2,-\R/\St) --  (-\R*\St*\ST/2,\R/\St)
+    --  (0,\R*\St) -- (\R*\St*\ST/2,\R/\St) 
+    -- (\R*\St*\ST/2,-\R/\St) -- (0,- \R*\St);
+\draw  (-\R*\St*\ST/2,\R/\St) -- (0,0) -- (\R*\St*\ST/2,\R/\St);
+\draw (-\R*\St*\ST/4, -3*\R*\St/4) -- (-\R*\St*\ST/4,\R*\St/4 )
+    -- (\R*\St*\ST/4, 3*\R*\St/4) ;
+\draw (\R*\St*\ST/4, -3*\R*\St/4) -- (\R*\St*\ST/4,\R*\St/4 )
+    -- (-\R*\St*\ST/4, 3*\R*\St/4) ;
+\draw  (-\R*\St*\ST/2,{\R/\St - \R*\St/4}) -- (0,- \R*\St/4)
+    -- (\R*\St*\ST/2,{\R/\St - \R*\St/4});
+\draw  (-\R*\St*\ST/2,{\R/\St - \R*\St/2}) -- (0,- \R*\St/2)
+    -- (\R*\St*\ST/2,{\R/\St - \R*\St/2});
+\draw  (-\R*\St*\ST/2,{\R/\St - 3*\R*\St/4}) -- (0,- 3*\R*\St/4)
+    -- (\R*\St*\ST/2,{\R/\St - 3*\R*\St/4});
+\draw(-\R*\St*\ST/8,-0.25*7*\R/\St) -- (-\R*\St*\ST/8,{-0.25*7*\R/\St + \R*\St})
+    -- (3*\R*\St*\ST/8,5*0.25*\R/\St);
+\draw(\R*\St*\ST/8,-0.25*7*\R/\St) -- (\R*\St*\ST/8,{-0.25*7*\R/\St + \R*\St})
+    -- (-3*\R*\St*\ST/8,5*0.25*\R/\St);
+\draw (-3*\R*\St*\ST/8,-5*0.25*\R/\St)
+    -- (-3*\R*\St*\ST/8,{-5*0.25*\R/\St + \R*\St})
+    -- (\R*\St*\ST/8,0.25*7*\R/\St);
+\draw (3*\R*\St*\ST/8,-5*0.25*\R/\St)
+    -- (3*\R*\St*\ST/8,{-5*0.25*\R/\St + \R*\St})
+    -- (-\R*\St*\ST/8,0.25*7*\R/\St);
+\undef\R
+\undef\ST
+\undef\St
+\end{tikzpicture}}
 
@@
 we describe our practical embeddability test.
-
-\begin{wrapfigure}{r}{0.25\textwidth}
-\begin{center}
-\begin{tikzpicture}[scale=1.9,thick,
-    partial ellipse/.style args={#1:#2:#3}{
-            insert path={+ (#1:#3) arc (#1:#2:#3)}
-        }]
-    \draw (0,0) [partial ellipse=360:180:1 and 0.3] ;
-    \draw[gray,dashed] (0,0) [partial ellipse=0:180:1 and 0.3] ;
-    \draw (0,0) [rotate=315,partial ellipse=360:180:1 and 0.3] ;
-    \draw[gray,dashed] (0,0) [rotate=315,partial ellipse=0:180:1 and 0.3] ;
-    \draw (0,0) circle (1) ;
-    \node[circle,inner sep=1pt,fill] at (90:1) {};
-    \node[above] at (90:1) {$v$};
-    \node[circle,inner sep=1pt,fill] at (45:1) {};
-    \node[right] at (45:1) {$w$};
-\end{tikzpicture}
-\end{center}
-\caption{squarefreeness of an embeddable graph
-\label{fig:crossproduct}}
-\end{wrapfigure}
 
@@
 \end{dfn}
+
 \begin{dfn}
@@
 \end{fact}
+
 To prove there is no Kochen-Specker system on~$17$ points,
@@
 \end{prop}
+
+\parpic[r]{%
+\begin{tikzpicture}[scale=1.0,thick,
+    partial ellipse/.style args={#1:#2:#3}{
+            insert path={+ (#1:#3) arc (#1:#2:#3)}
+        }]
+    \draw (0,0) [partial ellipse=360:180:1 and 0.3] ;
+    \draw[gray,dashed] (0,0) [partial ellipse=0:180:1 and 0.3] ;
+    \draw (0,0) [rotate=315,partial ellipse=360:180:1 and 0.3] ;
+    \draw[gray,dashed] (0,0) [rotate=315,partial ellipse=0:180:1 and 0.3] ;
+    \draw (0,0) circle (1) ;
+    \node[circle,inner sep=1pt,fill] at (90:1) {};
+    \node[above] at (90:1) {$v$};
+    \node[circle,inner sep=1pt,fill] at (45:1) {};
+    \node[right] at (45:1) {$w$};
+\end{tikzpicture}}
+
 \begin{proof}
     Given two non antipodal points~$v\neq w$.
-    See Figure~\ref{fig:crossproduct}.
+    See the figure on the right.
     Consider the points orthogonal to~$v$.
@@
 \end{proof}
+
 The squarefreeness is a considerable restriction.  There are
@@
 
-\section{Embeddability}
+\section{Embeddability}\label{sec:emb}
 Our computation has yielded over nine-thousand non-010-colorable graphs.
@@
 
-\label{sec:emb}
-\begin{wrapfigure}{r}{0.30\textwidth}
-\begin{center}
-    \begin{tikzpicture}[thick,scale=1.4,
-        p/.style={circle, draw, fill=white,
+\piccaption{One of the two minimal non-embeddable graphs \label{fig:unemb-10-2}}
+\parpic[r]{%
+\begin{tikzpicture}[thick,scale=1.4,%
+        p/.style={circle, draw, fill=white,%
                         inner sep=0pt, minimum width=5pt}]
@@
                 (330-30:1.3) node[p]{}
-
                 (90:0.7) node[above]{$w$}
@@
                 (330+30:1.3) node[right]{$p_5$}
-                (330-30:1.3) node[below]{$a$}
-                ;
-        \end{tikzpicture}
-\end{center}
-\caption{One of the two minimal
-        non-embeddable graphs
-\label{fig:unemb-10-2}}
-\end{wrapfigure}
+                (330-30:1.3) node[below]{$a$} ;
+\end{tikzpicture}}
 
@@
     \begin{center}
-    \begin{tikzpicture}[thick,scale=1.4,
+    \begin{tikzpicture}[thick,scale=1.0,
         p/.style={circle, draw, fill=white,
@@
         \qquad
-    \begin{tikzpicture}[thick,scale=1.4,
+    \begin{tikzpicture}[thick,scale=1.0,
         p/.style={circle, draw, fill=white,

### 2014-09-13 code/comp6/010cc/010cc.cpp
Add comp6

@@ -1,0 +1,214 @@
+#include <functional>
+#include <iostream>
+#include <fstream>
+#include <array>
+#include <set>
+#include <stack>
+#include <stdexcept>
+#include <cassert>
+#include <vector>
+
+typedef uint8_t node_t;
+typedef std::array<node_t, 3> triangle_t;
+
+class base_graph
+{
+public:
+    void find_010coloring();
+};
+
+std::array<node_t, 3> triangle(node_t a, node_t b, node_t c)
+{
+    std::array<node_t, 3> ret = {{a, b, c}};
+    if(ret[0] > ret[1])
+        std::swap(ret[1], ret[0]);
+    if(ret[1] > ret[2])
+        std::swap(ret[1], ret[2]);
+    return ret;
+}
+
+template<std::size_t N>
+class graph : public base_graph
+{
+    static_assert(0 < N && N < 62, "Only 0 < N < 62 is supported");
+
+private:
+    std::array<std::array<bool, N>, N> adj;
+    std::array<std::vector<node_t>, N> neigh;
+    std::array<std::vector<triangle_t>, N> triangles_of;
+
+public:
+    graph (std::string graph6)
+                : adj(), neigh(), triangles_of()
+    {
+        // First, sanity checks.
+        if(graph6.empty())
+            throw std::runtime_error("graph6 should not be empty");
+        unsigned int runtime_N =  graph6[0] - 63;
+        if(runtime_N != N)
+            throw std::runtime_error("graph size unsupported");
+
+        // Now, load the adjacency matrix and neighbourhood lists
+        int x = 0, y = 1;
+        for(int i = 1; i < graph6.length(); i++) {
+            for(int j = 0; j < 6; j++) {
+                bool value = (((graph6[i] - 63) >> (5 - j)) & 1) == 1;
+                this->adj[x][y] = value;
+                this->adj[y][x] = value;
+                if(value) {
+                    this->neigh[x].push_back(y);
+                    this->neigh[y].push_back(x);
+                }
+                if(++x == y) {
+                    y++;
+                    if (y == N)
+                        break;
+                    x = 0;
+                }
+            }
+            if (y == N)
+                break;
+        }
+
+        // Now, find the triangles
+        for(node_t x = 0; x < N; x++) {
+            for(node_t y : this->neigh[x]) {
+                if (y < x) continue;
+                for(node_t z: this->neigh[y]) {
+                    if (z < y) continue;
+                    if(!this->adj[z][x]) continue;
[…]

### 2014-09-13 code/comp6/client.py

@@ -1,0 +1,75 @@
+import sys
+import zmq
+import json
+import time
+import random
+import threading
+import subprocess
+import multiprocessing
+
+class Client(object):
+    def worker(self):
+        while True:
+            with self.cond:
+                self.cond.wait()
+                if self.task == -1:
+                    continue
+                task = self.task
+            print 'working on task %s' % task
+            self.result = subprocess.Popen(['./run-batch', str(task)],
+                                stdout=subprocess.PIPE,
+                                stderr=subprocess.PIPE).communicate()
+            print '     done!'
+            with self.cond:
+                self.cond.notify()
+
+    def main(self):
+        self.cond = threading.Condition()
+        self.result = None
+        self.task = -1
+        self.start_time = 0
+        identifier = str(random.randint(0,10000000))
+        if len(sys.argv) > 1:
+            identifier = sys.argv[1] + '-' + identifier
+        self.worker = threading.Thread(target=self.worker)
+        self.worker.start()
+        zctx = zmq.Context()
+        s = zctx.socket(zmq.REQ)
+        s.connect('tcp://sw.w-nz.com:5558')
+        while True:
+            s.send_json(['ping', identifier, self.task])
+            s.recv_json()
+            with self.cond:
+                if self.result is not None:
+                    while True:
+                        s.send_json(['result', identifier, self.task,
+                                                    json.dumps(self.result),
+                                            time.time() - self.start_time])
+                        msg = s.recv_json()
+                        if msg[0] == 'ok':
+                            break
+                        print 'result not accepted: %s; retrying' % msg
+                        time.sleep(1)
+                    self.task = -1
+                    self.result = None
+                if self.task == -1:
+                    s.send_json(['request', identifier])
+                    msg = s.recv_json()
+                    if msg[0] == 'done':
+                        return
+                    self.task = msg[1]
+                    self.start_time = time.time()
+                    self.cond.notify()
+                self.cond.wait(60)
+
+def main():
+    Client().main()
+
+if __name__ == '__main__':
+    if len(sys.argv) > 2:
+        cpu_count = int(sys.argv[2])
+    else:
+        cpu_count = multiprocessing.cpu_count()
+    for i in xrange(cpu_count):
+        print 'starting worker', i
+        threading.Thread(target=main).start()

### 2014-09-13 code/comp6/server.py

@@ -1,0 +1,105 @@
+import zmq
+import time
+import json
+import os.path
+import kyotocabinet
+
+BITS=100000
+
+class Server(object):
+    def __init__(self):
+        self.workers = {}
+    def open_db(self):
+        path = 'db.kch'
+        self.db = kyotocabinet.DB()
+        if os.path.exists(path):
+            print 'Opening db ...'
+            if not self.db.open(path, kyotocabinet.DB.OWRITER):
+                raise RuntimeError
+            n_replaced = 0
+            for key in self.db.match_prefix('result-'):
+                data = json.loads(self.db.get(key))
+                result = json.loads(data[0])
+                if not 'graphs generated' in result[1]:
+                    n_replaced += 1
+                    self.db.set(key.replace('result-', 'todo-'), None)
+                    self.db.remove(key)
+            for key in self.db.match_prefix('working-'):
+                if time.time() - json.loads(self.db.get(
+                        'worker-'+self.db.get(key)))[1] > 200:
+                    n_replaced += 1
+                    self.db.set(key.replace('working-', 'todo-'), None)
+                    self.db.remove(key)
+            print '  reset {} results'.format(n_replaced)
+            print '    done'
+            return
+        print 'Creating db ...'
+        if not self.db.open(path, kyotocabinet.DB.OWRITER |
+                                  kyotocabinet.DB.OCREATE):
+            raise RuntimeError
+        for i in xrange(BITS):
+            self.db.set('todo-{}'.format(i), None)
+        print '     done'
+
+    def main(self):
+        self.open_db()
+        self.last_sync = 0
+        zctx = zmq.Context()
+        s = zctx.socket(zmq.REP)
+        s.bind('tcp://*:5558')
+        while True:
+            msg = s.recv_json()
+            if time.time() - self.last_sync > 100:
+                self.last_sync = time.time()
+                self.db.synchronize()
+            if not isinstance(msg, list) or not msg:
+                s.send_json(['error', 'malformed request'])
+                continue
+            if msg[0] == 'ping':
+                if (not len(msg) == 3 or not isinstance(msg[1], str)
+                                or not isinstance(msg[2], int)):
+                    s.send_json(['error', 'malformed request'])
+                    continue
+                worker_id = msg[1]
+                task = msg[2]
+                self.db.set('worker-{}'.format(worker_id),
+                            json.dumps([task, time.time()]))
+                s.send_json(['pong'])
+                continue
[…]

### 2014-09-13 code/comp6/stats.py

@@ -1,0 +1,91 @@
+import json
+import time
+import math
+import os.path
+import kyotocabinet
+from pprint import pprint
+
+def avg(l):
+    return sum(l)/len(l)
+def stddev(l):
+    a = avg(l)
+    r = 0.0
+    for x in l:
+        r += (a - x)**2
+    return math.sqrt(r / len(l))
+
+class Program(object):
+    def main(self):
+        path = 'db.kch'
+        self.db = kyotocabinet.DB()
+        if not self.db.open(path, kyotocabinet.DB.OREADER
+                                | kyotocabinet.DB.ONOLOCK):
+            raise RuntimeError
+        n_workers = 0
+        names = dict()
+        n_todo = 0
+        for key in self.db.match_prefix('todo-'):
+            n_todo += 1
+        n_stale = 0
+        n_working = 0
+        for key in self.db.match_prefix('working-'):
+            if time.time() - json.loads(self.db.get(
+                    'worker-'+self.db.get(key)))[1] > 200:
+                n_stale += 1
+            else:
+                n_working += 1
+        print 'n_stale:', n_stale
+        print 'n_working', n_working
+        print 'n_todo:', n_todo
+        for key in self.db.match_prefix('worker-'):
+            if time.time() - json.loads(self.db.get(key))[1] > 200:
+                continue
+            n_workers += 1
+            name = key.split('-',1)[1].rsplit('-',1)[0]
+            if not name in names:
+                names[name] = 0
+            names[name] += 1
+        print 'number of workers: ', n_workers
+        pprint(names)
+        total_duration = 0.0
+        total_graphs = 0
+        eff_by_worker = {}
+        N = 0
+        N_error = 0
+        candidates = []
+        for key in self.db.match_prefix('result-'):
+            data = json.loads(self.db.get(key))
+            result = json.loads(data[0])
+            if not 'graphs generated' in result[1]:
+                N_error += 1
+                continue
+            candidates.extend(filter(lambda x: x, result[0].split('\n')))
+            duration = data[2]
+            graphs_generated =  int(result[1].split(
+                            'graphs generated')[0].split('>Z')[1])
+            worker = data[1].split('-')[0]
+            if not worker in eff_by_worker:
+                eff_by_worker[worker] = []
+            eff_by_worker[worker].append(float(graphs_generated) / duration)
+            total_graphs += graphs_generated
+            N += 1
[…]

### 2014-09-13 paper/main.tex
paper: improve placing of ks31

@@
 %  Based on  <http://www.latex-community.org/viewtopic.php?f=4&t=2111>
-\begin{tikzpicture}[scale=0.8,
+\begin{tikzpicture}[scale=0.75,
         style={inner sep=0pt,%

### 2014-09-13 paper/main.tex
paper: Add links to computations

@@
 and the deciding whether such graphs are embeddable.
-
 In Section~\ref{sec:ilb},
@@
 we describe our practical embeddability test.
+
+The software and results of the various computations performed for
+this paper, can be found here\cite{GH}.
 
@@
     There are no embeddable graphs with fixed color~$1$
-    on less than 17 vertices.
+    on less than 17 vertices.%
+        \footnote{Source code at \texttt{code/comp5.py} of \cite{GH}.}
 \end{comp}
@@
     omitting the trivial types~$\{(0),(1)\}$
-    and~$\{(0,0),(0,1),(1,0),(1,1)\}$.
+    and~$\{(0,0),(0,1),(1,0),(1,1)\}$.%
+        \footnote{Source code at \texttt{code/comp5.py} of \cite{GH}.}
     \begin{center}
@@
     Let~$C_n$ denote the number of non-010 colorable squarefree
-    graphs with minimal vertex order 3 on~$n$ nodes.  Then:
+    graphs with minimal vertex order 3 on~$n$ nodes.  Then:%
+        \footnote{Source code at \texttt{code/comp6} of \cite{GH}.}
 
@@
 subgraphs, can be found here:
-\url{http://kochen-specker.info/candidates/}}
+\url{http://kochen-specker.info/candidates/}.
+The source code for this computation can be found
+at~\texttt{code/comp2.py} of~\cite{GH}.
+}
 \end{comp}

### 2014-09-15 paper/main.tex
paper: second submission to QPL
paper: laatste wijzigingen

@@
-\documentclass[adraft,copyright,creativecommons]{eptcs}
+\documentclass[submission,copyright,creativecommons]{eptcs}
 
@@
 
-\usepackage{algorithm2e}
+\usepackage{algpseudocode}
 \usepackage{breakurl}
@@
 \usepackage{tikz}
+
+% begin vertical rule patch for algorithmicx
+% http://tex.stackexchange.com/questions/144840/
+\makeatletter
+\newcommand*{\algrule}[1][\algorithmicindent]{\makebox[#1][l]{\hspace*{.5em}\vrule height .75\baselineskip depth .25\baselineskip}}%
+
+\newcount\ALG@printindent@tempcnta
+\def\ALG@printindent{%
+    \ifnum \theALG@nested>0%
+        \ifx\ALG@text\ALG@x@notext%
+            % do nothing
+            \addvspace{-3pt}%
+        \else
+            \unskip
+            % 
+            \ALG@printindent@tempcnta=1
+            \loop
+                \algrule[\csname ALG@ind@\the%
+                    \ALG@printindent@tempcnta\endcsname]%
+                \advance \ALG@printindent@tempcnta 1
+            \ifnum \ALG@printindent@tempcnta<\numexpr\theALG@nested+1\relax%
+            \repeat
+        \fi
+    \fi
+    }%
+% 
+\patchcmd{\ALG@doentity}{\noindent\hskip\ALG@tlm}%
+        {\ALG@printindent}{}{\errmessage{failed to patch}}
+\makeatother
+% end vertical line thingy
 
@@
 
-Consider the following experiment. Measure the spin of a spin 1
-particle along a direction. For example, shoot a deuterium
-atom through a certain fixed inhomogeneous magnetic field,
-such as that in the Stern-Gerlach experiment. Depending
-on the direction it was shot through the field, the particle will
-move undisturbed or deviate.
+Consider the following experiment.  Shoot a deuterium
+atom (or another neutral spin 1 particle)
+through a certain fixed inhomogeneous magnetic field,
+such as that in the Stern-Gerlach experiment.
+The particle will then move undisturbed or deviate.
+What we have done is measure the spin component%
+\footnote{%
+As we are only interested in whether the particle deviates or
+    not, we actually only consider the square of the spin component.}
+ of the particle
+along a certain direction.  This direction depends on the specifics of the
+field and the movement of the particle.
 
@@
 A graph~$G$ is called~\keyword{embeddable},
-if it occurs as a subgraph of an orthogonality graph;
-that is: if there is a finite subset~$S$ of the projective plane,
+if it occurs as a subgraph of an orthogonality graph.
+That is: if there is a finite subset~$S$ of the projective plane,
 such that~$G \leq G(S)$.
@@
 \begin{prop}\label{prop:ks-biconn}
-A minimal Kochen-Specker graph is edge-biconnected,
-that is: removing any single edge leaves the graph connected.
+A minimal Kochen-Specker graph is edge-biconnected.
+That is: removing any single edge leaves the graph connected.
 \end{prop}
@@
    The \keyword{type~$t$ of~$(v_1,\ldots,v_n)$ (in~$G$)}
-   is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be colored;
-   that is: $t = \{ (c(v_1), \ldots, c(v_n));
+   is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be colored.
+   That is: $t = \{ (c(v_1), \ldots, c(v_n));
                 \ c \in C(G) \}.$ 
@@
 
-\begin{algorithm}
-    \While{there are unassigned vertices}{
-        pick an unassigned vertex~$v$\;
-        assign~$V(v)=v$\;
-        mark~$v$ as free\;
+\begin{algorithmic}[5]
+    \While{there are unassigned vertices}
+        \State pick an unassigned vertex~$v$
+        \State assign~$V(v)=v$ 
+        \State mark~$v$ as free
         \While{there are unassigned vertices adjacent to two
-                different assigned vertices}{
-            pick such a vertex~$w$ adjacent to the assigned ~$w_1$ and~$w_2$\;
-            assign~$V(w)=V(w_1) \times V(w_2)$\;
-            mark edges~$(v,w_1)$ and~$(v,w_2)$ as accounted for\;
-        }   
-    }
-    \For{each pair of vertices~$(v_1, v_2)$}{
-        \If{$(v_1,v_2)$ is not an edge} {
-            record requirement:~``$V(v_1)$ is not collinear to $V(v_2)$''\;
-        }
-    }
-    \For{each edge~$(v_1,v_2)$ not accounted for}{
-        record requirement:~``$V(v_1)$ is orthogonal to~$V(v_2)$''\;
-    }
-\end{algorithm}
+                different assigned vertices}
+            \State pick such a vertex~$w$ adjacent to the
+                    assigned ~$w_1$ and~$w_2$
+            \State assign~$V(w)=V(w_1) \times V(w_2)$
+            \State mark edges~$(v,w_1)$ and~$(v,w_2)$ as accounted for
+        \EndWhile
+    \EndWhile
+    \pagebreak[2]
+    \For{each pair of vertices~$(v_1, v_2)$}
+        \If{$(v_1,v_2)$ is not an edge}
+            \State record requirement:~``$V(v_1)$
+                    is not collinear to $V(v_2)$''
+        \EndIf
+    \EndFor
+    \For{each edge~$(v_1,v_2)$ not accounted for}
+        \State record requirement:~``$V(v_1)$ is orthogonal to~$V(v_2)$''
+    \EndFor
+\end{algorithmic}
+
 At two points in the algorithm, there is a choice which vertex to pick.

### 2014-09-18 paper/main.tex
paper: remove superfluous line

@@
 Therefore, we demonstrate: a Kochen-Specker system must have at least
-22 points.
-At the time of writing,
-we are computing
-whether there is a KS system of 22 vertices.\footnote{
+22 points.\footnote{%
 The authors have a wager whether there is a minimal KS system of less

### 2014-10-23 paper/main.tex
paper: remove linebreak after abstract per request
"-" -> "--"

@@
 \begin{abstract}
-    At the heart of the Conway's Free Will theorems and Kochen and Specker's
+    At the heart of the Conway--Kochen's
+    Free Will theorems and Kochen and Specker's
         argument against non-contextual hidden variable theories
@@
 \end{abstract}
-
-\clearpage
 

### 2014-10-23 paper/main.tex
Revert " "-" -> "--""
This reverts commit b440b7a1ae94b34e7fb09bc72303b2c4ee36082e.

--- 2014-09-18 paper/main.tex
@@
 \begin{abstract}
-    At the heart of the Conway's Free Will theorems and Kochen and Specker's
+    At the heart of the Conway-Kochen's
+    Free Will theorems and Kochen and Specker's
         argument against non-contextual hidden variable theories

### 2014-10-23 paper/main.tex
conway kochen free will theorem

--- 2014-09-18 paper/main.tex
@@
 \begin{abstract}
-    At the heart of the Conway's Free Will theorems and Kochen and Specker's
+    At the heart of the Conway-Kochen
+    Free Will Theorem and Kochen and Specker's
         argument against non-contextual hidden variable theories

### 2014-12-09 paper/main.tex
paper: modified for ngc

@@
-\documentclass[submission,copyright,creativecommons]{eptcs}
-
-\providecommand{\event}{QPL2014}
+\documentclass{report}
+\usepackage{ohm-ngc}
 
@@
 \usepackage{picins}
-\usepackage{amsthm}
 \usepackage{tikz}
@@
 
-\newcounter{main}
-\newtheorem{prop}[main]{Proposition}
-\newtheorem{comp}[main]{Computation}
-\newtheorem{cor}[main]{Corollary}
-\newtheorem{thm}[main]{Theorem}
-\newtheorem{lem}[main]{Lemma}
-\newtheorem{fact}[main]{Fact}
-\theoremstyle{definition}
-\newtheorem{dfn}[main]{Definition}
-\newtheorem{exa}[main]{Example}
-\newtheorem*{spin}{SPIN Axiom \cite{ck09}}
-\theoremstyle{remark}
-\newtheorem{rem}[main]{Remark}
-
-\title{A Kochen-Specker system has at least 22 vectors\\
-        {\small (extended abstract)%
-            \footnote{This is a condensed version of the article
-                        to be published in Ohmsha-Springer's
-                        New Generation Computing.
-                    One can find a preprint of the full version at
-                    \url{http://westerbaan.name/\~bas/math/ks.pdf}}}}
-\def\titlerunning{A Kochen-Specker system has at least 22 vectors
-            (extended abstract)}
-\def\authorrunning{Uijlen \& Westerbaan}
-
-\author{Sander Uijlen
-    \institute{Radboud Universiteit}
-    \email{suijlen@cs.ru.nl}
-\and
-    Bas Westerbaan
-    \institute{Radboud Universiteit}
-    \email{bwesterb@cs.ru.nl}}
+\newtheorem{fact}{Fact}[section]
+\newtheorem{comp}{Computation}[section]
+\newtheorem{spin}{SPIN Axiom \cite{ck09}}[section]
 
 \begin{document}
-
-\maketitle
+\title{A Kochen-Specker system has at least 22 vectors}
+\author{Sander Uijlen}{Radboud Universiteit, The Netherlands}
+\author{Bas Westerbaan}{Radboud Universiteit, The Netherlands}
+\E-mail{\{suijlen,bwesterb\}@cs.ru.nl}
 
@@
 
+\clearpage
 \section{Introduction}
@@
 \subsection{The experiment}
-
 Consider the following experiment.  Shoot a deuterium
@@
 axiom, that is:
-\begin{spin}
+\begin{spin} % TODO
     Given three pairwise orthogonal directions.
@@
 Their argument is based on the existence of a Kochen-Specker system.
-\begin{dfn}
+\begin{definition}
     A \keyword{Kochen-Specker (KS) system} is
@@
     \end{enumerate}
-\end{dfn}
+\end{definition}
 % TODO shorten this proof.
@@
 
-\begin{dfn}
+\begin{definition}
 Given a finite subset~$S$ of the projective plane
@@
 are orthogonal.
-\end{dfn}
-\begin{dfn}
+\end{definition}
+\begin{definition}
 A graph~$G$ is called~\keyword{embeddable},
@@
 such that~$G \leq G(S)$.
-\end{dfn}
+\end{definition}
 % TODO add remark on subgraph definition?
-\begin{dfn}
+\begin{definition}
 A graph is called~\keyword{010-colorable}
@@
 \end{enumerate}
-\end{dfn}
-
-\begin{dfn}
+\end{definition}
+
+\begin{definition}
 A \keyword{Kochen-Specker graph}
 is a embeddable graph that is not 010-colorable.
-\end{dfn}
+\end{definition}
 It is an easy, but important, consequence of the definitions that:
@@
 Luckily, we can restrict ourselves to certain classes of graphs.
-\begin{prop}[\cite{aow11}]
+\begin{proposition}[\cite{aow11}]
     An embeddable graph is squarefree.
@@
         For us the complete graph on four vertices is not squarefree.}
-\end{prop}
+\end{proposition}
 
@@
 Next, we show we can restrict ourselves to connected graphs.
-\begin{prop}[\cite{aow11}]\label{prop:ks-conn}
+\begin{proposition}[\cite{aow11}]\label{prop:ks-conn}
     A minimal Kochen-Specker graph is connected.
-\end{prop}
+\end{proposition}
 \begin{proof}
@@
 we consider another restriction.
-\begin{prop}
+\begin{proposition}
     A minimal Kochen-Specker graph has minimal vertex-order three.
     That is: every vertex is adjacent to at least three other vertices.
-\end{prop}
+\end{proposition}
 \begin{proof}
@@
 We continue with a strengthening of Proposition~\ref{prop:ks-conn}.
-\begin{prop}\label{prop:ks-biconn}
+\begin{proposition}\label{prop:ks-biconn}
 A minimal Kochen-Specker graph is edge-biconnected.
 That is: removing any single edge leaves the graph connected.
-\end{prop}
+\end{proposition}
 We need some preparation, before we can prove this Proposition.
-\begin{dfn}
+\begin{definition}
 Given a graph~$G$ and a vertex~$v$ of~$G$.
@@
 the vertex~$v$ is assigned color~$c$.
-\end{dfn}
+\end{definition}
 We are interested in these graphs because of the following observation.
-\begin{lem}\label{lem:biconn}
+\begin{lemma}\label{lem:biconn}
 If there is an embeddable graph~$G$ on~$n$ vertices with a vertex
@@
 then there is a Kochen-Specker graph on~$2n$ vertices.
-\end{lem}
+\end{lemma}
 \begin{proof}
@@
 
-\begin{prop}\label{prop:ks-triconn}
+\begin{proposition}\label{prop:ks-triconn}
     A minimal Kochen-Specker graph is edge-triconnected.
     That is: removing any two edges keeps the graph connected.
-\end{prop}
+\end{proposition}
 
@@
 
-\begin{dfn}
+\begin{definition}
    Given a graph~$G$ together with selected vertices~$v_1, \ldots, v_n \in G$.
@@
     called an~\keyword{$n$-type}.
-\end{dfn}
-
-\begin{exa}
+\end{definition}
+
+\begin{example}
 \begin{itemize}
@@
 \end{itemize}
-\end{exa}
+\end{example}
 
@@
         Thus both~$\#A \geq 17$ and~$\#B \geq 17$.
-        Hence~$\#G \geq 34$. Contradiction with minimality. \qedhere
+        Hence~$\#G \geq 34$. Contradiction with minimality. % \qedhere
     \end{enumerate}
@@
 \end{comp}
-\begin{cor}
+\begin{corollary}
     Any unconnected
@@
     With 20 vertices, there are exactly 25 of these.
-\end{cor}
+\end{corollary}
 This justifies, at this stage, not checking for connectedness.
@@
 First we give a pen-and-paper example.
-\begin{prop}\label{prop:unemb-10-2}
+\begin{proposition}\label{prop:unemb-10-2}
 The graph in Figure~\ref{fig:unemb-10-2}
 is not embeddable.
-\end{prop}
+\end{proposition}
 \begin{proof}
@@
 \begin{equation*}
-1 > |v_1w_2\left<v,w\right>| = |v_1w_2(v_1w_1+v_2w_2)| = 1. \qedhere
+1 > |v_1w_2\left<v,w\right>| = |v_1w_2(v_1w_1+v_2w_2)| = 1. %\qedhere
 \end{equation*}
@@
 
-\section{Acknowledgments}
+\begin{acknowledgment}
 We wish to thank the following for their generous contribution to the
@@
 for drawing figures for the draft version of this paper.
+\end{acknowledgment}
 
@@
 \bibliography{main}{}
-\bibliographystyle{eptcs}
+\bibliographystyle{kenta}
 

### 2014-12-10 paper/main.tex
add draft watermark
paper: triangles!

@@
 \usepackage{ohm-ngc}
+
+\usepackage{draftwatermark}
 
@@
 to graphs with minimal vertex order 3.
+We present another considerable restriction.
+
+\begin{proposition}\label{prop:ks-enoughtriangles}
+    In a minimal Kochen-Specker graph,
+        every node is part of a triangle.
+\end{proposition}
+\begin{proof}
+    Given any graph~$G$ together with a vertex~$v$ of~$G$
+        such that~$v$ is not part of a triangle.
+    Let~$G'$ denote~$G$ without~$v$.
+    Suppose~$G'$ is~$010$-colorable.
+    Then so is~$G$, as we may extend the coloring of~$G'$
+        to~$G$ by coloring~$v$ with~$0$.
+    Contraposing: if~$G$ is not~$010$-colorable,
+        then~$G'$ is not~$010$-colorable.
+    Clearly, if~$G$ is embeddable, then so is~$G'$.
+    Thus if~$G$ is Kochen-Specker, then so is~$G'$.
+\end{proof}
+There are only TODO squarefree non-isomorphic 
+graphs on 17 vertices with minimal vertex degree 3
+where every node is part of a triangle.
+Unfortunately,
+we could not find an efficient algorithm
+to restrict the enumeration of graphs
+to those where every node is part of a triangle.
 

### 2014-12-11 paper/main.tex
Add comp7 & fix some text
Signed-off-by: Bas Westerbaan <bas@westerbaan.name>

@@
     In a minimal Kochen-Specker graph,
-        every node is part of a triangle.
+        every vertex is part of a triangle.
 \end{proposition}
@@
 \end{proof}
-There are only TODO squarefree non-isomorphic 
+There are only~${\sim}10^5$ squarefree non-isomorphic 
 graphs on 17 vertices with minimal vertex degree 3
-where every node is part of a triangle.
+where every vertex is part of a triangle.
 Unfortunately,
@@
 to restrict the enumeration of graphs
-to those where every node is part of a triangle.
+to those where every vertex is part of a triangle.
 

### 2014-12-11 code/comp7.py

@@ -1,0 +1,23 @@
+""" Determines proporition of squarefree graphs with minimal
+    vertex degree three for which every node is in a triangle. """
+
+from nauty import geng
+
+from colorability import find_triangles, build_node_to_triangles
+
+def comp7():
+    n = 0
+    N = 0
+    prev_n = 0
+    for g in geng():
+        if len(g) != prev_n:
+            prev_n = len(g)
+            print len(g), n, N
+        N += 1
+        triangles = find_triangles(g)
+        node_to_triangles = build_node_to_triangles(g, triangles)
+        if all(node_to_triangles.values()):
+            n += 1
+
+if __name__ == '__main__':
+    comp7()

### 2014-12-11 paper/main.tex
proof of, etc

@@
 \usepackage{ohm-ngc}
+
+% Adaptation of ohm-ngc's proof environment to allow for a
+% "Proof of Proposition 2    lorum ipsum ..."
+\newenvironment{proofof}[1]{%
+    \par\noindent{\bfseries #1}\hskip2em}{%
+    \hfill \BK\vskip7pt}
 
@@
 
-\piccaption{Conway's 31 vector Kochen-Specker system\label{fig:conway31}}
+\piccaption{John Conway's 31 vector Kochen-Specker system\label{fig:conway31}}
 \parpic[r]{%
@@
 We follow \cite{aow11} and reduce the search for Kochen-Specker systems
-to the search of a certain class of graphs.
+to the search for a certain class of graphs.
 First note that in a Kochen-Specker system we may replace a point with its
@@
 A graph~$G$ is called~\keyword{embeddable},
-if it occurs as a subgraph of an orthogonality graph.
+if it is as a subgraph of an orthogonality graph.
 That is: if there is a finite subset~$S$ of the projective plane,
@@
 if there is a~$\{0,1\}$-coloring of the vertices,
-such that
+with
 \begin{enumerate}
@@
 A \keyword{Kochen-Specker graph}
-is a embeddable graph that is not 010-colorable.
+is an embeddable graph that is not 010-colorable.
 \end{definition}
@@
 is edge-biconnected.
-\begin{proof}[Proof of Proposition~\ref{prop:ks-biconn}]
+\begin{proofof}{Proof of Proposition~\ref{prop:ks-biconn}}
 Given a minimal Kochen-Specker graph~$G$.
@@
 Contradiction with~$G$'s minimality.
-\end{proof}
+\end{proofof}
 
@@
 
-\begin{proof}[Proof of Proposition~\ref{prop:ks-triconn}]
+\begin{proofof}{Proof of Proposition~\ref{prop:ks-triconn}}
 
@@
     \end{enumerate}
-\end{proof}
+\end{proofof}
 

### 2014-12-11 paper/main.tex
more fixes

@@
 \end{comp}
-
-We are ready to prove that a minimal Kochen-Specker graph
+We are ready to prove that a minimal KS graph
 is edge-biconnected.
@@
 Given a minimal Kochen-Specker graph~$G$.
-\parpic(0cm,1cm)[r]{%
+\parpic[r]{%
+%\parpic(0cm,1cm)[r]{%
     \begin{tikzpicture}[thick,scale=1.4,
@@
                 ;
-        \end{tikzpicture}}
+        \end{tikzpicture}}\vspace{-0.6mm}
 Recall it must be connected.  Suppose it is not edge-biconnected.
 Then there must be an edge~$(a,b)$ in~$G$,
-which removal disconnects~$G$.
+whose removal disconnects~$G$.
 Thus~$G$ decomposes into two connected graphs~$A$ and~$B$
-such that~$a\in A$, $b \in B$ and~$(a,b)$ is the only edge
+with~$a\in A$, $b \in B$ and~$(a,b)$ is the only edge
 between~$A$ and~$B$.
@@
 \end{proposition}
-
-Again, we need some preparation for the proof.  First, we generalize
+Again, we need some preparation.  First, we generalize
 the notion of fixed color.
-
 \begin{definition}
@@
 
-\parpic(0cm,1cm)[r]{%
+%\parpic(0cm,1cm)[r]{%
+\parpic[r]{%
 \begin{tikzpicture}[thick,scale=1.4,
@@
     and~$b_1,b_2 \in B$
-    such that~$(a_1, b_1)$ and~$(a_2,b_2)$
+    s.t.~$(a_1, b_1)$ and~$(a_2,b_2)$
     are the only edges between~$A$ and~$B$.
@@
             Note~$\#B \geq 17$ by Computation~\ref{comp:types}.
-
             Suppose there is a coloring of~$A$
@@
             Hence, by Computation~\ref{comp:types}, $\#B \geq 15$.
-
             Suppose~$a_1$ is not adjacent to~$a_2$.

### 2014-12-14 paper/main.tex
updates

@@
 
+\usepackage{etoolbox}
+
 % Adaptation of ohm-ngc's proof environment to allow for a
 % "Proof of Proposition 2    lorum ipsum ..."
-\newenvironment{proofof}[1]{%
-    \par\noindent{\bfseries #1}\hskip2em}{%
-    \hfill \BK\vskip7pt}
+\newbool{qedplaced} % keeps track whether QED symbol has been placed
+\newenvironment{myproof}[1][Proof]{%
+    \newcommand{\placeqed}{\ifmmode\quad\hbox{\hfill \BK}%
+                    \else\hfill \BK\vskip7pt%
+                    \fi}%
+    \global\boolfalse{qedplaced}%
+    \newcommand{\qed}{\ifbool{qedplaced}% Places QED symbol if not placed
+                              {}%
+                              {\placeqed\global\booltrue{qedplaced}}}%
+    \par\noindent{\bfseries #1}\hskip2em}{\qed}
 
@@
 \usepackage{paralist}
-\usepackage{etoolbox}
 \usepackage{graphicx}
@@
 is edge-biconnected.
-\begin{proofof}{Proof of Proposition~\ref{prop:ks-biconn}}
+\begin{myproof}[Proof of Proposition~\ref{prop:ks-biconn}]
 Given a minimal Kochen-Specker graph~$G$.
@@
 Contradiction with~$G$'s minimality.
-\end{proofof}
+\end{myproof}
 
@@
 
-\begin{proofof}{Proof of Proposition~\ref{prop:ks-triconn}}
+\begin{myproof}[Proof of Proposition~\ref{prop:ks-triconn}]
 
@@
         Thus both~$\#A \geq 17$ and~$\#B \geq 17$.
-        Hence~$\#G \geq 34$. Contradiction with minimality. % \qedhere
+        Hence~$\#G \geq 34$. Contradiction with minimality. \qed
     \end{enumerate}
-\end{proofof}
+\end{myproof}
 
@@
 \end{proposition}
-\begin{proof}
+\begin{myproof}
 Suppose it is embeddable.
@@
 Since~$v$ and~$w$ are not collinear,
-we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
-Now we find the contradiction:
+we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$. Thus
 \begin{equation*}
-1 > |v_1w_2\left<v,w\right>| = |v_1w_2(v_1w_1+v_2w_2)| = 1. %\qedhere
+1 > |v_1w_2\left<v,w\right>| = |v_1w_2(v_1w_1+v_2w_2)| = 1,
 \end{equation*}
-\end{proof}
+which is a contradiction.  Apparently the graph is not embeddable.
+\end{myproof}
 

### 2014-12-14 paper/main.tex
fix another parpic
remove lines for algorithmic (they were broken). fix some other things

@@
 \usepackage{tikz}
-
-% begin vertical rule patch for algorithmicx
-% http://tex.stackexchange.com/questions/144840/
-\makeatletter
-\newcommand*{\algrule}[1][\algorithmicindent]{\makebox[#1][l]{\hspace*{.5em}\vrule height .75\baselineskip depth .25\baselineskip}}%
-
-\newcount\ALG@printindent@tempcnta
-\def\ALG@printindent{%
-    \ifnum \theALG@nested>0%
-        \ifx\ALG@text\ALG@x@notext%
-            % do nothing
-            \addvspace{-3pt}%
-        \else
-            \unskip
-            % 
-            \ALG@printindent@tempcnta=1
-            \loop
-                \algrule[\csname ALG@ind@\the%
-                    \ALG@printindent@tempcnta\endcsname]%
-                \advance \ALG@printindent@tempcnta 1
-            \ifnum \ALG@printindent@tempcnta<\numexpr\theALG@nested+1\relax%
-            \repeat
-        \fi
-    \fi
-    }%
-% 
-\patchcmd{\ALG@doentity}{\noindent\hskip\ALG@tlm}%
-        {\ALG@printindent}{}{\errmessage{failed to patch}}
-\makeatother
-% end vertical line thingy
 
@@
 % TODO give intermediate computation.
-
+We propose a new method,
+which for all graphs we considered,
+could decide
+whether they were embeddable or not.
+First we give a pen-and-paper example.
 \piccaption{One of the two minimal non-embeddable graphs \label{fig:unemb-10-2}}
@@
 \end{tikzpicture}}
-
-We propose a new method,
-which for all graphs we considered,
-could decide
-whether they were embeddable or not.
-First we give a pen-and-paper example.
+\par\noindent\parbox{\textwidth}{%
 \begin{proposition}\label{prop:unemb-10-2}
@@
 is not embeddable.
-\end{proposition}
+\end{proposition}}
 \begin{myproof}
 Suppose it is embeddable.
-Consider~$p_1$.
+Consider (the point associated to)~$p_1$.
 It is orthogonal to both~$a$ and~$v$.
@@
 Similarly, $p_2$ is collinear to~$v \times p_1 = v \times (v \times a)$.
-Continuing in this fashion,
-we see that
+Continuing in this fashion, walking the circumference of the graph,
+we find
+\par\vskip-\lastskip % HACK to prevent eq. to mess with parpic
 \begin{equation}\label{eq:ue1}
@@
         \While{there are unassigned vertices adjacent to two
-                different assigned vertices}
+                distinct assigned}
             \State pick such a vertex~$w$ adjacent to the
@@
 \begin{comp}\label{comp:unemb20}
-    Every squarefree graph of minimal vertex order three
-    that is not 010-colorable
-    of order less than or equal to 20
+Every squarefree graph of minimal vertex order three that is
+not~$010$-colorable of order less than or equal 21
+contains an unembeddable subgraph.%
+\footnote{A list of these graphs together with their unembeddable
+subgraphs, can be found here:
+\url{http://kochen-specker.info/candidates/}.
+The source code for this computation can be found
+at~\texttt{code/comp2.py} of~\cite{GH}.}
+In particular, those of order less than or equal 20
     contains, as a subgraph, one of the following three graphs:
@@
 For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.
-
-Every squarefree graph of minimal vertex order three that is
-not~$010$-colorable of order 21 contains an unembeddable subgraph.%
-\footnote{A list of these graphs together with their unembeddable
-subgraphs, can be found here:
-\url{http://kochen-specker.info/candidates/}.
-The source code for this computation can be found
-at~\texttt{code/comp2.py} of~\cite{GH}.
-}
 \end{comp}

### 2014-12-14 paper/main.tex
paper: more improvements
paper: two typos

@@
    is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be colored.
-   That is: $t = \{ (c(v_1), \ldots, c(v_n));
-                \ c \in C(G) \}.$ 
+        That is:
+\begin{equation*}
+  t = \{ (c(v_1), \ldots, c(v_n));
+                \ c \in C(G) \}.
+\end{equation*}
     A type of~$n$ vertices is
@@
 
-Just as verteces with fixed color are rare,
+Just as vertices with fixed color are rare,
 we are interested in types,
@@
     Then it splits into two graphs~$A$ and~$B$
-    together with verteces~$a_1, a_2 \in A$
+    together with vertices~$a_1, a_2 \in A$
     and~$b_1,b_2 \in B$

### 2014-12-14 paper/main.tex
paper: more updates

@@
 \newtheorem{comp}{Computation}[section]
-\newtheorem{spin}{SPIN Axiom \cite{ck09}}[section]
 
@@
 axiom, that is:
-\begin{spin} % TODO
+
+% HACK to make a unnumbered "SPIN Axiom" theore environment.
+\begingroup\makeatletter\def\@thmcounter#1{}
+\newtheorem{spin}{SPIN Axiom}
+\begin{spin}[see \cite{ck09}]
     Given three pairwise orthogonal directions.
@@
 \end{spin}
+\def\@thmcounter#1{\noexpand\arabic{#1}}\makeatother\endgroup
+% /HACK
+
 Their argument is based on the existence of a Kochen-Specker system.
@@
 Luckily, we can restrict ourselves to certain classes of graphs.
-\begin{proposition}[\cite{aow11}]
+\begin{proposition}[by \cite{aow11}]
     An embeddable graph is squarefree.
@@
 Next, we show we can restrict ourselves to connected graphs.
-\begin{proposition}[\cite{aow11}]\label{prop:ks-conn}
+\begin{proposition}[by \cite{aow11}]\label{prop:ks-conn}
     A minimal Kochen-Specker graph is connected.
@@
     with minimal vertex order 3 and they have 10 vertices.
+    \footnote{They are the first five graphs
+            listed on:
+    \url{http://kochen-specker.info/smallGraphs/}.}
 \end{comp}
@@
 of a few vertices.  Then we derived cross-product expressions for the
-remaining vertices.  Finally, we find an equation relating some of
+remaining vertices.  Finally, we found an equation relating some of
 the cross-product expressions and show it is unsatisfiable.
@@
 of every squarefree graph with minimal vertex order three of
-less than 15, except for one.\footnote{
+less than 15, except for one.\footnote{%
 A list of all squarefree graphs with minimal vertex order three of

### 2014-12-14 paper/main.tex
paper: move AOW's graph to the right

@@
 We have verified the main result of \cite{aow11}:
-\begin{comp}
-There is a unique non-010-colorable squarefree connected graph on~$17$ or less
-vertices:
-\begin{center}
+
+\piccaption{AOW's 17-vertex non-010-colorable graph\label{fig:noncol17}}
+\parpic[r]{%
     \begin{tikzpicture}[thick,scale=1.4,
@@
             (90:0.2) -- (180:0.8)
-
             (90:0.2) -- (90:0.6) node[p]{}
@@
             (90:0.6) -- (90-45:1)
-
             (90:0.2) -- (270:0.4) node[p]{}
@@
         ;
-    \end{tikzpicture}
-\end{center}
+    \end{tikzpicture}}
+
+\begin{comp}
+There is only one non-010-colorable squarefree connected graph
+on less than~$18$ vertices.  See Figure~\ref{fig:noncol17}.
 It is not embeddable, as the graph in Figure~\ref{fig:unemb-10-2}

### 2014-12-14 paper/main.tex
paper: more improvements
paper: inline graphs

@@
 \newtheorem{comp}{Computation}[section]
+\newtheorem{algorithm}{Algorithm}[section]
 
@@
     An embeddable graph is squarefree.
-    That is: it does not contain the square as a subgraph.%
+    That is: it does not contain the square%
+    ~\begin{tikzpicture}[thick,scale=0.15,
+                        line width=0.4pt,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=2pt}]
+            \draw
+            (0,0) node[p]{} -- (0,1) node[p]{} --
+            (1,1) node[p]{} -- (1,0) node[p]{} -- (0,0);
+    \end{tikzpicture}
+as a subgraph.%
     \footnote{Some authors call a graph squarefree if it does not
         contain the square as induced subgraph.
-        For them the complete graph on four vertices is squarefree.
+        For them the complete graph on four vertices%
+    ~\begin{tikzpicture}[thick,scale=0.15,
+                        line width=0.3pt,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=2pt}]
+            \draw
+            (0,0) node[p]{} -- (0,1) node[p]{} --
+            (1,1) node[p]{} -- (1,0) node[p]{} -- (0,0)
+            (0,0) -- (1,1)
+            (1,0) -- (0,1);
+    \end{tikzpicture}
+        is squarefree.
         We follow~Weisstein\cite{sf-weisstein} and~Sloane\cite{sf-sloane} and
@@
 is edge-biconnected.
+\vskip7pt% HACK add space because the proof is not after a thm
 \begin{myproof}[Proof of Proposition~\ref{prop:ks-biconn}]
@@
 \begin{itemize}
-\item The triangle has~$3$-type~$\{(1,0,0), (0,1,0), (0,0,1)\}$.
+\item The triangle~\begin{tikzpicture}[thick,scale=0.15,
+                        line width=0.4pt,
+        p/.style={circle, draw, fill=white,
+                        inner sep=0pt, minimum width=2pt}]
+            \draw
+            (210:0.5) node[p]{} -- (330:0.5) node[p]{} --
+            (90:0.5) node[p]{} -- (210:0.5);
+    \end{tikzpicture}
+has~$3$-type~$\{(1,0,0), (0,1,0), (0,0,1)\}$.
 \item Every vertex in a Kochen-Specker graph has type~$\emptyset$.
@@
 
+\begin{algorithm}
 \begin{algorithmic}[5]
@@
 \end{algorithmic}
+\end{algorithm}
 

### 2014-12-15 paper/main.tex
paper: sander's wijzigingen

@@
                 difference arises from considering either squared
-                spin measurements $S^2$, or $1-S^2$.
+                spin measurements $S_v^2$, or $1-S_v^2$
+                for spin in direction~$v$.
                 %Then the colors correspond directly to the squared measurement
@@
 A point on the sphere obviously corresponds to a direction in space.
-Because of this, the term point, vector and direction
+Because of this, the terms point, vector and direction
 can be used interchangeably. Antipodal points correspond to opposite
@@
 required more time than would be gained by reducing the number of graphs.
+Cf.~Corollary~\ref{cor:littlecon}.
 
@@
    Given a graph~$G$ together with selected vertices~$v_1, \ldots, v_n \in G$.
-   Let~$C(G)$ denote the set of $010$-colorings of~$G$.
    The \keyword{type~$t$ of~$(v_1,\ldots,v_n)$ (in~$G$)}
@@
 \begin{equation*}
-  t = \{ (c(v_1), \ldots, c(v_n));
-                \ c \in C(G) \}.
+  t = \bigl\{ (c(v_1), \ldots, c(v_n)); \ 
+        c \colon G \to \{0,1\},\ 
+        \text{$c$ is a $010$-coloring of~$G$} \bigr\}
 \end{equation*}
@@
     and~$b_1,b_2 \in B$
-    s.t.~$(a_1, b_1)$ and~$(a_2,b_2)$
+    such that~$(a_1, b_1)$ and~$(a_2,b_2)$
     are the only edges between~$A$ and~$B$.
@@
 \end{comp}
-\begin{corollary}
+\begin{corollary}\label{cor:littlecon}
     Any unconnected
@@
 \end{equation*}
-which is a contradiction.  Apparently the graph is not embeddable.
+which is a contradiction.  Apparently this graph is not embeddable.
 \end{myproof}
@@
 
-In this way, there were still a few (010-colorable) graphs of which
-we could not decide embeddability.
-With some guessing,
-we determined embeddings for these graphs by hand.
+Using this algorithm, we could determine embeddability for most graphs
+with less than 15 vertices.  However, there were  a few (010-colorable)
+graphs for which the algorithm would not terminate and hence
+their embeddability was not decided.
+By hand, we found with trial and error an embedding for
+some of these graphs.
 Once we knew the troublesome graphs were embeddable,
@@
 requires a new insight.
-For instance: another restriction on which graphs to consider.
+Most likely, this would be a new
+practical restriction on which graphs to consider.
+In particular
+(following Prop.~\ref{prop:ks-enoughtriangles})
+a fast algorithm to enumerate
+square-free graphs of minimal vertex order three
+where each node is part of a triangle,
+would probably allow to test for KS-systems
+up to and including 24 vertices.
 

### 2014-12-15 paper/main.tex
paper: Attribute Spitters

@@
 
-We are grateful to prof.~McKay for discussing
-the feasibility of certain graph restrictions and
-to Judith van Stegeren\footnote{\url{http://jd7h.com/}}
-for drawing figures for the draft version of this paper.
+We are grateful to prof.~McKay for discussing the feasibility of
+certain graph restrictions; to Bas Spitters for showing us the
+article of Arends, Ouaknine and Wampler \emph{and} to Judith van
+Stegeren\footnote{\url{http://jd7h.com/}} for drawing figures for
+the draft version of this paper.
 \end{acknowledgment}

### 2014-12-16 paper/main.tex
paper: fix bibtex
paper: sander's text on non-3d systems

@@
 
-% TODO refer to non-3d systems
+Before we continue, we would like to make two remarks.
+\begin{enumerate}
+\item
+There is an inconsistency in the literature on how to count the number
+of points in a Kochen-Specker system.  One way is to take the minimal
+number of points required to be non-010-colorable.
+This is how we count in this paper.
+Another way is to require that for every pair of orthogonal points
+in the system, there is a third point orthogonal to both.
+This way of counting is used in~\cite{pavivcic}.
+\item
+We consider three-dimensional Kochen-Specker systems
+for two reasons.
+First, every 3-dimensional system can be seen as a higher-dimensional
+            system.   The converse if false.
+Secondly, the size of the least Kochen-Specker system
+            in dimension 4 and higher has already been found.\cite{pavivcic}
+            The three-dimensional case is still open.
+\end{enumerate}
+
 \subsection{Overview}
@@
 \bibliography{main}{}
-\bibliographystyle{kenta}
+\bibliographystyle{plain}
 

### 2014-12-16 paper/main.tex
paper: fix spacing of parpics
paper: udpates
Signed-off-by: Bas Westerbaan <bas@westerbaan.name>

@@
 \title{A Kochen-Specker system has at least 22 vectors}
-\author{Sander Uijlen}{Radboud Universiteit, The Netherlands}
-\author{Bas Westerbaan}{Radboud Universiteit, The Netherlands}
+\author{Sander Uijlen \& Bas Westerbaan}{Radboud Universiteit, The Netherlands}
 \E-mail{\{suijlen,bwesterb\}@cs.ru.nl}
@@
 \undef\St
+\node[]() at (0, -4){}; % HACK to get spacing right
 \end{tikzpicture}}
@@
         ;
+\node[]() at (0, -1.4){}; % HACK to get spacing right
     \end{tikzpicture}}
@@
 whether they were embeddable or not.
-First we give a pen-and-paper example.
+But first, we give two pen-and-paper proofs of the non-embeddability
+of the following graph. 
+These two pictures
+represent a single connected graph of minimal vertex degree three.
+For presentation,
+some vertices have been drawn twice.
+\begin{center}
+\footnotesize
+\begin{tikzpicture}[thick,scale=1.4,%
+        p/.style={circle, fill=white,%
+                        inner sep=1pt, minimum width=5pt}]
+            \draw
+                (0:0) node [p] {18} (0:0.8) node [p] {3}
+                (60:0.8) node [p] {10} (120:0.8) node [p] {0}
+                (180:0.8) node [p] {5} (240:0.8) node [p] {12}
+                (300:0.8) node [p] {9} (15:1.2) node [p] {13}
+                (45:1.2) node [p] {14} (75:1.2) node [p] {4}
+                (105:1.2) node [p] {6} (135:1.2) node [p] {15}
+                (165:1.2) node [p] {11} (195:1.2) node [p] {16}
+                (225:1.2) node [p] {1} (255:1.2) node [p] {7}
+                (285:1.2) node [p] {17} (315:1.2) node [p] {2}
+                (345:1.2) node [p] {8}
+                \foreach \x in {0, 120, 240} { (\x:0.8) -- (\x-60:0.8) }
+                \foreach \x in {0,60,120,180,240,300} {
+                    (0:0) -- (\x:0.8) (\x:0.8) -- (\x+15:1.2) 
+                    (\x:0.8) -- (\x-15:1.2) (\x+15:1.2) -- (\x-15:1.2)
+                }
+            ;
+    \end{tikzpicture}\qquad
+\begin{tikzpicture}[thick,scale=1.4,%
+        p/.style={circle, fill=white,%
+                        inner sep=1pt, minimum width=5pt}]
+            \draw
+                (90:0.3) node[p] {14} (210:0.3) node[p] {8}
+                (330:0.3) node[p] {1} (90:0.8) node[p] {4}
+                (210:0.8) node[p] {13} (330:0.8) node[p] {7}
+                (60:1.3) node[p] {17} (120:1.3) node[p] {11}
+                (180:1.3) node[p] {16} (240:1.3) node[p] {6}
+                (300:1.3) node[p] {15} (0:1.3) node[p] {2}
+                \foreach \x in {90, 210, 330} {
+                    (\x:0.3) -- (\x+120:0.3)
+                    (\x:0.3) -- (\x:0.8)
+                    (\x:0.8) -- (\x+30:1.3)
+                    (\x:0.8) -- (\x-30:1.3)
+                    (\x-30:1.3) -- (\x+30:1.3)
+                    (\x+30:1.3) -- (\x-30+120:1.3)
+                }
+            ;
+    \end{tikzpicture}
+\end{center}
+\subsection{First pen-and-paper proof}
+Of the graphs of 19 vertices,
+the above graph was the
+the only not excluded by the unembeddable graphs
+known by AOW.
+%[referentie naar IpD?GUbV?].
+In order to show that this graph is not embeddable,
+we proceded as follows.
+If the graph is embeddable, there is always an embedding such that
+the coordinates on the sphere of some triangle in this graph
+are~$(1,0,0)$, $(0,1,0)$ and~$(0,0,1)$.
+We chose to assign these values
+to the points 0, 10 and~18, respectively. For each of these vertices,
+any vertex adjacent to it
+must lie on a certain great circle.
+For example, we can assign to the vertices 9 and 3, which
+are both connected to 18, the coordinates~$(\sqrt{1-x^{2}}, -x,0)$
+and~$(x,\sqrt{1-x^{2}},0)$, for some $x\in(-1,1)$ and~$x\neq 0$.
+We find similar
+expressions for the coordinates of the other triangles containing the
+vertices 0, 10 and 18. The remaining vertices are assigned a variable
+value on the sphere, where we always choose the component containing
+the square-root positive, so we obtain coordinates of the
+form~$(a,b,\sqrt{1-a^2-b^2})$, for $a,b \in (-1,0)\cup (0,1)$.
+In this way we
+obtain no variables for the first triangle, 4 for attached triangles
+and then 16 for the left over vertices giving 20 unknowns. The
+relations for these variables are given by the fact that the inner
+product of the vectors should be zero if they are adjacent. Not
+counting the inner products that are trivially zero, we obtain 24
+equations. By tedious calculations, one
+derives the equation $x^2(1-x^2) = -1$, which is not satisfiable.
+Hence this system is not embeddable.
+
+If we only consider the points used to obtain the
+contradiction, we find an unembeddable subgraph of the system
+containing 13 vertices\footnote{These are the vetices
+0,2,3,4,6,8,9,10,13,14,15,17 and 18.}.
+In the appendix we included the  derivation of contradiction.
+In theory, this approach can be mechanically performed by
+a computer.  However, there are too many variables
+to solve this in reasonable time.
+
+\subsection{Second pen-and-paper proof}
+We give a second simpler pen-and-paper proof of the unembeddability
+of the graph, by showing a subgraph is unembeddable.
+
 \piccaption{One of the two minimal non-embeddable graphs \label{fig:unemb-10-2}}
@@
                 (330-30:1.3) node[below]{$a$} ;
+\node[]() at (0, -1.5){}; % HACK to get spacing right
 \end{tikzpicture}}
@@
 
+\subsection{An algorithm to decide embeddability}
 In the previous proof, we fixed, without loss of generality, the position
@@
 the cross-product expressions and show it is unsatisfiable.
-We can automate this reasoning as follows.
+We automate this reasoning as follows.
 

### 2014-12-16 paper/main.tex
paper: updates

@@
 in the system, there is a third point orthogonal to both.
-This way of counting is used in~\cite{pavivcic}.
+This way of counting is for example used in~\cite{pavivcic}.
 \item
@@
 
-% attribute mckay for answering questions?
-% TODO attribute geng
 \clearpage
@@
 
+\appendix
+\clearpage
+\section{Appendix}
+In this appendix, we show by hand, that indeed the 19 vertex graph
+considered at the start of section 3 %TODO
+is indeed not embeddable.
+We restrict the calculation to the~$13$-vertex unembeddable
+subgraph.
+
+\begin{center}
+    \footnotesize
+    \begin{tikzpicture}[thick,scale=3.5,
+        p/.style={fill=white, circle, inner sep=1pt, minimum width=45pt}]
+        \draw
+
+                (45:0.9) node[p, align=center]{0 \\ $(1,0,0)$} -- (65:1.3) node[p, align=center]{6 \\$(0,x_3,\cdot)$}
+                (45:0.9)  -- (25:1.3) node[p, align=center]{15 \\ $(0,\cdot,\text{-}x_3)$}
+                (65:1.3)  -- (25:1.3)
+
+                (90+45:0.9) node[p, align=center]{3 \\ $(\cdot,\text{-}x_1,0)$} -- (90+65:1.3) node[p, align=center]{8 \\ $(g,h,\cdot)$ }
+                (90+45:0.9)  -- (90+25:1.3) node[p, align=center]{13 \\ $(e,f,\cdot)$}
+                (90+65:1.3)  -- (90+25:1.3)
+
+                (180+45:0.9) node[p, align=center]{10 \\ $(0,1,0)$} -- (180+65:1.3) node[p, align=center]{4 \\ $(x_2,0,\cdot)$}
+                (180+45:0.9)  -- (180+25:1.3) node[p, align=center]{14 \\ $(\cdot,0,\text{-}x_2)$}
+                (180+65:1.3)  -- (180+25:1.3)
+
+                (270+45:0.9) node[p, align=center]{9 \\$(x_1,\cdot,0)$} -- (270+65:1.3) node[p, align=center]{2 \\ $(a,b,\cdot)$}
+                (270+45:0.9)  -- (270+25:1.3) node[p, align=center]{17 \\ $(c,d,\cdot)$}
+                (270+65:1.3)  -- (270+25:1.3)
+
+            \foreach \x in {0,180} {
+                (\x+25:1.3) -- (\x-25:1.3)
+                (\x+65:1.3) -- (\x+115:1.3)
+           }
+
+        (90:0.6) -- (90:0.6)  node[p, align=center]{18 \\ $(0,0,1)$}
+        (90:0.6) -- (270+45:0.9)
+        (90:0.6) -- (270-45:0.9)
+        (90:0.6) -- (90+45:0.9)
+	(90:0.6) -- (90-45:0.9)
+
+(45:1)--(225:1)
+(135:1)--(315:1) ;
+    \end{tikzpicture}
+\end{center}
+
+This is the subgraph and we have given the vertices names corresponding
+to the original, 19 vertex, graph. Below each name is a coordinate
+on the sphere (whenever there is a $\cdot$ in the coordinate, we
+mean that this component of the coordinate is fixed by the other
+components, as it must be a point on the sphere, where we do take
+this $\cdot$ to be the positive square root).
+
+Now if two vertices are connected via an edge, we have that the
+inner product of their coordinates is zero. Vertices connected to
+the points 0,10 and 18 are trivially zero. The non trivial equations
+then become:
+\begin{align}
+&ax_1 + b\sqrt{1-{x_1}^2}=0\\
+&cx_1 + d\sqrt{1-{x_1}^2}=0\\
+&e\sqrt{1-{x_1}^2} - fx_1=0\\
+&g\sqrt{1-{x_1}^2} - hx_1=0\\
+&g\sqrt{1-{x_2}^2} - x_2 \sqrt{1-g^2-h^2}=0\\
+&cx_2 + \sqrt{1-{x_2}^2}\sqrt{1-c^2-d^2}=0\\
+&b\sqrt{1-{x_3}^2} - x_3 \sqrt{1-a^2-b^2}=0\\
+&fx_3 + \sqrt{1-e^2-f^2}\sqrt{1-{x_3}^2}=0\\
+&ac + bd + \sqrt{1-a^2-b^2}\sqrt{1-c^2 - d^2} = 0\\
+&eg + fh + \sqrt{1-e^2-f^2}\sqrt{1-g^2 - h^2} = 0
+\end{align}
+
+From the equations containing only $x_1$ we can eliminate four variables.
+\begin{equation}
+b = \frac{-a x_1}{\sqrt{1-{x_1}^2}} \tag{1'}
+\end{equation}
+\begin{equation}
+d = \frac{-c x_1}{\sqrt{1-{x_1}^2}} \tag{2'}
+\end{equation}
+\begin{equation}
+e = \frac{f x_1}{\sqrt{1-{x_1}^2}} \tag{3'}
+\end{equation}
+\begin{equation}
+g = \frac{h x_1}{\sqrt{1-{x_1}^2}} \tag{4'}
+\end{equation}
+Note that this is possible since all $x_i \in (-1,1)\backslash
+\{0\}$ since otherwise we would have coinciding coordinates of
+vertices in the graph.
+
+Substituting these expressions in the other six equations and using
+that~$\tfrac{x}{1-x} +1 = \tfrac{1}{1-x}$, we obtain:
+
+\begin{align}
+&\frac{h x_1}{\sqrt{1-{x_1}^2}}\sqrt{1-{x_2}^2} - x_2 \sqrt{1-\frac{h^2}{1-{x_1}^2}}=0 \tag{5'}\\
+&cx_2 + \sqrt{1-{x_2}^2}\sqrt{1-\frac{c^2}{1-{x_1}^2}}=0 \tag{6'}\\
+&\frac{-a x_1}{\sqrt{1-{x_1}^2}}\sqrt{1-{x_3}^2} - x_3 \sqrt{1-\frac{a^2}{1-{x_1}^2}}=0\tag{7'}\\
+&fx_3 + \sqrt{1-\frac{f^2}{1-{x_1}^2}}\sqrt{1-{x_3}^2}=0\tag{8'}\\
+&ac + \frac{-a x_1}{\sqrt{1-{x_1}^2}}\frac{-c x_1}{\sqrt{1-{x_1}^2}} + \sqrt{1-\frac{a^2}{1-{x_1}^2}}\sqrt{1-\frac{c^2}{1-{x_1}^2}} = 0 \tag{9'}\\
+&\frac{f x_1}{\sqrt{1-{x_1}^2}}\frac{h x_1}{\sqrt{1-{x_1}^2}} + fh + \sqrt{1-\frac{f^2}{1-{x_1}^2}}\sqrt{1-\frac{h^2}{1-{x_1}^2}} = 0\tag{10'}
+\end{align}
+
+These can then almost trivially be rewritten to:
+
+\begin{align}
+&h x_1\sqrt{1-{x_2}^2} - x_2 \sqrt{1-{x_1}^2 - h^2}=0 \tag{5''}\label{5''}\\
+&cx_2\sqrt{1-{x_1}^2} + \sqrt{1-{x_2}^2}\sqrt{1-{x_1}^2 - c^2}=0\tag{6''}\label{6''}\\
+&a x_1\sqrt{1-{x_3}^2} + x_3 \sqrt{1-{x_1}^2 - a^2}=0\tag{7''}\label{7''}\\
+&fx_3\sqrt{1-{x_1}^2} + \sqrt{1-{x_1}^2 - f^2}\sqrt{1-{x_3}^2}=0\tag{8''}\label{8''}\\
+&ac + \sqrt{1- {x_1}^2 - a^2}\sqrt{1- {x_1}^2 - c^2} = 0\tag{9''}\label{9''}\\
+&fh + \sqrt{1- {x_1}^2 - f^2}\sqrt{1- {x_1}^2 - h^2} = 0\tag{10''}\label{10''}
+\end{align}
+
+It then follows from \eqref{7''}, \eqref{9''} and \eqref{6''} that:
+\begin{eqnarray*}
+\frac{\sqrt{1-{x_3}^2}}{x_3} &=& \frac{-\sqrt{1-{x_1}^2 - a^2}}{ax_1}\\
+&=& \frac{c}{x_1 \sqrt{1-{x_1}^2 -c^2}}\\
+&=& \frac{-\sqrt{1-{x_2}^2}}{x_1 x_2 \sqrt{1-{x_1}^2}}
+\end{eqnarray*}
+
+While using \eqref{5''},\eqref{10''} and \eqref{8''} we find that:
+\begin{eqnarray*}
+\frac{\sqrt{1-{x_2}^2}}{x_2} &=& \frac{\sqrt{1-{x_1}^2 - h^2}}{hx_1}\\
+&=& \frac{-f}{x_1 \sqrt{1-{x_1}^2 -f^2}}\\
+&=& \frac{\sqrt{1-{x_3}^2}}{x_1 x_3 \sqrt{1-{x_1}^2}}
+\end{eqnarray*}
+
+Inserting this expression for $\frac{\sqrt{1-{x_2}^2}}{x_2}$ into
+the expression for $\frac{\sqrt{1-{x_3}^2}}{x_3}$ we obtain the
+equation:
+
+\begin{equation}
+\frac{\sqrt{1-{x_3}^2}}{x_3} = \frac{-\sqrt{1-{x_3}^2}}{x_3 {x_1}^2  (1-{x_1}^2)}
+\end{equation}
+But then, eliminating the $x_3$ terms we find:
+\begin{equation}
+{x_1}^2 (1-{x_1}^2) = -1
+\end{equation}
+Which is our desired contradicton.
+
 \end{document}

### 2015-01-30 paper/main.tex
paper: remove watermark
Related work & typo's

@@
     \par\noindent{\bfseries #1}\hskip2em}{\qed}
-
-\usepackage{draftwatermark}
 
@@
 we describe our practical embeddability test.
-
+In Section~\ref{sec:related},
+we discuss related work.
 The software and results of the various computations performed for
 this paper, can be found here\cite{GH}.
+
+This is an extended version of the paper published in proceedings
+of the 11th workshop on Quantum Physics and Logic.\cite{UW14}
+Compared to the results presented on the workshop,
+we have improved the lower-bound to~$22$.
+We also include new restrictions on the graph enumeration
+of which we would like to emphasize~Prop.~\ref{prop:ks-enoughtriangles},
+which promises to push the bound when implemented efficiently.
+Furthermore we expand on methods to prove unembeddability.
+Finally we discuss how the related results in the literature
+compare to ours.
 
@@
     \end{tikzpicture}.
-    Because the two vertices are adjacent, this occurance of
+    Because the two vertices are adjacent, this occurrence of
     the type is called trivial.
@@
     \item Suppose~$a_1 = a_2$ and~$b_1 \neq b_2$.
-            This leads to a contradication in the same way
+            This leads to a contradiction in the same way
             as in case~\ref{tricase1}.
@@
 In order to show that this graph is not embeddable,
-we proceded as follows.
+we proceeded as follows.
 If the graph is embeddable, there is always an embedding such that
@@
 contradiction, we find an unembeddable subgraph of the system
-containing 13 vertices\footnote{These are the vetices
+containing 13 vertices\footnote{These are the vertices
 0,2,3,4,6,8,9,10,13,14,15,17 and 18.}.
@@
 
+\section{Related work}
+\label{sec:related}
+This paper improves the lower bound of
+Arends, Ouaknine and Wampler\cite{a09,aow11}
+on the size of a minimal Kochen-Specker system.
+Pavi\v{c}i\'c, Merlet, McKay and Norman\cite{pavivcic}
+count the size of a Kochen-Specker system differently
+and have demonstrated a different bound.
+\begin{definition}
+    A Kochen-Specker graph is called \keyword{complete}
+        if every edge is part of a triangle.
+\end{definition}
+Pavi\v{c}i\'c et al have given a lower bound on the size
+of the minimal \emph{complete} Kochen-Specker system.
+Similar to our approach, they reduce the problem
+to a question about discrete structures
+which can be considered as a subclass of graphs.
+They discovered several uncolorable candidates
+of which they could prove unembeddability
+using either symbolic computation or interval analysis.
+
+Clearly, a bound on the minimal size of a KS system
+is also a bound on the minimal size of a complete system.
+Conversely, a KS system can be completed by adding missing vectors.
+However, it is not clear how many vectors have to be added in general
+to complete the system.
+A history of the different bounds on 3-dimensional Kochen-Specker
+systems is given in the table below.
+\begin{center}
+    \begin{tabular}{llll}
+        Authors & year & KS & complete KS \\ \hline
+        Kochen and Specker\cite{ks} & 1975 & $\leq 117$ & $\leq 192$ \\
+        Penrose, Peres\cite{peres}
+                (independendly)& 1991 & $\leq 33$ & $\leq 57$ \\
+        Conway & $\sim$1995 & $\leq 31$ & $\leq 51$ \\
+        UW\cite{UW14} & 2014 & $\geq 22$ & $\geq 22$ \\
+        Pavi\v{c}i\'c et al.\cite{pavivcic}& 2004 & $?$ & $\geq 30$ \\
+    \end{tabular}
+\caption{A history of the bounds on the size of the minimal
+    (complete) three-dimensional Kochen-Specker system.}
+\end{center}
+
 \section{Conclusion and future research}
@@
 Enumerating all candidate graphs of less than 31 vertices
-is computationally infeasable.
+is computationally infeasible.
 To bridge the enormous the gap between 22 and 31,
@@
 considered at the start of section 3 %TODO
-is indeed not embeddable.
+is not embeddable.
 We restrict the calculation to the~$13$-vertex unembeddable
@@
 \end{equation}
-Which is our desired contradicton.
+Which is our desired contradiction.
 

### 2015-05-11 paper/main.tex
nonconnected -> disconnected

@@
 \begin{proof}
-    Suppose~$G$ is a non-connected Kochen-Specker graph.
+    Suppose~$G$ is a disconnected Kochen-Specker graph.
     Then one of its components is not 010-colorable.
@@
 There are only~${\sim}10^9$ non-isomorphic squarefree graphs on~$17$
-vertices that are not connected.
+vertices that are disconnected.
 In our computations, checking for connectedness
@@
 they seem to be of little use as a practical restriction.
-Concerning excluding unconnected graphs:
+Concerning excluding disconnected graphs:
 % TODO Add argument or computation for {2,3}-connected
@@
 \begin{corollary}\label{cor:littlecon}
-    Any unconnected
+    Any disconnected
     squarefree graph with minimal vertex order 3

### 2015-05-11 paper/main.tex
typo
paper: remove superfluous "do"

@@
 A graph~$G$ is called~\keyword{embeddable},
-if it is as a subgraph of an orthogonality graph.
+if it is a subgraph of an orthogonality graph.
 That is: if there is a finite subset~$S$ of the projective plane,
@@
 mean that this component of the coordinate is fixed by the other
-components, as it must be a point on the sphere, where we do take
+components, as it must be a point on the sphere, where we take
 this $\cdot$ to be the positive square root).

### 2015-05-11 paper/main.tex
paper: /010/-colored
paper: add count

@@
    The \keyword{type~$t$ of~$(v_1,\ldots,v_n)$ (in~$G$)}
-   is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be colored.
+   is the set of all possible ways $\{v_1, \ldots, v_n\}$ can be $010$-colored.
         That is:
@@
 We enumerated all squarefree graphs with minimal vertex
-order 3 on less than or equal~$21$ vertices,
+order 3 on less than or equal~$21$ vertices\footnote{%
+On 21 vertices, we found 24,888,945,914,244 such graphs.},
 using the~\texttt{geng} util of the nauty software package,

### 2015-05-19 paper/main.tex
paper: if -> is
paper: neater sentence
paper: add more graphs

@@
 First, every 3-dimensional system can be seen as a higher-dimensional
-            system.   The converse if false.
+            system.   The converse is false.
 Secondly, the size of the least Kochen-Specker system
@@
     They intersect in precisely two antipodal points.
-    Hence, if~$c$ and~$d$ are both orthogonal to~$v$ and~$w$,
+    Hence, if~$c$ and~$d$ are points, both orthogonal to~$v$ and~$w$,
     then~$c$ and~$d$ are equivalent.
@@
     squarefree connected graphs
-    with minimal vertex order 3 and they have 10 vertices.
-    \footnote{They are the first five graphs
+    with minimal vertex order 3 and they have 10
+    vertices.\footnote{They are the first five graphs
             listed on:
     \url{http://kochen-specker.info/smallGraphs/}.}
+    Three of them are embeddable and shown below.
+    The remaining two
+    are depicted in Comptutation~\ref{comp:unemb20}.
+\begin{center}
+\begin{tikzpicture}[thick,scale=1.0,
+    p/.style={circle, draw, fill=white,
+                    inner sep=0pt, minimum width=5pt}]
+
+        \draw
+                (30:0.5) node[p]{} -- (65:1.3) node[p]{}
+                (30:0.5)  -- (25:1.3) node[p]{}
+                (65:1.3)  -- (25:1.3)
+
+                (150:0.5) node[p]{} -- (155:1.3) node[p]{}
+                (150:0.5)  -- (115:1.3) node[p]{}
+                (155:1.3)  -- (115:1.3)
+
+            (90:0.5) node[p]{}
+            (90:-0.5) node[p]{}
+            (115:1.3) -- (65:1.3) -- (90:0.5) -- (115:1.3)
+            (90:0.5) -- (90:-0.5)
+            (350:1) node[p]{}
+            (190:1) node[p]{}
+            (155:1.3) -- (190:1) -- (270:0.5) -- (350:1) -- (25:1.3)
+            (150:0.5) -- (350:1)
+            (30:0.5) -- (190:1)
+            ;
+    \end{tikzpicture}
+    \qquad
+    \begin{tikzpicture}[thick,scale=1.0,
+    p/.style={circle, draw, fill=white,
+                    inner sep=0pt, minimum width=5pt}]
+        \draw
+            \foreach \x in {0,60,120,180,240,300} {
+                (\x:1.3) node[p]{} -- (\x+60:1.3) node[p]{}
+            }
+	(0,0) node[p]{} -- (270:0.5) node[p]{}
+	(0,0) -- (45:0.5) node[p]{} -- (300:1.3)
+	(0:0) -- (135:0.5) node[p]{} -- (240:1.3)
+	(180:1.3) -- (270:0.5) -- (0:1.3)
+	(60:1.3) -- (135:0.5)
+	(120:1.3) -- (45:0.5);
+    \end{tikzpicture}
+    \qquad
+    \begin{tikzpicture}[thick,scale=1.0,
+    p/.style={circle, draw, fill=white,
+                    inner sep=0pt, minimum width=5pt}]
+                            \draw
+                (30:0.5) node[p]{} -- (65:1.3) node[p]{}
+                (30:0.5)  -- (25:1.3) node[p]{}
+                (65:1.3)  -- (25:1.3)
+                (150:0.5) node[p]{} -- (155:1.3) node[p]{}
+                (150:0.5)  -- (115:1.3) node[p]{}
+                (155:1.3)  -- (115:1.3)
+                (90:1) node[p]{}
+                (90:-0.5) node[p]{}
+                (65:1.3) -- (90:1) -- (115:1.3)
+                (90:1) -- (90:-0.5)
+                (350:1) node[p]{}
+                (190:1) node[p]{}
+                (155:1.3) -- (190:1) -- (270:0.5) -- (350:1) -- (25:1.3)
+                (150:0.5) -- (350:1)
+                (30:0.5) -- (190:1) ;
+\end{tikzpicture}
+\end{center}
 \end{comp}

### 2015-05-21 paper/main.tex
paper: second cross product unemb. proof

@@
 \newcommand{\keyword}[1]{\textbf{#1}}
+\newcommand{\coll}{\mathrel{\text{coll.}}}
 
@@
 See Proposition~\ref{prop:unemb-10-2}.
-For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.
+For the second graph, we also have a similar direct proof.
+See Proposition~\ref{prop:unemb-10-1} in the appendix.
+The third graph is shown to not be embeddable using our
+algorithm.
 \end{comp}
@@
 
+% Apparently, \appendix does not work properly.
+% Thus, we set \thesubsection by hand
 \appendix
 \clearpage
-\section{Appendix}
+\renewcommand{\thesubsection}{\Alph{subsection}}
+\renewcommand{\thesection}{\Alph{subsection}}
+\section*{Appendix}
+\subsection{Original pen-and-paper proof}
 In this appendix, we show by hand, that indeed the 19 vertex graph
@@
 
+\subsection{Second cross-product unembeddability proof}
+
+\parpic(5cm,3.5cm)[r]{%
+\begin{tikzpicture}[thick,scale=1.4,
+    p/.style={circle, draw, fill=white,
+                    inner sep=0pt, minimum width=5pt}]
+        \draw
+            \foreach \x in {0,180} {
+                (\x+15:1.3) node[p]{} -- (\x+90:0.9) node[p]{}
+                (\x+15:1.3) -- (\x+45:0.5) node[p]{}
+                (\x+45:0.5) -- (\x+90:0.9)
+                (\x-15:1.3) node[p]{} -- (\x-90:0.9)
+                (\x-15:1.3) -- (\x-45:0.5) node[p]{}
+                (\x-45:0.5) -- (\x-90:0.9)
+                (\x-15:1.3) -- (\x+15:1.3)
+            }
+            (45:0.5) -- (45+180:0.5)
+            (45+90:0.5) -- (45+90+180:0.5)
+            (90:0.9) node[above] {$x$}
+            (270:0.9) node[below] {$w$}
+            (15:1.3) node[right] {$y$}
+            (-15:1.3) node[right] {$v$}
+            (45:0.5) node[below] {$z$}
+            (-45:0.5) node[above] {$p_5$}
+            (135:0.5) node[below] {$p_1$}
+            (-135:0.5) node[above] {$p_4$}
+            (165:1.3) node[left] {$p_2$}
+            (-165:1.3) node[left] {$p_3$}
+        ;
+    \end{tikzpicture}}
+
+\par\noindent\parbox{\textwidth}{%
+\begin{proposition}\label{prop:unemb-10-1}
+The graph on the right is not embeddable.
+\end{proposition}}
+\begin{myproof}
+    Suppose it is embeddable.
+    Note (the point associated to) $p_5$
+        is orthogonal to (the points associated to) $v$ and~$w$.
+    As~$v \neq w$, we see~$p_5$ is collinear
+    to cross-product~$v \times w$.
+    Similarly, we derive
+    \begin{align*}
+        p_1 &\coll x \times (v \times w) &
+        p_4 &\coll w \times z \\
+        p_2 &\coll x \times (x \times v \times w) &
+        p_3 & \coll w \times (w \times z) \\
+        \omit\rlap{$
+        p_3 \coll (w \times z) \times (x \times (x \times (v \times w)))$}
+    \end{align*}
+    And thus
+    \begin{equation}
+        (w \times (w \times z)) \times
+        ((w \times z) \times (x \times (x\times (v \times w)))) =0.
+            \label{eq:coll2}
+    \end{equation}
+    By rotating the embedding,
+        we may assume without loss of generality
+            that we have~$z=(0,0,1)$, $y=(0,1,0)$ and~$x=(1,0,0)$.
+    Write
+\begin{align*}
+    v&=\Bigl(v_1,0,\sqrt{1-v_1^2}\Bigr) &
+        w &= \Bigl(w_1,w_2,\sqrt{1-w_1^2-w_2^2}\Bigr).
+\end{align*}
+Now, equation~\ref{eq:coll2} becomes
+\begin{align*}
+    w_2^2 \Bigl(v_1-\sqrt{1-v_1^2}w_1\sqrt{1-w_1^2-w_2^2}\Bigr) &= 0 \\
+    w_1w_2\Bigl(\sqrt{1-v_1^2}w_1 \sqrt{1-w_1^2 - w_2^2} - v_1 \Bigr) &= 0.
+\end{align*}
+Note~$w_1 \neq 0$,
+for otherwise~$w$ would be orthogonal to~$x$,
+which would imply~$w = y$, quod non.
+Similarly~$w_2 \neq 0$, for otherwise~$w$ would be orthogonal to~$y$
+and then~$v = p_4$, quod non.
+And finally, $v_1 \neq 0$, for otherwise~$v$ would be orthogonal to~$x$
+and then~$x = p_5$, quod non.
+It follows~$v_1 - \sqrt{1-v_1^2}  w_1 \sqrt{1-w_1^2 - w_2^2} = 0$.
+Thus $v_1 = \sqrt{1-v_1^2}  w_1 \sqrt{1-w_1^2 - w_2^2}$.
+As $v$ is orthogonal to~$w$,
+we have~$-v_1 w_1 = \sqrt{1-v_1^2}\sqrt{1-w_1^2-w_2^2}$.
+Substituting this in the previous equation,
+we get~$v_1 = -v_1w_1^2$.
+That is: $w_1^2 = -1$, which is absurd.
+Apparently, the graph cannot be embeddable.
+\end{myproof}
+
 \end{document}
